### 2012-09-14 proposal/proposal.tex
initial commit of proposal file

\documentclass[a4paper, 11pt]{scrartcl}
\usepackage[utf8]{inputenc} 
\usepackage[T1]{fontenc}  
\usepackage[left=2.5cm, right=2.5cm, top=2.5cm, bottom=3cm, a4paper]{geometry}
\usepackage[pdftex]{graphics} 
\usepackage{mathptmx}
\usepackage{graphicx}

\title{\Large Out title here}
\author{\small Mattijs Meiboom, Robin Mills, Tijn Schouten, and Spyros Ioakeimidis}
\date{\small \today}

\setlength{\parindent}{0pt}

\begin{document}

\maketitle

\thispagestyle{empty}

\section{Introduction}

\end{document}

### 2012-09-14 proposal/proposal.tex
proposal update

@@
 \title{\Large Out title here}
-\author{\small Mattijs Meiboom, Robin Mills, Tijn Schouten, and Spyros Ioakeimidis}
+\author{\small Mattijs Meiboom (student number), Robin Mills (student number)\\
+			\small Tijn Schouten (student number), Spyros Ioakeimidis (2144603)}
 \date{\small \today}
@@
 
+This project is about ...
+
+\section{Application}
+
+Explain how you get data or simulator, how many examples, how many inputs, outputs, whether it is an RL, classification, or regression problem.
+
+\section{Methods}
+
+Explain the methods that you will compare. Identify the learning algorithm and the optimisation method.
+
+\section{Setup of Experiments}
+
+\section{Programming Language}
+
+The chosen programming language is Python. We have chosen Python because ...
+
+\section{Planning}
+
+Describe the different phases and the planning deadlines (until November)
+
 \end{document}

### 2012-09-14 proposal/proposal.tex
Added my student number
added student number
removed hello :), added student number
added some introduction

@@
 \title{\Large Out title here}
-\author{\small Mattijs Meiboom (student number), Robin Mills (student number)\\
-			\small Tijn Schouten (student number), Spyros Ioakeimidis (2144603)}
+\author{\small Mattijs Meiboom (s1398342), Robin Mills (student number)\\
+			\small Tijn Schouten (s1836560), Spyros Ioakeimidis (2144603)}
 \date{\small \today}
@@
 
-This project is about ...
+This project is about classifying fMRI data in order to detect the cognitive states associated with viewing different stimuli.
+Specifically, viewing sentences versus viewing pictures will evoke different patterns of activity in the brain.
+By using classifiers we will try to capture these patterns by predicting which stimuli a participant is looking at, based on the brain activity during that time.
 

### 2012-09-14 proposal/proposal.tex
minor changes
Added section on programmign language, extended gitignore

--- 2012-09-14 proposal/proposal.tex
@@
 
-\title{\Large Out title here}
-\author{\small Mattijs Meiboom (student number), Robin Mills (student number)\\
-			\small Tijn Schouten (student number), Spyros Ioakeimidis (2144603)}
+\title{\Large Bayer network classifiers for fRMI}
+\author{\small Mattijs Meiboom (s1398342), Robin Mills (s1544179)\\
+			\small Tijn Schouten (s1836560), Spyros Ioakeimidis (s2144603)}
 \date{\small \today}
@@
 
-The chosen programming language is Python. We have chosen Python because ...
+We have chosen to implement our classification algorithms using Python.
+To be specific,  the well-known SciPy library.
+SciPy is open-source software for mathematics, science, and engineering and offers similar functionality compared to MatLab.
+Even though our datasets have been stored in MatLab's \textit{.mat} format, SciPy can work with this as well.
+
+The main motivation for chosing Python over MatLab is our familiarity with Python, its performance and its clean coding style.
+As an added benefit, anyone is able to reproduce our classification results even if they don't own a MatLab license.
+
+Note that in the end, the specific choice is trivial with respect to our results. Both MatLab and Python (with some added libraries) support similar functionality.
 

### 2012-09-14 proposal/proposal.tex

@@
 
-\title{\Large Bayer network classifiers for fRMI}
+\title{\Large Bayes network classifiers for fRMI}
 \author{\small Mattijs Meiboom (s1398342), Robin Mills (s1544179)\\

### 2012-09-14 proposal/proposal.tex
minrochanges
spelling
proposal update

@@
 \section{Programming Language}
+\label{sec:language}
 
@@
 
-The main motivation for chosing Python over MatLab is our familiarity with Python, its performance and its clean coding style.
+The main motivation for choosing Python over MatLab is our familiarity with Python, its performance and its clean coding style.
 As an added benefit, anyone is able to reproduce our classification results even if they don't own a MatLab license.
@@
 
+\begin{description}
+	\item[11th - 21st September]\-\\
+	Research ...
+	\item[22nd - 28th September]\-\\
+	Model definition ...
+	\item[29th - 19th October]\-\\
+	Implementation and Parameter estimation ...
+	\item[20th - 22nd October]\-\\
+	Preparing the presentation ...
+	\item[28th October]\-\\
+	Final delivery of the document and the implementation
+\end{description}
+
 \end{document}

### 2012-09-14 proposal/proposal.tex
Added info to application section

--- 2012-09-14 proposal/proposal.tex
@@
 
-This project is about classifying fMRI data in order to detect the cognitive states associated with viewing different stimuli.
+This project is about classifying fMRI data in order to detect the cognitive states of subjects associated with viewing different stimuli.
 Specifically, viewing sentences versus viewing pictures will evoke different patterns of activity in the brain.
@@
 
-Explain how you get data or simulator, how many examples, how many inputs, outputs, whether it is an RL, classification, or regression problem.
+For our project, we are using a collection of data from the StarPlus project\footnote{StarPlus project - http://www.cs.cmu.edu/afs/cs.cmu.edu/project/theo-81/www/}.
+This dataset contains the results from a trial experiment performed at Carnegie Mellon University.
+The data was originally collected my Marcel Just \textit{et al}.
+
+The dataset contains data on 6 test subjects that each completed a number of trials.
+For each trial (see section \ref{sec:setup}), 54 fMRI snapshots were made of specific regions of their brain.
+Each snapshot consists of a list of 3d pixels (so called voxels).
+The dimensions of these snapshots are 64x64x8 voxels, amounting to a total of 5700 voxels per snapshot (this is less than 64 * 64 * 8 because only specific regions were recorded).
+
+Along with each snapshot, metadata is available on the state of the trial (e.g.
+what stimulus was active?) and what user the snapshot belongs to.
+For each entire trial, we also have data on the trial specifics and the result of the trial.
+
+Since we will be trying to classify trial snapshots to determine what stimulus was active when the snapshot was taken, our application deals with a typical classification problem.
 
@@
 \section{Setup of Experiments}
+\label{sec:setup}
 
@@
 
+22nd/24th - presentation
+31st - final document
+
 \end{document}

### 2012-09-14 proposal/proposal.tex

@@
 
-22nd/24th - presentation
-31st - final document
+\begin{description}
+	\item[11th - 21st September]\-\\
+	Research ...
+	\item[22nd - 28th September]\-\\
+	Model definition ...
+	\item[29th - 19th October]\-\\
+	Implementation and Parameter estimation ...
+	\item[20th - 22nd October]\-\\
+	Preparing the presentation ...
+	\item[28th October]\-\\
+	Final delivery of the document and the implementation
+\end{description}
 

### 2012-09-14 proposal/proposal.tex
setup of experiments

--- 2012-09-14 proposal/proposal.tex
@@
 
+Participants performed an experiment while laying in an fMRI scanner.
+The experiment consisted of sets of trials where participants were shown a first stimulus for four seconds, then a blank screen for another four seconds, and then a second stimulus for four seconds, which was followed by a rest period of 15 seconds.
+For the first half of the trials, a picture was used as first stimulus and a sentence was used as second stimulus, while the second half of trials showed a sentence as the first stimulus and a picture as the second.
+Participants had four seconds after the second stimulus appeared to indicate that the sentence and picture belonged together, after which the stimulus disappeared.
+If participants did not press a button the stimulus would disappear after four seconds.
+
 \section{Programming Language}
-\label{sec:language}
 
@@
 
-\begin{description}
-	\item[11th - 21st September]\-\\
-	Research ...
-	\item[22nd - 28th September]\-\\
-	Model definition ...
-	\item[29th - 19th October]\-\\
-	Implementation and Parameter estimation ...
-	\item[20th - 22nd October]\-\\
-	Preparing the presentation ...
-	\item[28th October]\-\\
-	Final delivery of the document and the implementation
-\end{description}
-
 \end{document}

### 2012-09-14 proposal/proposal.tex

--- 2012-09-14 proposal/proposal.tex
@@
 
+Participants performed an experiment while laying in an fMRI scanner.
+The experiment consisted of sets of trials where participants were shown a first stimulus for four seconds, then a blank screen for another four seconds, and then a second stimulus for four seconds, which was followed by a rest period of 15 seconds.
+For the first half of the trials, a picture was used as first stimulus and a sentence was used as second stimulus, while the second half of trials showed a sentence as the first stimulus and a picture as the second.
+Participants had four seconds after the second stimulus appeared to indicate that the sentence and picture belonged together, after which the stimulus disappeared.
+If participants did not press a button the stimulus would disappear after four seconds.
+
 \section{Programming Language}

### 2012-09-14 proposal/proposal.tex
planning

@@
 This dataset contains the results from a trial experiment performed at Carnegie Mellon University.
-The data was originally collected my Marcel Just \textit{et al}.
+The data was originally collected by Marcel Just \textit{et al}.
 
 The dataset contains data on 6 test subjects that each completed a number of trials.
-For each trial (see section \ref{sec:setup}), 54 fMRI snapshots were made of specific regions of their brain.
+For each trial (cf.
+Section~\ref{sec:setup}), 54 fMRI snapshots were made of specific regions of their brain.
 Each snapshot consists of a list of 3d pixels (so called voxels).
@@
 	\item[11th - 21st September]\-\\
-	Research ...
+	During this time period, the group members will conduct an initial research. The research will focus on the Bayes network classifiers and especially on the TAN tree algorithm.
 	\item[22nd - 28th September]\-\\
-	Model definition ...
+	During this time period, the definition of the model will be realised. 
 	\item[29th - 19th October]\-\\

### 2012-09-14 proposal/proposal.tex
Fixed trial numbers

--- 2012-09-14 proposal/proposal.tex
@@
 Each snapshot consists of a list of 3d pixels (so called voxels).
-The dimensions of these snapshots are 64x64x8 voxels, amounting to a total of 5700 voxels per snapshot (this is less than 64 * 64 * 8 because only specific regions were recorded).
+The dimensions of these snapshots are 64x64x8 voxels, amounting to a total of approximately 5000 voxels per snapshot (this is less than 64 * 64 * 8 because only specific regions were recorded).
 

### 2012-09-14 proposal/proposal.tex

@@
 This dataset contains the results from a trial experiment performed at Carnegie Mellon University.
-The data was originally collected my Marcel Just \textit{et al}.
+The data was originally collected by Marcel Just \textit{et al}.
 

### 2012-09-14 proposal/proposal.tex
added bib file, planning
updated application

--- 2012-09-14 proposal/proposal.tex
@@
 
-The dataset contains data on 6 test subjects that each completed a number of trials.
-For each trial (cf.
-Section~\ref{sec:setup}), 54 fMRI snapshots were made of specific regions of their brain.
-Each snapshot consists of a list of 3d pixels (so called voxels).
-The dimensions of these snapshots are 64x64x8 voxels, amounting to a total of 5700 voxels per snapshot (this is less than 64 * 64 * 8 because only specific regions were recorded).
+The dataset contains data on 6 subjects that each completed a number of trials.
+For each trial (see section \ref{sec:setup}), a brain scan was made every 500 ms, totalling to approximately 54 scans per trial.
+Each snapshot consists of a list of approximately 5000 voxels (three dimensional pixels), of predefined regions of the brain.
 
 Along with each snapshot, metadata is available on the state of the trial (e.g.
-what stimulus was active?) and what user the snapshot belongs to.
+what stimulus was shown) and what user the snapshot belongs to.
 For each entire trial, we also have data on the trial specifics and the result of the trial.

### 2012-09-14 proposal/proposal.tex
added bibliography section
Added initial section on methods

--- 2012-09-14 proposal/proposal.tex
@@
 \section{Application}
+\label{sec:application}
 
@@
 \section{Methods}
+\label{sec:methods}
 
-Explain the methods that you will compare. Identify the learning algorithm and the optimisation method.
+The dataset we use originated from a project that researched classifying cognitive state using Gaussian Naive Bayes classifiers.
+This resulted in an accuracy of approximately 85\%.
+According to literature\footnote{Bayesian network classifiers - N.
+Friedman, D.
+Geiger, and M.
+Goldszmidt}, a higher accuracy could be obtained by using a so-called \textit{Tree Augmented Naive Bayesian network classifier}, or TAN.
+We will use this classification method to investigate whether we are indeed able to reach a higher accuracy.
+
+To verify our results, we will use the original Naive Bayesian classifier as a benchmark and compare them to our TAN results.
 

### 2012-09-14 proposal/proposal.tex

--- 2012-09-14 proposal/proposal.tex
@@
 \usepackage[left=2.5cm, right=2.5cm, top=2.5cm, bottom=3cm, a4paper]{geometry}
+\usepackage{hyperref}
 \usepackage[pdftex]{graphics} 
@@
 	\item[22nd - 28th September]\-\\
-	During this time period, the definition of the model will be realised. 
+	During this time period, the definition of the model will be realised. \cite{CV:2009kl}
 	\item[29th - 19th October]\-\\
@@
 
+\nocite{*} %DO NOT UNCOMMENT, only for test purpose
+\bibliographystyle{IEEEtran}
+\bibliography{bibliography}
+
 \end{document}

### 2012-09-14 proposal/proposal.tex
added reference

--- 2012-09-14 proposal/proposal.tex
@@
 
-For our project, we are using a collection of data from the StarPlus project\footnote{StarPlus project - http://www.cs.cmu.edu/afs/cs.cmu.edu/project/theo-81/www/}.
+For our project, we are using a collection of data from the StarPlus project.
 This dataset contains the results from a trial experiment performed at Carnegie Mellon University.
-The data was originally collected by Marcel Just \textit{et al}.
+The data was originally collected by Marcel Just \textit{et al} \cite{StarPlus:2004kx}.
 
@@
 
+%\cite{Friedman:1997gw}
+\nocite{*} %DO NOT UNCOMMENT, only for test purpose
+\bibliographystyle{IEEEtran}
+\bibliography{bibliography}
+
 \end{document}

### 2012-09-14 proposal/proposal.tex

--- 2012-09-14 proposal/proposal.tex
@@
 	\item[22nd - 28th September]\-\\
-	During this time period, the definition of the model will be realised. \cite{CV:2009kl}
+	During this time period, the definition of the model will be realised.
 	\item[29th - 19th October]\-\\
@@
 
+%\cite{Friedman:1997gw}
 \nocite{*} %DO NOT UNCOMMENT, only for test purpose

### 2012-09-14 proposal/proposal.tex
bib and proposal update

@@
 This resulted in an accuracy of approximately 85\%.
-According to literature\footnote{Bayesian network classifiers - N.
-Friedman, D.
-Geiger, and M.
-Goldszmidt}, a higher accuracy could be obtained by using a so-called \textit{Tree Augmented Naive Bayesian network classifier}, or TAN.
+According to literature \cite{Friedman:1997gw}, a higher accuracy could be obtained by using a so-called \textit{Tree Augmented Naive Bayesian network classifier}, or TAN.
 We will use this classification method to investigate whether we are indeed able to reach a higher accuracy.
@@
 
-Describe the different phases and the planning deadlines (until November)
+In this section we describe the various phases and the planning deadlines that the group will follow.
 
@@
 	\item[22nd - 28th September]\-\\
-	During this time period, the definition of the model will be realised.
+	During this time period, the definition of the various models will be realised.
 	\item[29th - 19th October]\-\\
@@
 	\item[20th - 22nd October]\-\\
-	Preparing the presentation ...
+	During this time period the group members will prepare the presentation of the project.
 	\item[28th October]\-\\
-	Final delivery of the document and the implementation
+	Until the aforementioned date the final version of the document and the implementation will be delivered.
 \end{description}
 
-%\cite{Friedman:1997gw}
 \nocite{*} %DO NOT UNCOMMENT, only for test purpose

### 2012-09-14 proposal/proposal.tex
spelling and writing stuff

--- 2012-09-14 proposal/proposal.tex
@@
 \section{Application}
+\label{sec:application}
 
-For our project, we are using a collection of data from the StarPlus project.
+For our project, we are using a collection of data from the StarPlus project\footnote{StarPlus project - http://www.cs.cmu.edu/afs/cs.cmu.edu/project/theo-81/www/}.
 This dataset contains the results from a trial experiment performed at Carnegie Mellon University.
-The data was originally collected by Marcel Just \textit{et al} \cite{StarPlus:2004kx}.
+The data was originally collected by Marcel Just \textit{et al}.
 
@@
 For each trial (see section \ref{sec:setup}), a brain scan was made every 500 ms, totalling to approximately 54 scans per trial.
-Each snapshot consists of a list of approximately 5000 voxels (three dimensional pixels), of predefined regions of the brain.
+Each scan consists of a list of approximately 5000 voxels (three dimensional pixels), of predefined regions of the brain.
+Along with each trial, metadata is available on the type of trial (e.g.
+what stimulus was shown), the result of the trial, and to which subject the scan belongs.
 
-Along with each snapshot, metadata is available on the state of the trial (e.g.
-what stimulus was shown) and what user the snapshot belongs to.
-For each entire trial, we also have data on the trial specifics and the result of the trial.
-
-Since we will be trying to classify trial snapshots to determine what stimulus was active when the snapshot was taken, our application deals with a typical classification problem.
+Since we will be trying to classify brain patterns to determine what stimulus was active when the scan was taken, our application deals with a typical classification problem.
 
 \section{Methods}
+\label{sec:methods}
 
-Explain the methods that you will compare. Identify the learning algorithm and the optimisation method.
+The dataset we use originated from a project that investigated classifying cognitive state using Gaussian Naive Bayes classifiers.
+This resulted in an accuracy of approximately 85\%.
+According to literature, a higher accuracy could be obtained by using a so-called \textit{Tree Augmented Naive Bayesian network classifier}, or TAN\footnote{Bayesian network classifiers - N.
+Friedman, D.
+Geiger, and M.
+Goldszmidt}.
+We will use this classification method to investigate whether we are indeed able to reach a higher accuracy.
+
+To verify our results, we will use the original Naive Bayesian classifier as a benchmark and compare it to our TAN results.
 
@@
 	\item[22nd - 28th September]\-\\
-	During this time period, the definition of the model will be realised.
+	During this time period, the definition of the model will be realised. \cite{CV:2009kl}
 	\item[29th - 19th October]\-\\
@@
 
-%\cite{Friedman:1997gw}
 \nocite{*} %DO NOT UNCOMMENT, only for test purpose

### 2012-09-14 proposal/proposal.tex

--- 2012-09-14 proposal/proposal.tex
@@
 For each trial (see section \ref{sec:setup}), a brain scan was made every 500 ms, totalling to approximately 54 scans per trial.
-Each snapshot consists of a list of approximately 5000 voxels (three dimensional pixels), of predefined regions of the brain.
+Each scan consists of a list of approximately 5000 voxels (three dimensional pixels), of predefined regions of the brain.
+Along with each trial, metadata is available on the type of trial (e.g.
+what stimulus was shown), the result of the trial, and to which subject the scan belongs.
 
-Along with each snapshot, metadata is available on the state of the trial (e.g.
-what stimulus was shown) and what user the snapshot belongs to.
-For each entire trial, we also have data on the trial specifics and the result of the trial.
-
-Since we will be trying to classify trial snapshots to determine what stimulus was active when the snapshot was taken, our application deals with a typical classification problem.
+Since we will be trying to classify brain patterns to determine what stimulus was active when the scan was taken, our application deals with a typical classification problem.
 
@@
 This resulted in an accuracy of approximately 85\%.
-According to literature \cite{Friedman:1997gw}, a higher accuracy could be obtained by using a so-called \textit{Tree Augmented Naive Bayesian network classifier}, or TAN.
+According to literature, a higher accuracy could be obtained by using a so-called \textit{Tree Augmented Naive Bayesian network classifier}, or TAN\cite{Friedman:1997gw}.
 We will use this classification method to investigate whether we are indeed able to reach a higher accuracy.
 
-To verify our results, we will use the original Naive Bayesian classifier as a benchmark and compare them to our TAN results.
+To verify our results, we will use the original Naive Bayesian classifier as a benchmark and compare it to our TAN results.
 

### 2012-09-14 proposal/proposal.tex
proposal update
spelling

@@
 
-\title{\Large Bayes network classifiers for fRMI}
+\title{\Large Bayes network classifiers for fMRI}
 \author{\small Mattijs Meiboom (s1398342), Robin Mills (s1544179)\\
@@
 The dataset contains data on 6 subjects that each completed a number of trials.
-For each trial (see section \ref{sec:setup}), a brain scan was made every 500 ms, totalling to approximately 54 scans per trial.
+For each trial (cf.
+Section~\ref{sec:setup}), a brain scan was made every 500 ms, totalling to approximately 54 scans per trial.
 Each scan consists of a list of approximately 5000 voxels (three dimensional pixels), of predefined regions of the brain.
@@
 This resulted in an accuracy of approximately 85\%.
-According to literature, a higher accuracy could be obtained by using a so-called \textit{Tree Augmented Naive Bayesian network classifier}, or TAN\cite{Friedman:1997gw}.
+According to literature, a higher accuracy could be obtained by using a so-called \textit{Tree Augmented Naive Bayesian network classifier}, or TAN \cite{Friedman:1997gw}.
 We will use this classification method to investigate whether we are indeed able to reach a higher accuracy.

### 2012-09-14 proposal/proposal.tex
update section

--- 2012-09-14 proposal/proposal.tex
@@
 For each trial (see section \ref{sec:setup}), a brain scan was made every 500 ms, totalling to approximately 54 scans per trial.
-Each scan consists of a list of approximately 5000 voxels (three dimensional pixels), of predefined regions of the brain.
-Along with each trial, metadata is available on the type of trial (e.g.
-what stimulus was shown), the result of the trial, and to which subject the scan belongs.
+Each snapshot consists of a list of approximately 5000 voxels (three dimensional pixels), of predefined regions of the brain.
 
-Since we will be trying to classify brain patterns to determine what stimulus was active when the scan was taken, our application deals with a typical classification problem.
+Along with each snapshot, metadata is available on the state of the trial (e.g.
+what stimulus was shown) and what user the snapshot belongs to.
+For each entire trial, we also have data on the trial specifics and the result of the trial.
+
+Since we will be trying to classify trial snapshots to determine what stimulus was active when the snapshot was taken, our application deals with a typical classification problem.
 
@@
 This resulted in an accuracy of approximately 85\%.
-According to literature, a higher accuracy could be obtained by using a so-called \textit{Tree Augmented Naive Bayesian network classifier}, or TAN\cite{Friedman:1997gw}.
+According to literature \cite{Friedman:1997gw}, a higher accuracy could be obtained by using a so-called \textit{Tree Augmented Naive Bayesian network classifier}, or TAN.
 We will use this classification method to investigate whether we are indeed able to reach a higher accuracy.
 
-To verify our results, we will use the original Naive Bayesian classifier as a benchmark and compare it to our TAN results.
+To verify our results, we will use the original Naive Bayesian classifier as a benchmark and compare them to our TAN results.
 
@@
 Participants performed an experiment while laying in an fMRI scanner.
-The experiment consisted of sets of trials where participants were shown a first stimulus for four seconds, then a blank screen for another four seconds, and then a second stimulus for four seconds, which was followed by a rest period of 15 seconds.
-For the first half of the trials, a picture was used as first stimulus and a sentence was used as second stimulus, while the second half of trials showed a sentence as the first stimulus and a picture as the second.
+The experiment consisted of sets of trials where participants were presented a first stimulus for four seconds, followed by a blank screen lasting four seconds, and then a second stimulus for four seconds, followed by a rest period of 15 seconds.
+For the first half of the trials, a picture was presented as the first stimulus and a sentence as second stimulus, while for the second half of trials, a sentencewas presented as the first stimulus and a picture as the second.
 Participants had four seconds after the second stimulus appeared to indicate that the sentence and picture belonged together, after which the stimulus disappeared.

### 2012-09-14 proposal/proposal.tex
proposal section

@@
 The experiment consisted of sets of trials where participants were presented a first stimulus for four seconds, followed by a blank screen lasting four seconds, and then a second stimulus for four seconds, followed by a rest period of 15 seconds.
-For the first half of the trials, a picture was presented as the first stimulus and a sentence as second stimulus, while for the second half of trials, a sentencewas presented as the first stimulus and a picture as the second.
+For the first half of the trials, a picture was presented as the first stimulus and a sentence as the second stimulus, while for the second half of trials, a sentence was presented as the first stimulus and a picture as the second.
 Participants had four seconds after the second stimulus appeared to indicate that the sentence and picture belonged together, after which the stimulus disappeared.

### 2012-09-14 proposal/proposal.tex

--- 2012-09-14 proposal/proposal.tex
@@
 Participants performed an experiment while laying in an fMRI scanner.
-The experiment consisted of sets of trials where participants were shown a first stimulus for four seconds, then a blank screen for another four seconds, and then a second stimulus for four seconds, which was followed by a rest period of 15 seconds.
-For the first half of the trials, a picture was used as first stimulus and a sentence was used as second stimulus, while the second half of trials showed a sentence as the first stimulus and a picture as the second.
+The experiment consisted of sets of trials where participants were presented a first stimulus for four seconds, followed by a blank screen lasting four seconds, and then a second stimulus for four seconds, followed by a rest period of 15 seconds.
+For the first half of the trials, a picture was presented as the first stimulus and a sentence as the second stimulus, while for the second half of trials, a sentence was presented as the first stimulus and a picture as the second.
 Participants had four seconds after the second stimulus appeared to indicate that the sentence and picture belonged together, after which the stimulus disappeared.

### 2012-09-14 proposal/proposal.tex
prop sec

@@
 Participants performed an experiment while laying in an fMRI scanner.
-The experiment consisted of sets of trials where participants were presented a first stimulus for four seconds, followed by a blank screen lasting four seconds, and then a second stimulus for four seconds, followed by a rest period of 15 seconds.
+The experiment consisted of a sets of trials where participants were presented a first stimulus for four seconds, followed by a blank screen lasting four seconds, and then a second stimulus for four seconds, followed by a rest period of 15 seconds.
 For the first half of the trials, a picture was presented as the first stimulus and a sentence as the second stimulus, while for the second half of trials, a sentence was presented as the first stimulus and a picture as the second.

### 2012-09-14 proposal/proposal.tex
small spelling
Added citation instead of footnote.

--- 2012-09-14 proposal/proposal.tex
@@
 
-For our project, we are using a collection of data from the StarPlus project.
+For our project, we are using a collection of data from the StarPlus project\footnote{StarPlus project - http://www.cs.cmu.edu/afs/cs.cmu.edu/project/theo-81/www/}.
 This dataset contains the results from a trial experiment performed at Carnegie Mellon University.
-The data was originally collected by Marcel Just \textit{et al} \cite{StarPlus:2004kx}.
+The data was originally collected by Marcel Just \textit{et al}.
 
@@
 This resulted in an accuracy of approximately 85\%.
-According to literature \cite{Friedman:1997gw}, a higher accuracy could be obtained by using a so-called \textit{Tree Augmented Naive Bayesian network classifier}, or TAN.
+According to literature\cite{Friedman:1997gw}, a higher accuracy could be obtained by using a so-called \textit{Tree Augmented Naive Bayesian network classifier}, or TAN.
 We will use this classification method to investigate whether we are indeed able to reach a higher accuracy.
@@
 Participants performed an experiment while laying in an fMRI scanner.
-The experiment consisted of sets of trials where participants were presented a first stimulus for four seconds, followed by a blank screen lasting four seconds, and then a second stimulus for four seconds, followed by a rest period of 15 seconds.
-For the first half of the trials, a picture was presented as the first stimulus and a sentence as the second stimulus, while for the second half of trials, a sentence was presented as the first stimulus and a picture as the second.
+The experiment consisted of sets of trials where participants were shown a first stimulus for four seconds, then a blank screen for another four seconds, and then a second stimulus for four seconds, which was followed by a rest period of 15 seconds.
+For the first half of the trials, a picture was used as first stimulus and a sentence was used as second stimulus, while the second half of trials showed a sentence as the first stimulus and a picture as the second.
 Participants had four seconds after the second stimulus appeared to indicate that the sentence and picture belonged together, after which the stimulus disappeared.
@@
 
-In this section we describe the various phases and the planning deadlines that the group will follow.
+Describe the different phases and the planning deadlines (until November)
 
@@
 	\item[22nd - 28th September]\-\\
-	During this time period, the definition of the various models will be realised.
+	During this time period, the definition of the model will be realised. \cite{CV:2009kl}
 	\item[29th - 19th October]\-\\
@@
 	\item[20th - 22nd October]\-\\
-	During this time period the group members will prepare the presentation of the project.
+	Preparing the presentation ...
 	\item[28th October]\-\\
-	Until the aforementioned date the final version of the document and the implementation will be delivered.
+	Final delivery of the document and the implementation
 \end{description}

### 2012-09-14 proposal/proposal.tex

--- 2012-09-14 proposal/proposal.tex
@@
 Along with each trial, metadata is available on the type of trial (e.g.
-what stimulus was shown), the result of the trial, and to which subject the scan belongs.
+what stimulus was shown), the result of the trial, and to which subject the scans belong.
 
@@
 This resulted in an accuracy of approximately 85\%.
-According to literature, a higher accuracy could be obtained by using a so-called \textit{Tree Augmented Naive Bayesian network classifier}, or TAN \cite{Friedman:1997gw}.
+According to literature\cite{Friedman:1997gw}, a higher accuracy could be obtained by using a so-called \textit{Tree Augmented Naive Bayesian network classifier}, or TAN.
 We will use this classification method to investigate whether we are indeed able to reach a higher accuracy.

### 2012-09-14 proposal/proposal.tex
minor changes

--- 2012-09-14 proposal/proposal.tex
@@
 	\item[29th - 19th October]\-\\
-	Implementation and Parameter estimation ...
+	Implementation and Parameter estimation.
 	\item[20th - 22nd October]\-\\
@@
 	\item[28th October]\-\\
-	Until the aforementioned date the final version of the document and the implementation will be delivered.
+	Until this date the final version of the document and the implementation will be delivered.
 \end{description}

### 2012-09-14 proposal/proposal.tex
Minor typo/grammar issues

--- 2012-09-14 proposal/proposal.tex
@@
 
-This project is about classifying fMRI data in order to detect the cognitive states of subjects associated with viewing different stimuli.
+The proposed project is about classifying fMRI data in order to detect the cognitive states of subjects associated with viewing different stimuli.
 Specifically, viewing sentences versus viewing pictures will evoke different patterns of activity in the brain.
-By using classifiers we will try to capture these patterns by predicting which stimuli a participant is looking at, based on the brain activity during that time.
+By using classifiers we will try to capture these patterns by predicting what stimulus a participant is looking at, based on the brain activity during that time.
 

### 2012-09-14 proposal/proposal.tex
minor changes

--- 2012-09-14 proposal/proposal.tex
@@
 This resulted in an accuracy of approximately 85\%.
-According to literature\cite{Friedman:1997gw}, a higher accuracy could be obtained by using a so-called \textit{Tree Augmented Naive Bayesian network classifier}, or TAN.
+According to literature \cite{Friedman:1997gw}, a higher accuracy could be obtained by using a so-called \textit{Tree Augmented Naive Bayesian network classifier}, or TAN.
 We will use this classification method to investigate whether we are indeed able to reach a higher accuracy.

### 2012-09-15 proposal/proposal.tex
small writing

@@
 Section~\ref{sec:setup}), a brain scan was made every 500 ms, totalling to approximately 54 scans per trial.
-Each scan consists of a list of approximately 5000 voxels (three dimensional pixels), of predefined regions of the brain.
+Each scan consists of a list of approximately 5000 voxels (three dimensional pixels), containing information on the activation of predefined regions of the brain.
 Along with each trial, metadata is available on the type of trial (e.g.
@@
 To be specific,  the well-known SciPy library.
-SciPy is open-source software for mathematics, science, and engineering and offers similar functionality compared to MatLab.
-Even though our datasets have been stored in MatLab's \textit{.mat} format, SciPy can work with this as well.
+SciPy is open-source software for mathematics, science, and engineering and offers similar functionality as MatLab.
+Additionally, SciPy supports the MatLab's \textit{.mat} format, in which our data is stored.
 
 The main motivation for choosing Python over MatLab is our familiarity with Python, its performance and its clean coding style.
-As an added benefit, anyone is able to reproduce our classification results even if they don't own a MatLab license.
+As an added benefit, no MatLab license is required to reproduce our results.
 

### 2012-09-16 proposal/proposal.tex
writing

@@
 
-The proposed project is about classifying fMRI data in order to detect the cognitive states of subjects associated with viewing different stimuli.
+The proposed project is about classifying fMRI data in order to detect the cognitive states that are associated with viewing different stimuli.
 Specifically, viewing sentences versus viewing pictures will evoke different patterns of activity in the brain.
@@
 This dataset contains the results from a trial experiment performed at Carnegie Mellon University.
-The data was originally collected by Marcel Just \textit{et al} \cite{StarPlus:2004kx}.
+The data was originally collected by Marcel Just \textit{et a.l} \cite{StarPlus:2004kx}.
 
@@
 Participants performed an experiment while laying in an fMRI scanner.
-The experiment consisted of a sets of trials where participants were presented a first stimulus for four seconds, followed by a blank screen lasting four seconds, and then a second stimulus for four seconds, followed by a rest period of 15 seconds.
+The experiment consisted of a set of trials in which participants were presented a first stimulus for four seconds, followed by a blank screen lasting four seconds, and then a second stimulus for four seconds, followed by a rest period of 15 seconds.
 For the first half of the trials, a picture was presented as the first stimulus and a sentence as the second stimulus, while for the second half of trials, a sentence was presented as the first stimulus and a picture as the second.

### 2012-09-17 proposal/proposal.tex
minor changes in proposal
proposal update

@@
 Participants had four seconds after the second stimulus appeared to indicate that the sentence and picture belonged together, after which the stimulus disappeared.
-If participants did not press a button the stimulus would disappear after four seconds.
+If participants did not press the button the stimulus would disappear after those four seconds.
 
@@
 	\item[11th - 21st September]\-\\
-	During this time period, the group members will conduct an initial research. The research will focus on the Bayes network classifiers and especially on the TAN tree algorithm.
+	During this time period, the group members will conduct an initial research.
+The research will focus on the Bayes network classifiers and especially on the TAN tree algorithm.
+The purpose of this phase is to understand the problem.
 	\item[22nd - 28th September]\-\\
-	During this time period, the definition of the various models will be realised.
+	During this time period, the definition of the various models will be realised. A set of models is needed to approximate the output function.
 	\item[29th - 19th October]\-\\
-	Implementation and Parameter estimation.
+	Choosing and implementing a learning algorithm. This phase contains also the estimation of the parameters of the learning algorithm.
 	\item[20th - 22nd October]\-\\

### 2012-09-17 proposal/proposal.tex
proposal update

@@
 This dataset contains the results from a trial experiment performed at Carnegie Mellon University.
-The data was originally collected by Marcel Just \textit{et a.l} \cite{StarPlus:2004kx}.
+The data was originally collected by Marcel Just \emph{et al.} \cite{StarPlus:2004kx}.
 
@@
 This resulted in an accuracy of approximately 85\%.
-According to literature \cite{Friedman:1997gw}, a higher accuracy could be obtained by using a so-called \textit{Tree Augmented Naive Bayesian network classifier}, or TAN.
+According to literature \cite{Friedman:1997gw}, a higher accuracy could be obtained by using a so-called \emph{Tree Augmented Naive Bayesian network classifier}, or TAN.
 We will use this classification method to investigate whether we are indeed able to reach a higher accuracy.
@@
 Participants had four seconds after the second stimulus appeared to indicate that the sentence and picture belonged together, after which the stimulus disappeared.
-If participants did not press the button the stimulus would disappear after those four seconds.
+If participants did not press the button the stimulus would disappear after four seconds.
 
@@
 SciPy is open-source software for mathematics, science, and engineering and offers similar functionality as MatLab.
-Additionally, SciPy supports the MatLab's \textit{.mat} format, in which our data is stored.
+Additionally, SciPy supports the MatLab's \emph{.mat} format, in which our data is stored.
 
@@
 \section{Planning}
+\label{sec:planning}
 
@@
 	\item[29th - 19th October]\-\\
-	Choosing and implementing a learning algorithm. This phase contains also the estimation of the parameters of the learning algorithm.
+	During this time period we will choose and implement a learning algorithm. Furthermore, this phase contains the estimation of the parameters of the algorithm.
 	\item[20th - 22nd October]\-\\

### 2012-09-19 article/article.tex
initial commit article

\documentclass[preprint,journal]{vgtc}
\let\ifpdf\relax
%% Uncomment one of the lines above depending on where your paper is
%% in the conference process. ``review'' and ``widereview'' are for review
%% submission, ``preprint'' is for pre-publication, and the final version
%% doesn't use a specific qualifier. Further, ``electronic'' includes
%% hyperreferences for more convenient online viewing.

%% Please use one of the ``review'' options in combination with the
%% assigned online id (see below) ONLY if your paper uses a double blind
%% review process. Some conferences, like IEEE Vis and InfoVis, have NOT
%% in the past.

%% Please note that the use of figures is not permitted on the first page
%% of the journal version.  Figures should begin on the second page and be
%% in CMYK or Grey scale format, otherwise, colour shifting may occur
%% during the printing process.  Papers submitted with figures on the
%% first page will be refused.

%% These three lines bring in essential packages: ``mathptmx'' for Type 1
%% typefaces, ``graphicx'' for inclusion of EPS figures. and ``times''
%% for proper handling of the times font family.

\usepackage{mathptmx}
\usepackage{graphicx,xcolor,booktabs, tabularx,amsmath}
\usepackage{times}
\newcommand{\todo}[1]{\textbf{\textcolor{blue}{Todo: #1 }}}
%% We encourage the use of mathptmx for consistent usage of times font
%% throughout the proceedings. However, if you encounter conflicts
%% with other math-related packages, you may want to disable it.

%% If you are submitting a paper to a conference for review with a double
%% blind reviewing process, please replace the value ``0'' below with your
%% OnlineID. Otherwise, you may safely leave it at ``0''.
\onlineid{0}

%% declare the category of your paper, only shown in review mode
\vgtccategory{}

%% allow for this line if you want the electronic option to work properly
%\vgtcinsertpkg

%% In preprint mode you may define your own headline.
\preprinttext{}

%% Paper title.

\title{Bayes network classifiers for fMRI}

%% This is how authors are specified in the journal style

%% indicate IEEE Member or Student Member in form indicated below
\author{Mattijs Meiboom, Robin Mills, Tijn Schouten, Spyros Ioakeimidis}
\authorfooter{
\item
  Mattijs Meiboom is with University of Groningen, E-mail: .
\item
  Robin Mills is with University of Groningen, E-mail: .
\item
  Tijn Schouten is University of Groningen, E-mail: .
\item
  Spyros Ioakeimidis  is with University of Groningen, E-mail: spyrosikmd@gmail.com.
}

%other entries to be set up for journal
%\shortauthortitle{}

%% Abstract section.
\abstract{
 This is the abstract section...
} % end of abstract

%% Keywords that describe your work. Will show as 'Index Terms' in journal
%% please capitalize first letter and insert punctuation after last keyword
\keywords{}

%% Copyright space is enabled by default as required by guidelines.
%% It is disabled by the 'review' option or via the following command:
 \nocopyrightspace

\begin{document}

%% The ``\maketitle'' command must be the first command after the
%% ``\begin{document}'' command. It prepares and prints the title block.

%% the only exception to this rule is the \firstsection command
\firstsection{Introduction}

\maketitle

\bibliographystyle{abbrv}
%%use following if all content of bibtex file should be shown
\nocite{*}
\bibliography{references}
\end{document}

### 2012-09-20 src/framework.py
initial commit of the source file, ignore .mat files

@@ -1,0 +1,28 @@
+#!/usr/bin/env python
+
+import scipy.io as sio
+
+subject = sio.loadmat('data-starplus-04799-v7.mat')
+
+#                         trial scan voxel
+# subject_data['data'][0] [t]   [s]  [v]
+#print subject['data'][0][0][1][2]
+
+NUMBER_OF_TRIALS = 54
+NUMBER_OF_VOXELS = 4949
+
+def get_valid_trials(subject):
+    '''Returns the valid indexes of the trials for a subject.'''
+    return [i for i in range(NUMBER_OF_TRIALS)
+                if subject['info'][0]['cond'][i] > 1 ]
+
+#print get_valid_trials(subject)
+
+ROIs = ['CALC', 'LIPL', 'LT', 'LTRIA', 'LOPER', 'LIPS', 'LDLPFC']
+
+def get_rois_index(subject):
+    '''Returns the valid indexes of the voxels for a subject.'''
+    return [i for i in range(NUMBER_OF_VOXELS)
+                if subject['meta']['colToROI'][0][0][i] in ROIs]
+    
+print len(get_rois_index(subject))

### 2012-09-21 src/framework.py
update framework example

@@ -2,27 +2,39 @@
 
 import scipy.io as sio
 
-subject = sio.loadmat('data-starplus-04799-v7.mat')
+'''This is the structure of the data corresponding to every subject.
+t = Trial
+s = Scan
+v = Voxel
+subject['data'][t][0][s][v]
 
-#                         trial scan voxel
-# subject_data['data'][0] [t]   [s]  [v]
-#print subject['data'][0][0][1][2]
+For example if we want to access the value of the 1567th voxel of the 20th trial and the 11th scan then we would have the following:
 
-NUMBER_OF_TRIALS = 54
-NUMBER_OF_VOXELS = 4949
+subject['data'][19][0][10][1566]'''
+
+NUM_OF_TRIALS = 54
+NUM_OF_VOXELS = 4949
+LOW_SCAN_INDEX = 4
+HIGH_SCAN_INDEX = 14
+ROIS = ['CALC', 'LIPL', 'LT', 'LTRIA', 'LOPER', 'LIPS', 'LDLPFC']
 
 def get_valid_trials(subject):
     '''Returns the valid indexes of the trials for a subject.'''
-    return [i for i in range(NUMBER_OF_TRIALS)
-                if subject['info'][0]['cond'][i] > 1 ]
+    return (index for index in range(NUM_OF_TRIALS)
+                if subject['info'][0]['cond'][index] > 1)
 
-#print get_valid_trials(subject)
-
-ROIs = ['CALC', 'LIPL', 'LT', 'LTRIA', 'LOPER', 'LIPS', 'LDLPFC']
-
-def get_rois_index(subject):
+def get_valid_voxels(subject):
     '''Returns the valid indexes of the voxels for a subject.'''
-    return [i for i in range(NUMBER_OF_VOXELS)
-                if subject['meta']['colToROI'][0][0][i] in ROIs]
+    return (index for index in range(NUM_OF_VOXELS)
+                if subject['meta']['colToROI'][0][0][index] in ROIS)
+
+def run():
+    subject = sio.loadmat('data-starplus-04799-v7.mat')
+    values = []
+    for ti in get_valid_trials(subject):
+        for si in range(LOW_SCAN_INDEX, HIGH_SCAN_INDEX):
+            for vi in get_valid_voxels(subject):
+                values.append(subject['data'][ti][0][si][vi])
     
-print len(get_rois_index(subject))
+if __name__ == "__main__":
+    run()

### 2012-09-23 src/framework.py
initial commit on naive bayes classifier

@@ -1,40 +1,16 @@
 #!/usr/bin/env python
 
 import scipy.io as sio
+from model import NaiveBayes
 
-'''This is the structure of the data corresponding to every subject.
-t = Trial
-s = Scan
-v = Voxel
-subject['data'][t][0][s][v]
+subject = sio.loadmat('data-starplus-04799-v7.mat')
 
-For example if we want to access the value of the 1567th voxel of the 20th trial and the 11th scan then we would have the following:
+naive_bayes = NaiveBayes(subject)
 
-subject['data'][19][0][10][1566]'''
+naive_bayes.train()
 
-NUM_OF_TRIALS = 54
-NUM_OF_VOXELS = 4949
-LOW_SCAN_INDEX = 4
-HIGH_SCAN_INDEX = 14
-ROIS = ['CALC', 'LIPL', 'LT', 'LTRIA', 'LOPER', 'LIPS', 'LDLPFC']
+#subject2 = sio.loadmat('data-starplus-04820-v7.mat')
 
-def get_valid_trials(subject):
-    '''Returns the valid indexes of the trials for a subject.'''
-    return (index for index in range(NUM_OF_TRIALS)
-                if subject['info'][0]['cond'][index] > 1)
+#scan = subject['data'][10][0][16]
 
-def get_valid_voxels(subject):
-    '''Returns the valid indexes of the voxels for a subject.'''
-    return (index for index in range(NUM_OF_VOXELS)
-                if subject['meta']['colToROI'][0][0][index] in ROIS)
-
-def run():
-    subject = sio.loadmat('data-starplus-04799-v7.mat')
-    values = []
-    for ti in get_valid_trials(subject):
-        for si in range(LOW_SCAN_INDEX, HIGH_SCAN_INDEX):
-            for vi in get_valid_voxels(subject):
-                values.append(subject['data'][ti][0][si][vi])
-    
-if __name__ == "__main__":
-    run()
+#naive_bayes.classify(scan)

### 2012-09-23 src/model.py

@@ -1,0 +1,131 @@
+'''
+This is the structure of the data corresponding to every subject.
+t = Trial
+s = Scan
+v = Voxel
+subject['data'][t][0][s][v]
+
+For example if we want to access the value of the 1567th voxel of the 20th trial
+and the 11th scan then we would have the following:
+
+subject['data'][19][0][10][1566]
+'''
+
+from __future__ import division
+import math
+from collections import namedtuple
+
+Observation = namedtuple('Observation', 'klass voxels')
+CondProbability = namedtuple('CondProbability', 'klass value')
+
+class NaiveBayes:
+
+	def __init__(self, subject):
+		self.subject = subject
+		self.num_of_trials = 0
+		self.num_of_voxels = 0
+		self.valid_trial_indexes = 0
+		self.valid_voxel_indexes = 0
+		self.features = []
+		self.priori_probabilities = {}
+		self.conditional_probabilities = []
+		self.rois = ['CALC', 'LIPL', 'LT', 'LTRIA', 'LOPER', 'LIPS', 'LDLPFC']
+		self.first_stimulus_index = 0
+		self.second_stimulus_index = 16
+
+	def _get_valid_trials(self):
+		'''Returns the valid indexes of trials for a subject accoring to condition.
+		We care only about conditions with value 2 or 3.'''
+		return [index for index in range(self.num_of_trials)
+					if self.subject['info'][0]['cond'][index] > 1]
+
+	def _get_valid_voxels(self):
+		'''Returns the valid indexes of voxels for a subject according to rois'''
+		return [index for index in range(self.num_of_voxels)
+					if self.subject['meta']['colToROI'][0][0][index] in self.rois]
+
+	def _get_voxel_values(self, trial_indexes, stimulus, voxel):
+		'''Returns the absolute value of the same index voxels.'''
+		return [math.fabs(self.subject['data'][trial_index][0][stimulus][voxel]) for trial_index in trial_indexes]
+
+	def _get_voxel_vector(self, trial_index, stimulus):
+		'''Returns the voxel vector data for the trial with index trial_index
+		and scan index as defined by stimulus.'''
+		return self.subject['data'][trial_index][0][stimulus]
+
+	def _get_first_stimulus(self, trial_index):
+		'''Returns the first stimulus of a trial with index trial_index.'''
+		return self.subject['info'][0][trial_index]['firstStimulus'][0]
+
+	def _get_voxels_of_same_index(self, voxel_index, c):
[…]

### 2012-09-23 src/model.py
naive bayes

@@ -97,7 +97,7 @@ class NaiveBayes:
 		self.priori_probabilities['S'] = total_sentence_observations / total_observations
 
 	def _compute_conditional_probabilities(self):
-		'''Computes the conditional probabilities of the each voxel. Only for these voxesl that
+		'''Computes the conditional probabilities of each voxel. Only for these voxels that
 		we define as valid. '''
 		total_sum_picture = 0
 		total_sum_sentence = 0
@@ -120,12 +120,12 @@ class NaiveBayes:
 		self.conditional_probabilities.append(sen_cond_probabilities)
 
 	def train(self):
-		''''''
+		'''Trains the classifier.'''
 		self._init_values()
 		self._extract_values()
 		self._compute_priori_probabilities()
 		self._compute_conditional_probabilities()
 
 	def classify(self, scan):
-		''''''
+		'''Classifies a new scan of data.'''
 		pass

### 2012-09-23 src/framework.py
naive bayes

@@ -13,4 +13,4 @@ naive_bayes.train()
 
 #scan = subject['data'][10][0][16]
 
-#naive_bayes.classify(scan)
+naive_bayes.classify()

### 2012-09-23 src/model.py

@@ -12,8 +12,8 @@ subject['data'][19][0][10][1566]
 '''
 
 from __future__ import division
-import math
 from collections import namedtuple
+import math
 
 Observation = namedtuple('Observation', 'klass voxels')
 CondProbability = namedtuple('CondProbability', 'klass value')
@@ -40,7 +40,7 @@ class NaiveBayes:
 					if self.subject['info'][0]['cond'][index] > 1]
 
 	def _get_valid_voxels(self):
-		'''Returns the valid indexes of voxels for a subject according to rois'''
+		'''Returns the valid indexes of voxels for a subject according to rois.'''
 		return [index for index in range(self.num_of_voxels)
 					if self.subject['meta']['colToROI'][0][0][index] in self.rois]
 
@@ -126,6 +126,17 @@ class NaiveBayes:
 		self._compute_priori_probabilities()
 		self._compute_conditional_probabilities()
 
-	def classify(self, scan):
-		'''Classifies a new scan of data.'''
-		pass
+	def classify(self):
+		'''Classifies a new scan of data. The following code is just an example as for classification
+		training data are used. This needs to be modified.'''
+		scan = self._get_voxel_vector(trial_index=4, stimulus=0) # It must show `Picture'
+		valid_scan = []
+		for index, trial_index in enumerate(self.valid_voxel_indexes):
+			valid_scan.append(math.fabs(scan[trial_index]))
+		p_picture_scan = math.log(self.priori_probabilities['P']) + \
+							math.fsum([voxel_value * math.log(self.conditional_probabilities[0][index][1]) for index, voxel_value in enumerate(valid_scan)])
+		p_sentence_scan = math.log(self.priori_probabilities['S']) + \
+							math.fsum([voxel_value * math.log(self.conditional_probabilities[1][index][1]) for index, voxel_value in enumerate(valid_scan)])
+		print "P(P|Scan): ", p_picture_scan
+		print "P(S|Scan): ", p_sentence_scan
+		print "Max: ", max(p_picture_scan, p_sentence_scan)

### 2012-09-23 src/model.py
naive bayes

@@ -22,10 +22,10 @@ class NaiveBayes:
 
 	def __init__(self, subject):
 		self.subject = subject
-		self.num_of_trials = 0
-		self.num_of_voxels = 0
-		self.valid_trial_indexes = 0
-		self.valid_voxel_indexes = 0
+		self.num_of_trials = None
+		self.num_of_voxels = None
+		self.valid_trial_indexes = None
+		self.valid_voxel_indexes = None
 		self.features = []
 		self.priori_probabilities = {}
 		self.conditional_probabilities = []

### 2012-09-23 src/model.py
naive bayes

@@ -16,7 +16,7 @@ from collections import namedtuple
 import math
 
 Observation = namedtuple('Observation', 'klass voxels')
-CondProbability = namedtuple('CondProbability', 'klass value')
+ConditionalProbability = namedtuple('ConditionalProbability', 'klass value')
 
 class NaiveBayes:
 
@@ -111,10 +111,10 @@ class NaiveBayes:
 		sen_cond_probabilities = []
 		for index, voxel_index in enumerate(self.valid_voxel_indexes):
 			pic_voxels_sum = math.fsum(self._get_voxels_of_same_index(voxel_index=voxel_index, c='P'))
-			cond_prob = CondProbability('P', pic_voxels_sum / total_sum_picture)
+			cond_prob = ConditionalProbability('P', pic_voxels_sum / total_sum_picture)
 			pic_cond_probabilities.append(cond_prob)
 			sen_voxels_sum = math.fsum(self._get_voxels_of_same_index(voxel_index=voxel_index, c='S'))
-			cond_prob = CondProbability('S', sen_voxels_sum / total_sum_sentence)
+			cond_prob = ConditionalProbability('S', sen_voxels_sum / total_sum_sentence)
 			sen_cond_probabilities.append(cond_prob)
 		self.conditional_probabilities.append(pic_cond_probabilities)
 		self.conditional_probabilities.append(sen_cond_probabilities)

### 2012-09-23 src/framework.py
naive bayes

@@ -11,6 +11,4 @@ naive_bayes.train()
 
 #subject2 = sio.loadmat('data-starplus-04820-v7.mat')
 
-#scan = subject['data'][10][0][16]
-
 naive_bayes.classify()

### 2012-09-23 src/model.py
naive bayes

@@ -137,6 +137,6 @@ class NaiveBayes:
 							math.fsum([voxel_value * math.log(self.conditional_probabilities[0][index][1]) for index, voxel_value in enumerate(valid_scan)])
 		p_sentence_scan = math.log(self.priori_probabilities['S']) + \
 							math.fsum([voxel_value * math.log(self.conditional_probabilities[1][index][1]) for index, voxel_value in enumerate(valid_scan)])
-		print "P(P|Scan): ", p_picture_scan
-		print "P(S|Scan): ", p_sentence_scan
+		print "P(Picture|Scan): ", p_picture_scan
+		print "P(Sentence|Scan): ", p_sentence_scan
 		print "Max: ", max(p_picture_scan, p_sentence_scan)

### 2012-09-23 src/model.py
naive bayes

@@ -134,9 +134,11 @@ class NaiveBayes:
 		for index, trial_index in enumerate(self.valid_voxel_indexes):
 			valid_scan.append(math.fabs(scan[trial_index]))
 		p_picture_scan = math.log(self.priori_probabilities['P']) + \
-							math.fsum([voxel_value * math.log(self.conditional_probabilities[0][index][1]) for index, voxel_value in enumerate(valid_scan)])
+							math.fsum([voxel_value * math.log(self.conditional_probabilities[0][index][1]) \
+								for index, voxel_value in enumerate(valid_scan)])
 		p_sentence_scan = math.log(self.priori_probabilities['S']) + \
-							math.fsum([voxel_value * math.log(self.conditional_probabilities[1][index][1]) for index, voxel_value in enumerate(valid_scan)])
+							math.fsum([voxel_value * math.log(self.conditional_probabilities[1][index][1]) \
+								for index, voxel_value in enumerate(valid_scan)])
 		print "P(Picture|Scan): ", p_picture_scan
 		print "P(Sentence|Scan): ", p_sentence_scan
 		print "Max: ", max(p_picture_scan, p_sentence_scan)

### 2012-09-23 src/model.py
naive bayes

@@ -15,8 +15,8 @@ from __future__ import division
 from collections import namedtuple
 import math
 
-Observation = namedtuple('Observation', 'klass voxels')
-ConditionalProbability = namedtuple('ConditionalProbability', 'klass value')
+Observation = namedtuple('Observation', 'klass voxels_vector')
+ConditionalProbability = namedtuple('ConditionalProbability', 'klass voxel_value')
 
 class NaiveBayes:
 
@@ -90,8 +90,8 @@ class NaiveBayes:
 	def _compute_priori_probabilities(self):
 		'''Computes the priori probabilities of the two classes P and S according to the formula:
 		P(Ci) = number of observations in class / number of all observations'''
-		total_picture_observations = len([obs for obs in self.features if obs[0] == 'P'])
-		total_sentence_observations = len([obs for obs in self.features if obs[0] == 'S'])
+		total_picture_observations = len([voxels_vector for klass, voxels_vector in self.features if klass == 'P'])
+		total_sentence_observations = len([voxels_vector for klass, voxels_vector in self.features if klass == 'S'])
 		total_observations = total_picture_observations + total_sentence_observations
 		self.priori_probabilities['P'] = total_picture_observations / total_observations
 		self.priori_probabilities['S'] = total_sentence_observations / total_observations

### 2012-09-23 src/model.py
naive bayes

@@ -98,7 +98,7 @@ class NaiveBayes:
 
 	def _compute_conditional_probabilities(self):
 		'''Computes the conditional probabilities of each voxel. Only for these voxels that
-		we define as valid. '''
+		we define as valid.'''
 		total_sum_picture = 0
 		total_sum_sentence = 0
 		for index, voxel_index in enumerate(self.valid_voxel_indexes):

### 2012-09-24 src/framework.py
naive bayes

@@ -1,14 +1,0 @@
-#!/usr/bin/env python
-
-import scipy.io as sio
-from model import NaiveBayes
-
-subject = sio.loadmat('data-starplus-04799-v7.mat')
-
-naive_bayes = NaiveBayes(subject)
-
-naive_bayes.train()
-
-#subject2 = sio.loadmat('data-starplus-04820-v7.mat')
-
-naive_bayes.classify()

### 2012-09-24 src/model.py

@@ -13,6 +13,7 @@ subject['data'][19][0][10][1566]
 
 from __future__ import division
 from collections import namedtuple
+import scipy.io as sio
 import math
 
 Observation = namedtuple('Observation', 'klass voxels_vector')
@@ -142,3 +143,9 @@ class NaiveBayes:
 		print "P(Picture|Scan): ", p_picture_scan
 		print "P(Sentence|Scan): ", p_sentence_scan
 		print "Max: ", max(p_picture_scan, p_sentence_scan)
+
+if __name__ == "__main__":
+	subject = sio.loadmat('data-starplus-04799-v7.mat')
+	naive_bayes = NaiveBayes(subject)
+	naive_bayes.train()
+	naive_bayes.classify()

### 2012-09-24 src/model.py
naive bayes

@@ -20,6 +20,7 @@ Observation = namedtuple('Observation', 'klass voxels_vector')
 ConditionalProbability = namedtuple('ConditionalProbability', 'klass voxel_value')
 
 class NaiveBayes:
+	'''This class implements a naive bayes classifier.'''
 
 	def __init__(self, subject):
 		self.subject = subject

### 2012-09-24 src/model.py
add data structure document, modified model

@@ -14,6 +14,7 @@ subject['data'][19][0][10][1566]
 from __future__ import division
 from collections import namedtuple
 import scipy.io as sio
+import scipy.stats as sis
 import math
 
 Observation = namedtuple('Observation', 'klass voxels_vector')
@@ -32,8 +33,10 @@ class NaiveBayes:
 		self.priori_probabilities = {}
 		self.conditional_probabilities = []
 		self.rois = ['CALC', 'LIPL', 'LT', 'LTRIA', 'LOPER', 'LIPS', 'LDLPFC']
-		self.first_stimulus_index = 0
-		self.second_stimulus_index = 16
+		#
+		self.first_stimulus_index = 20
+		#
+		self.second_stimulus_index = 40
 
 	def _get_valid_trials(self):
 		'''Returns the valid indexes of trials for a subject accoring to condition.
@@ -63,7 +66,7 @@ class NaiveBayes:
 		'''Returns the voxel absolute values of the same index. For example we want all the voxel
 		values with index 0 (means the first voxel value of each voxel vector) in order to
 		compute the conditional probability of this voxel: P(V1|Ci).'''
-		return [math.fabs(voxels[voxel_index]) for klass, voxels in self.features if klass == c]
+		return [voxels[voxel_index] for klass, voxels in self.features if klass == c]
 
 	def _init_values(self):
 		'''Extract some basic values that the classifier needs.'''
@@ -125,28 +128,68 @@ class NaiveBayes:
 		'''Trains the classifier.'''
 		self._init_values()
 		self._extract_values()
-		self._compute_priori_probabilities()
-		self._compute_conditional_probabilities()
+		#self._compute_priori_probabilities()
+		#self._compute_conditional_probabilities()
 
-	def classify(self):
+	def classify(self, subject):
 		'''Classifies a new scan of data. The following code is just an example as for classification
 		training data are used. This needs to be modified.'''
-		scan = self._get_voxel_vector(trial_index=4, stimulus=0) # It must show `Picture'
+		#####
+		#scan = self._get_voxel_vector(trial_index=25, stimulus=34)
+		#####
 		valid_scan = []
 		for index, trial_index in enumerate(self.valid_voxel_indexes):
-			valid_scan.append(math.fabs(scan[trial_index]))
-		p_picture_scan = math.log(self.priori_probabilities['P']) + \
[…]

### 2012-09-24 src/data_wrapper.py
added framework and data wrapper

@@ -1,0 +1,72 @@
+from collections import namedtuple
+
+Observation = namedtuple('Observation', 'klass voxels')
+
+class DataWrapper:
+	''''''
+
+	def __init__(self, subject):
+		''''''
+		self.subject = subject
+		self.num_of_voxels = None
+		self.num_of_trials = None
+		self.valid_trial_indexes = None
+		self.valid_voxel_indexes = None
+		self.rois = ['CALC', 'LIPL', 'LT', 'LTRIA', 'LOPER', 'LIPS', 'LDLPFC']
+		self.first_stimulus_index = 20
+		self.second_stimulus_index = 40
+		self.features = []
+
+	def _get_valid_trial_indexes(self):
+		'''Returns the valid indexes of trials for a subject accoring to condition.
+		We care only about conditions with value 2 or 3.'''
+		return [index for index in range(self.num_of_trials)
+					if self.subject['info'][0]['cond'][index] > 1]
+
+	def _get_valid_voxel_indexes(self):
+		'''Returns the valid indexes of voxels for a subject according to rois.'''
+		return [index for index in range(self.num_of_voxels)
+					if self.subject['meta']['colToROI'][0][0][index] in self.rois]
+
+	def _get_first_stimulus(self, trial_index):
+		'''Returns the first stimulus of a trial with index trial_index. This will
+		be either P if the subject saw a picture, or S if the subject saw a sentence.'''
+		return self.subject['info'][0][trial_index]['firstStimulus'][0]
+
+	def _get_voxels_of_same_scan(self, trial_index, scan_index):
+		'''Returns the voxel vector data for the trial with index trial_index
+		and scan index as defined by stimulus.'''
+		return self.subject['data'][trial_index][0][scan_index]
+
+	def _get_voxels_of_same_index(self, voxel_index, k):
+		'''Returns the voxel values of the same index. For example we want all the voxel
+		values with index 0 (means the first voxel value of each voxel vector) in order to
+		compute the conditional probability of this voxel: P(V1|Ci).'''
+		return [voxels[voxel_index] for klass, voxels in self.features if klass == k]
+
+	def _extract_features(self):
+		''''''
+		features = []
+		for index, trial_index in enumerate(self.valid_trial_indexes):
+			klass = self._get_first_stimulus(trial_index)
+			if klass == 'P':
+				features.append(Observation('P', self._get_voxels_of_same_scan(trial_index,
+																				self.first_stimulus_index)))
[…]

### 2012-09-24 src/framework.py

@@ -1,0 +1,4 @@
+!#/usr/bin/env python
+
+from data_wrapper import DataWrapper
+from models import NaiveBayes

### 2012-09-24 src/models.py

@@ -17,86 +17,87 @@ import scipy.io as sio
 import scipy.stats as sis
 import math
 
-Observation = namedtuple('Observation', 'klass voxels_vector')
+#Observation = namedtuple('Observation', 'klass voxels_vector')
 ConditionalProbability = namedtuple('ConditionalProbability', 'klass voxel_value')
 
 class NaiveBayes:
 	'''This class implements a naive bayes classifier.'''
 
-	def __init__(self, subject):
-		self.subject = subject
-		self.num_of_trials = None
-		self.num_of_voxels = None
-		self.valid_trial_indexes = None
-		self.valid_voxel_indexes = None
-		self.features = []
+	def __init__(self, data_wrapper):
+		#self.subject = subject
+		self.data_wrapper = data_wrapper
+		# self.num_of_trials = None
+		# self.num_of_voxels = None
+		# self.valid_trial_indexes = None
+		# self.valid_voxel_indexes = None
+		# self.features = []
 		self.priori_probabilities = {}
 		self.conditional_probabilities = []
-		self.rois = ['CALC', 'LIPL', 'LT', 'LTRIA', 'LOPER', 'LIPS', 'LDLPFC']
+		#self.rois = ['CALC', 'LIPL', 'LT', 'LTRIA', 'LOPER', 'LIPS', 'LDLPFC']
 		#
-		self.first_stimulus_index = 20
+		#self.first_stimulus_index = 20
 		#
-		self.second_stimulus_index = 40
+		#self.second_stimulus_index = 40
 
-	def _get_valid_trials(self):
-		'''Returns the valid indexes of trials for a subject accoring to condition.
-		We care only about conditions with value 2 or 3.'''
-		return [index for index in range(self.num_of_trials)
-					if self.subject['info'][0]['cond'][index] > 1]
+	# def _get_valid_trials(self):
+	# 	'''Returns the valid indexes of trials for a subject accoring to condition.
+	# 	We care only about conditions with value 2 or 3.'''
+	# 	return [index for index in range(self.num_of_trials)
+	# 				if self.subject['info'][0]['cond'][index] > 1]
 
-	def _get_valid_voxels(self):
-		'''Returns the valid indexes of voxels for a subject according to rois.'''
-		return [index for index in range(self.num_of_voxels)
-					if self.subject['meta']['colToROI'][0][0][index] in self.rois]
+	# def _get_valid_voxels(self):
[…]

### 2012-09-24 src/data_wrapper.py
refactor models and data wrapper

@@ -1,3 +1,16 @@
+'''
+This is the structure of the data corresponding to every subject.
+t = Trial
+s = Scan
+v = Voxel
+subject['data'][t][0][s][v]
+
+For example if we want to access the value of the 1567th voxel of the 20th trial
+and the 11th scan then we would have the following:
+
+subject['data'][19][0][10][1566]
+'''
+
 from collections import namedtuple
 
 Observation = namedtuple('Observation', 'klass voxels')
@@ -38,12 +51,6 @@ class DataWrapper:
 		and scan index as defined by stimulus.'''
 		return self.subject['data'][trial_index][0][scan_index]
 
-	def _get_voxels_of_same_index(self, voxel_index, k):
-		'''Returns the voxel values of the same index. For example we want all the voxel
-		values with index 0 (means the first voxel value of each voxel vector) in order to
-		compute the conditional probability of this voxel: P(V1|Ci).'''
-		return [voxels[voxel_index] for klass, voxels in self.features if klass == k]
-
 	def _extract_features(self):
 		''''''
 		features = []
@@ -55,9 +62,9 @@ class DataWrapper:
 				features.append(Observation('S', self._get_voxels_of_same_scan(trial_index,
 																				self.second_stimulus_index)))
 			else:
-				features.append(Observation('S', self._get_voxels_of_same_index(trial_index,
+				features.append(Observation('S', self._get_voxels_of_same_scan(trial_index,
 																				self.first_stimulus_index)))
-				features.append(Observation('P', self._get_voxels_of_same_index(trial_index,
+				features.append(Observation('P', self._get_voxels_of_same_scan(trial_index,
 																				self.second_stimulus_index)))
 		return features
 
@@ -69,4 +76,8 @@ class DataWrapper:
 		self.valid_voxel_indexes = self._get_valid_voxel_indexes()
 		self.features = self._extract_features()
 
-	
+	def get_voxels_of_same_index(self, voxel_index, k):
+		'''Returns the voxel values of the same index. For example we want all the voxel
+		values with index 0 (means the first voxel value of each voxel vector) in order to
+		compute the conditional probability of this voxel: P(V1|Ci).'''
+		return [voxels[voxel_index] for klass, voxels in self.features if klass == k]

### 2012-09-24 src/framework.py

@@ -1,4 +1,15 @@
-!#/usr/bin/env python
+#!/usr/bin/env python
 
 from data_wrapper import DataWrapper
 from models import NaiveBayes
+import scipy.io as sio
+
+subject = sio.loadmat('data-starplus-04799-v7.mat')
+
+data_wrapper = DataWrapper(subject)
+
+data_wrapper.extract_values()
+
+naive_bayes = NaiveBayes(data_wrapper)
+
+naive_bayes.train()

### 2012-09-24 src/models.py

@@ -1,97 +1,20 @@
-'''
-This is the structure of the data corresponding to every subject.
-t = Trial
-s = Scan
-v = Voxel
-subject['data'][t][0][s][v]
-
-For example if we want to access the value of the 1567th voxel of the 20th trial
-and the 11th scan then we would have the following:
-
-subject['data'][19][0][10][1566]
-'''
-
 from __future__ import division
 from collections import namedtuple
-import scipy.io as sio
 import scipy.stats as sis
 import math
 
-#Observation = namedtuple('Observation', 'klass voxels_vector')
 ConditionalProbability = namedtuple('ConditionalProbability', 'klass voxel_value')
 
 class NaiveBayes:
 	'''This class implements a naive bayes classifier.'''
 
 	def __init__(self, data_wrapper):
-		#self.subject = subject
 		self.data_wrapper = data_wrapper
-		# self.num_of_trials = None
-		# self.num_of_voxels = None
-		# self.valid_trial_indexes = None
-		# self.valid_voxel_indexes = None
-		# self.features = []
 		self.priori_probabilities = {}
-		self.conditional_probabilities = []
-		#self.rois = ['CALC', 'LIPL', 'LT', 'LTRIA', 'LOPER', 'LIPS', 'LDLPFC']
-		#
-		#self.first_stimulus_index = 20
-		#
-		#self.second_stimulus_index = 40
-
-	# def _get_valid_trials(self):
-	# 	'''Returns the valid indexes of trials for a subject accoring to condition.
-	# 	We care only about conditions with value 2 or 3.'''
-	# 	return [index for index in range(self.num_of_trials)
-	# 				if self.subject['info'][0]['cond'][index] > 1]
-
-	# def _get_valid_voxels(self):
-	# 	'''Returns the valid indexes of voxels for a subject according to rois.'''
-	# 	return [index for index in range(self.num_of_voxels)
-	# 				if self.subject['meta']['colToROI'][0][0][index] in self.rois]
-
-	# def _get_voxel_values(self, trial_indexes, stimulus, voxel):
-	# 	'''Returns the absolute value of the same index voxels.'''
-	# 	return [math.fabs(self.subject['data'][trial_index][0][stimulus][voxel]) for trial_index in trial_indexes]
-
-	# def _get_voxel_vector(self, trial_index, stimulus):
-	# 	'''Returns the voxel vector data for the trial with index trial_index
-	# 	and scan index as defined by stimulus.'''
[…]

### 2012-09-24 src/data_wrapper.py
minor changes

@@ -46,7 +46,8 @@ class DataWrapper:
 		be either P if the subject saw a picture, or S if the subject saw a sentence.'''
 		return self.subject['info'][0][trial_index]['firstStimulus'][0]
 
-	def _get_voxels_of_same_scan(self, trial_index, scan_index):
+	# non_private for now
+	def get_voxels_of_same_scan(self, trial_index, scan_index):
 		'''Returns the voxel vector data for the trial with index trial_index
 		and scan index as defined by stimulus.'''
 		return self.subject['data'][trial_index][0][scan_index]
@@ -57,14 +58,14 @@ class DataWrapper:
 		for index, trial_index in enumerate(self.valid_trial_indexes):
 			klass = self._get_first_stimulus(trial_index)
 			if klass == 'P':
-				features.append(Observation('P', self._get_voxels_of_same_scan(trial_index,
+				features.append(Observation('P', self.get_voxels_of_same_scan(trial_index,
 																				self.first_stimulus_index)))
-				features.append(Observation('S', self._get_voxels_of_same_scan(trial_index,
+				features.append(Observation('S', self.get_voxels_of_same_scan(trial_index,
 																				self.second_stimulus_index)))
 			else:
-				features.append(Observation('S', self._get_voxels_of_same_scan(trial_index,
+				features.append(Observation('S', self.get_voxels_of_same_scan(trial_index,
 																				self.first_stimulus_index)))
-				features.append(Observation('P', self._get_voxels_of_same_scan(trial_index,
+				features.append(Observation('P', self.get_voxels_of_same_scan(trial_index,
 																				self.second_stimulus_index)))
 		return features
 

### 2012-09-24 src/framework.py

@@ -13,3 +13,9 @@ data_wrapper.extract_values()
 naive_bayes = NaiveBayes(data_wrapper)
 
 naive_bayes.train()
+
+trial_index = 3
+scan_index = 49
+scan = subject['data'][trial_index][0][scan_index]
+
+naive_bayes.classify(scan)

### 2012-09-24 src/models.py

@@ -55,14 +55,10 @@ class NaiveBayes:
 		self._compute_means()
 		self._compute_standard_deviations()
 
-	def classify(self, subject):
-		'''Classifies a new scan of data. The following code is just an example as for classification
-		training data are used. This needs to be modified.'''
-		#####
-		#scan = self._get_voxel_vector(trial_index=25, stimulus=34)
-		#####
+	def classify(self, scan):
+		'''Classifies a new scan of data.'''
 		valid_scan = []
-		for index, trial_index in enumerate(self.valid_voxel_indexes):
+		for index, trial_index in enumerate(self.data_wrapper.valid_voxel_indexes):
 			valid_scan.append(scan[trial_index])
 		self.distributions = {}
 		distribution_picture = []
@@ -76,9 +72,6 @@ class NaiveBayes:
 													  self.standard_deviations_sentence[index]))
 		self.distributions['P'] = [math.log(value) for value in distribution_picture]
 		self.distributions['S'] = [math.log(value) for value in distribution_sentence]
-		#
-		#
-		#
 		sum_log_picture = math.fsum(self.distributions['P'])
 		sum_log_sentence = math.fsum(self.distributions['S'])
 		print "P(Picture|Scan) = ", sum_log_picture

### 2012-09-24 src/data_wrapper.py
added classification of all scans for one trial

@@ -43,13 +43,13 @@ class DataWrapper:
 
 	def _get_first_stimulus(self, trial_index):
 		'''Returns the first stimulus of a trial with index trial_index. This will
-		be either P if the subject saw a picture, or S if the subject saw a sentence.'''
+		be either `P' if the subject saw a picture, or `S' if the subject saw a sentence.'''
 		return self.subject['info'][0][trial_index]['firstStimulus'][0]
 
 	# non_private for now
 	def get_voxels_of_same_scan(self, trial_index, scan_index):
 		'''Returns the voxel vector data for the trial with index trial_index
-		and scan index as defined by stimulus.'''
+		and scan with index scan_index.'''
 		return self.subject['data'][trial_index][0][scan_index]
 
 	def _extract_features(self):

### 2012-09-24 src/framework.py

@@ -7,15 +7,13 @@ import scipy.io as sio
 subject = sio.loadmat('data-starplus-04799-v7.mat')
 
 data_wrapper = DataWrapper(subject)
-
 data_wrapper.extract_values()
 
 naive_bayes = NaiveBayes(data_wrapper)
-
 naive_bayes.train()
 
+# Classification starts here
 trial_index = 3
-scan_index = 49
-scan = subject['data'][trial_index][0][scan_index]
-
-naive_bayes.classify(scan)
+for scan_index in range(54):
+	scan = subject['data'][trial_index][0][scan_index]
+	naive_bayes.classify(scan, scan_index)

### 2012-09-24 src/models.py

@@ -55,7 +55,11 @@ class NaiveBayes:
 		self._compute_means()
 		self._compute_standard_deviations()
 
-	def classify(self, scan):
+	def _classification_result(self, sum_log_picture, sum_log_sentence):
+		''''''
+		return "Picture" if max(sum_log_picture, sum_log_sentence) == sum_log_picture else "Sentence"
+
+	def classify(self, scan, scan_index):
 		'''Classifies a new scan of data.'''
 		valid_scan = []
 		for index, trial_index in enumerate(self.data_wrapper.valid_voxel_indexes):
@@ -74,6 +78,7 @@ class NaiveBayes:
 		self.distributions['S'] = [math.log(value) for value in distribution_sentence]
 		sum_log_picture = math.fsum(self.distributions['P'])
 		sum_log_sentence = math.fsum(self.distributions['S'])
+		print "CLASSIFICATION OF SCAN: ", (scan_index + 1)
 		print "P(Picture|Scan) = ", sum_log_picture
 		print "P(Sentence|Scan) = ", sum_log_sentence
-		print "Class: ", "Picture" if max(sum_log_picture, sum_log_sentence) == sum_log_picture else "Sentence"
+		print "Class: ", self._classification_result(sum_log_picture, sum_log_sentence)

### 2012-09-25 src/data_wrapper.py
Removed pep8 errors.

@@ -15,70 +15,71 @@ from collections import namedtuple
 
 Observation = namedtuple('Observation', 'klass voxels')
 
+
 class DataWrapper:
-	''''''
+    ''''''
 
-	def __init__(self, subject):
-		''''''
-		self.subject = subject
-		self.num_of_voxels = None
-		self.num_of_trials = None
-		self.valid_trial_indexes = None
-		self.valid_voxel_indexes = None
-		self.rois = ['CALC', 'LIPL', 'LT', 'LTRIA', 'LOPER', 'LIPS', 'LDLPFC']
-		self.first_stimulus_index = 20
-		self.second_stimulus_index = 40
-		self.features = []
+    def __init__(self, subject):
+        ''''''
+        self.subject = subject
+        self.num_of_voxels = None
+        self.num_of_trials = None
+        self.valid_trial_indexes = None
+        self.valid_voxel_indexes = None
+        self.rois = ['CALC', 'LIPL', 'LT', 'LTRIA', 'LOPER', 'LIPS', 'LDLPFC']
+        self.first_stimulus_index = 20
+        self.second_stimulus_index = 40
+        self.features = []
 
-	def _get_valid_trial_indexes(self):
-		'''Returns the valid indexes of trials for a subject accoring to condition.
-		We care only about conditions with value 2 or 3.'''
-		return [index for index in range(self.num_of_trials)
-					if self.subject['info'][0]['cond'][index] > 1]
+    def _get_valid_trial_indexes(self):
+        '''Returns the valid indexes of trials for a subject accoring to condition.
+        We care only about conditions with value 2 or 3.'''
+        return [index for index in range(self.num_of_trials)
+                if self.subject['info'][0]['cond'][index] > 1]
 
-	def _get_valid_voxel_indexes(self):
-		'''Returns the valid indexes of voxels for a subject according to rois.'''
-		return [index for index in range(self.num_of_voxels)
-					if self.subject['meta']['colToROI'][0][0][index] in self.rois]
+    def _get_valid_voxel_indexes(self):
+        '''Returns the valid indexes of voxels for a subject according to rois.'''
+        return [index for index in range(self.num_of_voxels)
+                if self.subject['meta']['colToROI'][0][0][index] in self.rois]
 
-	def _get_first_stimulus(self, trial_index):
-		'''Returns the first stimulus of a trial with index trial_index. This will
-		be either `P' if the subject saw a picture, or `S' if the subject saw a sentence.'''
[…]

### 2012-09-25 src/framework.py

@@ -4,6 +4,7 @@ from data_wrapper import DataWrapper
 from models import NaiveBayes
 import scipy.io as sio
 
+# Loadmat should use $PATH, but somehow it fails for me (Mattijs)
 subject = sio.loadmat('data-starplus-04799-v7.mat')
 
 data_wrapper = DataWrapper(subject)
@@ -15,5 +16,5 @@ naive_bayes.train()
 # Classification starts here
 trial_index = 3
 for scan_index in range(54):
-	scan = subject['data'][trial_index][0][scan_index]
-	naive_bayes.classify(scan, scan_index)
+    scan = subject['data'][trial_index][0][scan_index]
+    naive_bayes.classify(scan, scan_index)

### 2012-09-25 src/ide.py

@@ -1,0 +1,8 @@
+
+
+def main(a):
+    if a:
+        print "hoi"
+
+if __name__ == '__main__':
+    main()

### 2012-09-25 src/models.py

@@ -5,80 +5,81 @@ import math
 
 ConditionalProbability = namedtuple('ConditionalProbability', 'klass voxel_value')
 
+
 class NaiveBayes:
-	'''This class implements a naive bayes classifier.'''
+    '''This class implements a naive bayes classifier.'''
 
-	def __init__(self, data_wrapper):
-		self.data_wrapper = data_wrapper
-		self.priori_probabilities = {}
-		self.means_picture = []
-		self.means_sentence = []
-		self.standard_deviations_picture = []
-		self.standard_deviations_sentence = []
+    def __init__(self, data_wrapper):
+        self.data_wrapper = data_wrapper
+        self.priori_probabilities = {}
+        self.means_picture = []
+        self.means_sentence = []
+        self.standard_deviations_picture = []
+        self.standard_deviations_sentence = []
 
-	def _compute_priori_probabilities(self):
-		'''Computes the priori probabilities of the two classes P and S according to the formula:
-		P(Ci) = number of observations in class / number of all observations'''
-		total_picture_observations = len([voxels_vector for klass, voxels_vector in self.data_wrapper.features if klass == 'P'])
-		total_sentence_observations = len([voxels_vector for klass, voxels_vector in self.data_wrapper.features if klass == 'S'])
-		total_observations = total_picture_observations + total_sentence_observations
-		self.priori_probabilities['P'] = total_picture_observations / total_observations
-		self.priori_probabilities['S'] = total_sentence_observations / total_observations
+    def _compute_priori_probabilities(self):
+        '''Computes the priori probabilities of the two classes P and S according to the formula:
+        P(Ci) = number of observations in class / number of all observations'''
+        total_picture_observations = len([voxels_vector for klass, voxels_vector in self.data_wrapper.features if klass == 'P'])
+        total_sentence_observations = len([voxels_vector for klass, voxels_vector in self.data_wrapper.features if klass == 'S'])
+        total_observations = total_picture_observations + total_sentence_observations
+        self.priori_probabilities['P'] = total_picture_observations / total_observations
+        self.priori_probabilities['S'] = total_sentence_observations / total_observations
 
-	def _compute_means(self):
-		''''''
-		for voxel_index in self.data_wrapper.valid_voxel_indexes:
[…]

### 2012-09-25 src/framework.py
Added some notes to article, added relative loading of data

@@ -5,6 +5,7 @@ from models import NaiveBayes
 import scipy.io as sio
 
 # Loadmat should use $PATH, but somehow it fails for me (Mattijs)
+# subject = sio.loadmat('../../fmri_project/data-starplus-04799-ROI.mat')
 subject = sio.loadmat('data-starplus-04799-v7.mat')
 
 data_wrapper = DataWrapper(subject)

### 2012-09-25 src/ide.py

@@ -1,8 +1,0 @@
-
-
-def main(a):
-    if a:
-        print "hoi"
-
-if __name__ == '__main__':
-    main()

### 2012-09-25 src/test.py

@@ -1,9 +1,0 @@
-from pylab import *
-
-n = 256
-X = np.linspace(-np.pi, np.pi, 256,endpoint=True)
-C,S = np.cos(X), np.sin(X)
-plot(X,C),plot(X,S)
-
-#savefig("../figures/exercice_1.png",dpi=72)
-show()

### 2012-09-25 src/framework.py
Added __main__ to framework to prevent auto-run after import

@@ -5,7 +5,7 @@ from models import NaiveBayes
 import scipy.io as sio
 
 # Loadmat should use $PATH, but somehow it fails for me (Mattijs)
-# subject = sio.loadmat('../../fmri_project/data-starplus-04799-ROI.mat')
+#subject = sio.loadmat('../../fmri_project/data-starplus-04799-v7.mat')
 subject = sio.loadmat('data-starplus-04799-v7.mat')
 
 data_wrapper = DataWrapper(subject)
@@ -14,8 +14,9 @@ data_wrapper.extract_values()
 naive_bayes = NaiveBayes(data_wrapper)
 naive_bayes.train()
 
-# Classification starts here
-trial_index = 3
-for scan_index in range(54):
-    scan = subject['data'][trial_index][0][scan_index]
-    naive_bayes.classify(scan, scan_index)
+if __name__ == '__main__':
+    # Classification starts here
+    trial_index = 3
+    for scan_index in range(54):
+        scan = subject['data'][trial_index][0][scan_index]
+        naive_bayes.classify(scan, scan_index)

### 2012-09-25 src/models.py
Added file for notes

@@ -79,7 +79,9 @@ class NaiveBayes:
         self.distributions['S'] = [math.log(value) for value in distribution_sentence]
         sum_log_picture = math.fsum(self.distributions['P'])
         sum_log_sentence = math.fsum(self.distributions['S'])
+        klass = self._classification_result(sum_log_picture, sum_log_sentence)
         print "CLASSIFICATION OF SCAN: ", (scan_index + 1)
         print "P(Picture|Scan) = ", sum_log_picture
         print "P(Sentence|Scan) = ", sum_log_sentence
-        print "Class: ", self._classification_result(sum_log_picture, sum_log_sentence)
+        print "Class: ", klass
+        return klass

### 2012-09-26 src/framework.py
Added a basic "result" to the output - still need to think about what we classify

@@ -4,8 +4,7 @@ from data_wrapper import DataWrapper
 from models import NaiveBayes
 import scipy.io as sio
 
-# Loadmat should use $PATH, but somehow it fails for me (Mattijs)
-#subject = sio.loadmat('../../fmri_project/data-starplus-04799-v7.mat')
+# Make sure the *.mat files are in the same directory
 subject = sio.loadmat('data-starplus-04799-v7.mat')
 
 data_wrapper = DataWrapper(subject)
@@ -14,9 +13,16 @@ data_wrapper.extract_values()
 naive_bayes = NaiveBayes(data_wrapper)
 naive_bayes.train()
 
+# Classification starts here
 if __name__ == '__main__':
-    # Classification starts here
     trial_index = 3
-    for scan_index in range(54):
+    score = 0
+    scanset = range(54)  # We probably want only certain scans classified
+    for scan_index in scanset:
         scan = subject['data'][trial_index][0][scan_index]
-        naive_bayes.classify(scan, scan_index)
+        # Obviously, expected should be derived
+        expected_klass = "Sentence"
+        predicted_klass = naive_bayes.classify(scan, scan_index)
+        if expected_klass == predicted_klass:
+            score += 1
+    print "Correctly classified %s out of %s scans" % (score, len(scanset))

### 2012-09-26 src/framework.py
added average matlab frameworks files

@@ -4,16 +4,25 @@ from data_wrapper import DataWrapper
 from models import NaiveBayes
 import scipy.io as sio
 
-subject = sio.loadmat('data-starplus-04799-v7.mat')
+files = [
+	'matlab_avergage_rois_04799.mat',
+	# 'matlab_avergage_rois_04820.mat',
+	# 'matlab_avergage_rois_04847.mat'
+]
 
-data_wrapper = DataWrapper(subject)
-data_wrapper.extract_values()
+for file in files:
+	subject = sio.loadmat(file)
 
-naive_bayes = NaiveBayes(data_wrapper)
-naive_bayes.train()
+	data_wrapper = DataWrapper(subject)
+	data_wrapper.extract_values()
 
+	naive_bayes = NaiveBayes(data_wrapper)
+	naive_bayes.train()
+
+
+class_subject = sio.loadmat('matlab_avergage_rois_04799.mat')
 # Classification starts here
 trial_index = 3
 for scan_index in range(54):
-	scan = subject['data'][trial_index][0][scan_index]
+	scan = class_subject['data'][trial_index][0][scan_index]
 	naive_bayes.classify(scan, scan_index)

### 2012-09-26 src/framework.py
minor change

@@ -5,7 +5,7 @@ from models import NaiveBayes
 import scipy.io as sio
 
 files = [
-	'matlab_avergage_rois_04799.mat',
+	'../matlab/matlab_avergage_rois_04799.mat',
 	# 'matlab_avergage_rois_04820.mat',
 	# 'matlab_avergage_rois_04847.mat'
 ]

### 2012-09-26 src/framework.py
Added function for getting expected class

@@ -1,37 +1,69 @@
 #!/usr/bin/env python
+'''
+Framework for training and classifying FMRI data.
 
+This framework offers a set of functions to train a classification algorithm
+and then use this to classify test data. For more info, please see the article
+accompanying this framework.
+
+@since September 2012
+'''
 from data_wrapper import DataWrapper
 from models import NaiveBayes
 import scipy.io as sio
 
 files = [
-	'../matlab/matlab_avergage_rois_04799.mat',
-	# 'matlab_avergage_rois_04820.mat',
-	# 'matlab_avergage_rois_04847.mat'
+    'matlab_avergage_rois_04799.mat',
+    # 'matlab_avergage_rois_04820.mat',
+    # 'matlab_avergage_rois_04847.mat'
 ]
 
 for file in files:
-	subject = sio.loadmat(file)
+    subject = sio.loadmat(file)
 
-	data_wrapper = DataWrapper(subject)
-	data_wrapper.extract_values()
+    data_wrapper = DataWrapper(subject)
+    data_wrapper.extract_values()
 
-	naive_bayes = NaiveBayes(data_wrapper)
-	naive_bayes.train()
+    naive_bayes = NaiveBayes(data_wrapper)
+    naive_bayes.train()
 
 
-class_subject = sio.loadmat('matlab_avergage_rois_04799.mat')
+def get_expected_class(subject, trial_index, scan_index):
+    '''
+    Checks the subject's info to see what class is expected when
+    classifying a scan from a certain trial. Trial index runs from 1 to 54.
+    Since we only know 'firstStimulus' for a trial, we'll map all scans before
+    [...] to firstStimulus and the rest to the secondStimulus.
+
+    According to data, trial_index 1..27 all have firstStimulus = 'P'.
+    Each trial has ~54 scans, so we'll asume that 27 is the halfwayPoint/
+
+    Returns either 'Picture' or 'Sentence'
+    '''
+    halfwayPoint = 27
+    first_stimulus = subject['info'][0][1][trial_index][0]
+    if first_stimulus == 'P':
+        if scan_index < halfwayPoint:
+            return 'Picture'
+        else:
+            return 'Sentence'
+    else:
+        if scan_index < halfwayPoint:
+            return 'Sentence'
+        else:
+            return 'Picture'
 
 # Classification starts here
 if __name__ == '__main__':
+    class_subject = sio.loadmat('matlab_avergage_rois_04799.mat')
     trial_index = 3
     score = 0
     scanset = range(54)  # We probably want only certain scans classified
[…]

### 2012-09-26 src/data_wrapper.py
generalize dta wrapper

@@ -15,68 +15,73 @@ from collections import namedtuple
 
 Observation = namedtuple('Observation', 'klass voxels')
 
+FIRST_STIMULUS_SCAN = 20
+SECOND_STIMULUS_SCAN = 40
 
 class DataWrapper:
     ''''''
 
-    def __init__(self, subject):
+    def __init__(self, subjects):
         ''''''
-        self.subject = subject
-        self.num_of_voxels = None
-        self.num_of_trials = None
-        self.valid_trial_indexes = None
-        self.valid_voxel_indexes = None
+        self.subjects = subjects
+        #self.num_of_voxels = None
+        #self.num_of_trials = None
+        #self.valid_trial_indexes = None
+        #self.valid_voxel_indexes = None
         self.rois = ['CALC', 'LIPL', 'LT', 'LTRIA', 'LOPER', 'LIPS', 'LDLPFC']
-        self.first_stimulus_index = 20
-        self.second_stimulus_index = 40
+        self.first_stimulus_index = FIRST_STIMULUS_SCAN
+        self.second_stimulus_index = SECOND_STIMULUS_SCAN
         self.features = []
 
-    def _get_valid_trial_indexes(self):
+    def _get_valid_trial_indexes(self, subject, num_of_trials):
         '''Returns the valid indexes of trials for a subject accoring to condition.
         We care only about conditions with value 2 or 3.'''
-        return [index for index in range(self.num_of_trials)
-                if self.subject['info'][0]['cond'][index] > 1]
+        return [index for index in range(num_of_trials)
+                if subject['info'][0]['cond'][index] > 1]
 
-    def _get_valid_voxel_indexes(self):
+    def _get_valid_voxel_indexes(self, num_of_voxels):
         '''Returns the valid indexes of voxels for a subject according to rois.'''
-        return [index for index in range(self.num_of_voxels)
+        return [index for index in range(num_of_voxels)
                 if self.subject['meta']['colToROI'][0][0][index] in self.rois]
 
-    def _get_first_stimulus(self, trial_index):
+    def _get_first_stimulus(self, subject, trial_index):
         '''Returns the first stimulus of a trial with index trial_index. This will
         be either `P' if the subject saw a picture, or `S' if the subject saw a sentence.'''
-        return self.subject['info'][0][trial_index]['firstStimulus'][0]
+        return subject['info'][0][trial_index]['firstStimulus'][0]
 
     # non_private for now
-    def get_voxels_of_same_scan(self, trial_index, scan_index):
[…]

### 2012-09-26 src/framework.py

@@ -4,33 +4,31 @@ from data_wrapper import DataWrapper
 from models import NaiveBayes
 import scipy.io as sio
 
-files = [
-	'../matlab/matlab_avergage_rois_04799.mat',
-	# 'matlab_avergage_rois_04820.mat',
-	# 'matlab_avergage_rois_04847.mat'
+FIRST_STIMULUS_SCAN = 20
+SECOND_STIMULUS_SCAN = 40
+
+subjects = [
+	sio.loadmat('../matlab/matlab_avergage_rois_04799.mat'),
+	sio.loadmat('../matlab/matlab_avergage_rois_04820.mat'),
+	sio.loadmat('../matlab/matlab_avergage_rois_04847.mat')
 ]
 
-for file in files:
-	subject = sio.loadmat(file)
+data_wrapper = DataWrapper(subjects)
+data_wrapper.extract_values()
 
-	data_wrapper = DataWrapper(subject)
-	data_wrapper.extract_values()
+naive_bayes = NaiveBayes(data_wrapper)
+naive_bayes.train()
 
-	naive_bayes = NaiveBayes(data_wrapper)
-	naive_bayes.train()
-
-
-class_subject = sio.loadmat('matlab_avergage_rois_04799.mat')
+class_subject = sio.loadmat('../matlab/matlab_avergage_rois_05675.mat')
 
 # Classification starts here
 if __name__ == '__main__':
-    trial_index = 3
     score = 0
     scanset = range(54)  # We probably want only certain scans classified
     for scan_index in scanset:
         scan = class_subject['data'][trial_index][0][scan_index]
         # Obviously, expected should be derived
-        expected_klass = "Sentence"
+        expected_klass = "Picture"
         predicted_klass = naive_bayes.classify(scan, scan_index)
         if expected_klass == predicted_klass:
             score += 1

### 2012-09-26 src/models.py

@@ -5,11 +5,13 @@ import math
 
 ConditionalProbability = namedtuple('ConditionalProbability', 'klass voxel_value')
 
+NUM_OF_VOXELS = 7
 
 class NaiveBayes:
     '''This class implements a naive bayes classifier.'''
 
     def __init__(self, data_wrapper):
+        ''''''
         self.data_wrapper = data_wrapper
         self.priori_probabilities = {}
         self.means_picture = []
@@ -28,7 +30,7 @@ class NaiveBayes:
 
     def _compute_means(self):
         ''''''
-        for voxel_index in self.data_wrapper.valid_voxel_indexes:
+        for voxel_index in range(NUM_OF_VOXELS):
             voxel_vector = self.data_wrapper.get_voxels_of_same_index(voxel_index, 'P')
             self.means_picture.append(math.fsum(voxel_vector) / len(voxel_vector))
             voxel_vector = self.data_wrapper.get_voxels_of_same_index(voxel_index, 'S')
@@ -36,14 +38,14 @@ class NaiveBayes:
 
     def _compute_standard_deviations(self):
         ''''''
-        for index, voxel_index in enumerate(self.data_wrapper.valid_voxel_indexes):
-            voxel_vector = self.data_wrapper.get_voxels_of_same_index(voxel_index, 'P')
+        for index in range(NUM_OF_VOXELS):
+            voxel_vector = self.data_wrapper.get_voxels_of_same_index(index, 'P')
             variance = 0
             for voxel_value in voxel_vector:
                 variance += ((voxel_value - self.means_picture[index]) ** 2) / (len(voxel_vector) - 1)
             self.standard_deviations_picture.append(math.sqrt(variance))
-        for index, voxel_index in enumerate(self.data_wrapper.valid_voxel_indexes):
-            voxel_vector = self.data_wrapper.get_voxels_of_same_index(voxel_index, 'S')
+        for index in range(NUM_OF_VOXELS):
+            voxel_vector = self.data_wrapper.get_voxels_of_same_index(index, 'S')
             variance = 0
             for voxel_value in voxel_vector:
                 variance += ((voxel_value - self.means_sentence[index]) ** 2) / (len(voxel_vector) - 1)

### 2012-09-26 src/data_wrapper.py
changes

@@ -33,7 +33,7 @@ class DataWrapper:
         self.second_stimulus_index = SECOND_STIMULUS_SCAN
         self.features = []
 
-    def _get_valid_trial_indexes(self, subject, num_of_trials):
+    def get_valid_trial_indexes(self, subject, num_of_trials):
         '''Returns the valid indexes of trials for a subject accoring to condition.
         We care only about conditions with value 2 or 3.'''
         return [index for index in range(num_of_trials)
@@ -79,7 +79,7 @@ class DataWrapper:
         for subject in self.subjects:
             num_of_trials = subject['meta']['ntrials'][0][0][0][0]
             #num_of_voxels = subject['meta']['nvoxels'][0][0][0][0]
-            valid_trial_indexes = self._get_valid_trial_indexes(subject, num_of_trials)
+            valid_trial_indexes = self.get_valid_trial_indexes(subject, num_of_trials)
             #valid_voxel_indexes = self._get_valid_voxel_indexes(num_of_voxels)
             self._extract_features(subject, valid_trial_indexes)
 

### 2012-09-26 src/framework.py

@@ -12,9 +12,6 @@ from data_wrapper import DataWrapper
 from models import NaiveBayes
 import scipy.io as sio
 
-FIRST_STIMULUS_SCAN = 20
-SECOND_STIMULUS_SCAN = 40
-
 subjects = [
 	sio.loadmat('../matlab/matlab_avergage_rois_04799.mat'),
 	sio.loadmat('../matlab/matlab_avergage_rois_04820.mat'),
@@ -52,16 +49,20 @@ def get_expected_class(subject, trial_index, scan_index):
         else:
             return 'Picture'
 
+FIRST_STIMULUS_SCAN = 20
+SECOND_STIMULUS_SCAN = 40
+
 # Classification starts here
 if __name__ == '__main__':
     class_subject = sio.loadmat('../matlab/matlab_avergage_rois_05675.mat')
+    num_of_trials = class_subject['meta']['ntrials'][0][0][0][0]
+    valid_trials = data_wrapper.get_valid_trial_indexes(class_subject, num_of_trials)
     score = 0
-    scanset = range(54)  # We probably want only certain scans classified
-    for scan_index in scanset:
-        scan = class_subject['data'][trial_index][0][scan_index]
-        # Obviously, expected should be derived
-        expected_class = get_expected_class(class_subject, trial_index, scan_index)
-        predicted_class = naive_bayes.classify(scan, scan_index)
-        if expected_class == predicted_class:
-            score += 1
+    for trial_index in valid_trials:
+    	for scan_index in [FIRST_STIMULUS_SCAN, SECOND_STIMULUS_SCAN]:
+    		scan = data_wrapper.get_voxels_of_same_scan(class_subject, trial_index, scan_index)
+        	expected_class = get_expected_class(class_subject, trial_index, scan_index)
+        	predicted_class = naive_bayes.classify(scan)
+        	if expected_class == predicted_class:
+        		score += 1
     print "Correctly classified %s out of %s scans" % (score, len(scanset))

### 2012-09-26 src/models.py

@@ -62,15 +62,15 @@ class NaiveBayes:
         ''''''
         return "Picture" if max(sum_log_picture, sum_log_sentence) == sum_log_picture else "Sentence"
 
-    def classify(self, scan, scan_index):
+    def classify(self, scan):
         '''Classifies a new scan of data.'''
-        valid_scan = []
-        for index, trial_index in enumerate(self.data_wrapper.valid_voxel_indexes):
-            valid_scan.append(scan[trial_index])
+        # valid_scan = []
+        # for index, trial_index in range(NUM_OF_VOXELS):
+        #     valid_scan.append(scan[trial_index])
         self.distributions = {}
         distribution_picture = []
         distribution_sentence = []
-        for index, voxel_value in enumerate(valid_scan):
+        for index, voxel_value in enumerate(scan):
             distribution_picture.append(sis.norm.pdf(voxel_value,
                                                      self.means_picture[index],
                                                      self.standard_deviations_picture[index]))
@@ -82,8 +82,9 @@ class NaiveBayes:
         sum_log_picture = math.fsum(self.distributions['P'])
         sum_log_sentence = math.fsum(self.distributions['S'])
         klass = self._classification_result(sum_log_picture, sum_log_sentence)
-        print "CLASSIFICATION OF SCAN: ", (scan_index + 1)
+        print "##########"
         print "P(Picture|Scan) = ", sum_log_picture
         print "P(Sentence|Scan) = ", sum_log_sentence
         print "Class: ", klass
+        print "##########"
         return klass

### 2012-09-26 src/data_wrapper.py
changes in model and framework

@@ -15,8 +15,8 @@ from collections import namedtuple
 
 Observation = namedtuple('Observation', 'klass voxels')
 
-FIRST_STIMULUS_SCAN = 20
-SECOND_STIMULUS_SCAN = 40
+FIRST_STIMULUS_SCAN = 17
+SECOND_STIMULUS_SCAN = 34
 
 class DataWrapper:
     ''''''
@@ -24,10 +24,6 @@ class DataWrapper:
     def __init__(self, subjects):
         ''''''
         self.subjects = subjects
-        #self.num_of_voxels = None
-        #self.num_of_trials = None
-        #self.valid_trial_indexes = None
-        #self.valid_voxel_indexes = None
         self.rois = ['CALC', 'LIPL', 'LT', 'LTRIA', 'LOPER', 'LIPS', 'LDLPFC']
         self.first_stimulus_index = FIRST_STIMULUS_SCAN
         self.second_stimulus_index = SECOND_STIMULUS_SCAN
@@ -78,9 +74,7 @@ class DataWrapper:
         ''''''
         for subject in self.subjects:
             num_of_trials = subject['meta']['ntrials'][0][0][0][0]
-            #num_of_voxels = subject['meta']['nvoxels'][0][0][0][0]
             valid_trial_indexes = self.get_valid_trial_indexes(subject, num_of_trials)
-            #valid_voxel_indexes = self._get_valid_voxel_indexes(num_of_voxels)
             self._extract_features(subject, valid_trial_indexes)
 
     def get_voxels_of_same_index(self, voxel_index, k):

### 2012-09-26 src/framework.py

@@ -12,6 +12,7 @@ from data_wrapper import DataWrapper
 from models import NaiveBayes
 import scipy.io as sio
 
+
 subjects = [
 	sio.loadmat('../matlab/matlab_avergage_rois_04799.mat'),
 	sio.loadmat('../matlab/matlab_avergage_rois_04820.mat'),
@@ -37,7 +38,7 @@ def get_expected_class(subject, trial_index, scan_index):
     Returns either 'Picture' or 'Sentence'
     '''
     halfwayPoint = 27
-    first_stimulus = subject['info'][0][1][trial_index][0]
+    first_stimulus = subject['info'][0][trial_index]['firstStimulus'][0]
     if first_stimulus == 'P':
         if scan_index < halfwayPoint:
             return 'Picture'
@@ -49,8 +50,8 @@ def get_expected_class(subject, trial_index, scan_index):
         else:
             return 'Picture'
 
-FIRST_STIMULUS_SCAN = 20
-SECOND_STIMULUS_SCAN = 40
+from data_wrapper import FIRST_STIMULUS_SCAN
+from data_wrapper import SECOND_STIMULUS_SCAN
 
 # Classification starts here
 if __name__ == '__main__':
@@ -58,6 +59,7 @@ if __name__ == '__main__':
     num_of_trials = class_subject['meta']['ntrials'][0][0][0][0]
     valid_trials = data_wrapper.get_valid_trial_indexes(class_subject, num_of_trials)
     score = 0
+    counter = 0
     for trial_index in valid_trials:
     	for scan_index in [FIRST_STIMULUS_SCAN, SECOND_STIMULUS_SCAN]:
     		scan = data_wrapper.get_voxels_of_same_scan(class_subject, trial_index, scan_index)
@@ -65,4 +67,5 @@ if __name__ == '__main__':
         	predicted_class = naive_bayes.classify(scan)
         	if expected_class == predicted_class:
         		score += 1
-    print "Correctly classified %s out of %s scans" % (score, len(scanset))
+        	counter += 1
+    print "Correctly classified %s out of %s scans" % (score, counter)

### 2012-09-26 src/framework.py
Added notes to logfile

@@ -13,9 +13,9 @@ from models import NaiveBayes
 import scipy.io as sio
 
 subjects = [
-	sio.loadmat('../matlab/matlab_avergage_rois_04799.mat'),
-	sio.loadmat('../matlab/matlab_avergage_rois_04820.mat'),
-	sio.loadmat('../matlab/matlab_avergage_rois_04847.mat')
+    sio.loadmat('../matlab/matlab_avergage_rois_04799.mat'),
+    sio.loadmat('../matlab/matlab_avergage_rois_04820.mat'),
+    sio.loadmat('../matlab/matlab_avergage_rois_04847.mat')
 ]
 
 data_wrapper = DataWrapper(subjects)
@@ -23,6 +23,7 @@ data_wrapper.extract_values()
 
 naive_bayes = NaiveBayes(data_wrapper)
 naive_bayes.train()
+
 
 def get_expected_class(subject, trial_index, scan_index):
     '''
@@ -59,10 +60,10 @@ if __name__ == '__main__':
     valid_trials = data_wrapper.get_valid_trial_indexes(class_subject, num_of_trials)
     score = 0
     for trial_index in valid_trials:
-    	for scan_index in [FIRST_STIMULUS_SCAN, SECOND_STIMULUS_SCAN]:
-    		scan = data_wrapper.get_voxels_of_same_scan(class_subject, trial_index, scan_index)
-        	expected_class = get_expected_class(class_subject, trial_index, scan_index)
-        	predicted_class = naive_bayes.classify(scan)
-        	if expected_class == predicted_class:
-        		score += 1
+        for scan_index in [FIRST_STIMULUS_SCAN, SECOND_STIMULUS_SCAN]:
+            scan = data_wrapper.get_voxels_of_same_scan(class_subject, trial_index, scan_index)
+            expected_class = get_expected_class(class_subject, trial_index, scan_index)
+            predicted_class = naive_bayes.classify(scan)
+            if expected_class == predicted_class:
+                score += 1
     print "Correctly classified %s out of %s scans" % (score, len(scanset))

### 2012-09-28 src/coord_data_wrapper.py
git status

@@ -1,0 +1,36 @@
+
+
+class CoordDataWrapper:
+	''''''
+
+	def __init__(self, subjects):
+		''''''
+		self.subjects = subjects
+		self.features = []
+
+	def extract_values(self):
+		''''''
+		features = []
+		subjects_coords = []
+		for subject_index in range(len(self.subjects)):
+			subject_coord = []
+			for z in range(8):
+				for x in range(64):
+					for y in range(64):
+						subject_coord.append(self.subjects[subject_index]['meta']['coordToCol'][0][0][x][y][z])
+			subjects_coords.append(subject_coord)
+
+		voxel_indexes = []
+		for coord_index in range(32768):
+			something = []
+			for subject_index in range(6):
+				something.append(subjects_coords[subject_index][coord_index])
+			voxel_indexes.append(something)
+
+		valid_coords = []
+		for coord in voxel_indexes:
+			if not 0 in coord:
+				valid_coords.append(coord)
+
+		for valid_coords_list in valid_coords:
+			print valid_coords_list

### 2012-09-28 src/test.py

@@ -1,0 +1,15 @@
+from coord_data_wrapper import CoordDataWrapper
+import scipy.io as sio
+
+subjects = [
+    sio.loadmat('../matlab/matlab_avergage_rois_04799.mat'),
+    sio.loadmat('../matlab/matlab_avergage_rois_04820.mat'),
+    sio.loadmat('../matlab/matlab_avergage_rois_04847.mat'),
+    sio.loadmat('../matlab/matlab_avergage_rois_05675.mat'),
+    sio.loadmat('../matlab/matlab_avergage_rois_05680.mat'),
+    sio.loadmat('../matlab/matlab_avergage_rois_05710.mat')
+]
+
+coord = CoordDataWrapper(subjects)
+
+coord.extract_values()

### 2012-10-16 article/article.tex
Added some notes to article, added relative loading of data
Added __main__ to framework to prevent auto-run after import
Article and bibtex update

@@
+%% This paper describes the results of the Machine Learning practical
+%% project. It should be approximately 8 pages long and contain the
+%% following sections:
+%%
+%% 1) Title, authors, abstract
+%% 2) Introduction
+%% 3) Application
+%% 4) Methods
+%% 5) Experimental Results
+%% 6) Discussion / Conclusion
+%% References
 \documentclass[preprint,journal]{vgtc}
@@
 \item
-  Mattijs Meiboom is with University of Groningen, E-mail: .
+  Mattijs Meiboom is with University of Groningen, E-mail: m.meiboom@student.rug.nl.
 \item
@@
 \item
-  Tijn Schouten is University of Groningen, E-mail: .
+  Tijn Schouten is University of Groningen, E-mail: t.m.schouten.2@student.rug.nl.
 \item
@@
 \abstract{
- This is the abstract section...
+Functional magnetic resonance imaging (fMRI) is a brain mapping technique that allows researchers to measure activity in a subject's brain.
+This technique is typically used to determine which areas of the brain are involved in performing a certain task.
+Using this knowledge, we may also be able to determine what task was performed by looking at fMRI data.
+In this paper, we will discuss the use of machine learning algorithms for the classification of fMRI data.
+By using classifiers we will try to predict what stimulus a subject is looking at, based on the brain activity during that time.
 } % end of abstract
@@
 %% please capitalize first letter and insert punctuation after last keyword
-\keywords{}
+\keywords{fMRI, Machine learning, Classification.}
 
@@
 %% ``\begin{document}'' command. It prepares and prints the title block.
-
 %% the only exception to this rule is the \firstsection command
 \firstsection{Introduction}
+\maketitle
+%% Now the actual contents of first section
+- basic terms introduced (fmri, classification, ...)
+- related work and results
+- state of the art: multivariate pattern analysis (\cite{Haxby2012852})
+- structure of rest of the paper
+\section{Application}
+- prediction
+- cognitive state
+\section{Methods}
+- supervised learning
+- naive bayes
+- multivariate
+- pca
+- comparison
+- classifying subtrials vs trials (ie 54 vs 27)
 
-\maketitle
+\subsection{Data acquisition and dimensions}
+The dataset used for our experiments was originally collected by Marcel Just et al.
+from Carnegie Mellon University's Center for Cognitive Brain Imaging and published as the StarPlus fMRI dataset.
+This set consists of fMRI data for 6 subjects that was obtained while having each subject participate in a number of trials.
+
+In these trials, subjects were presented with a stimulus on a computer screen.
+This stimulus was either a picture or a sentence.
+After a period of time, the stimulus would be replaced with a stimulus of the opposite type (i.e.
+picture with sentence and vice versa) after which the subject would indicate by pressing a button whether the sentence correctly described the picture.
+At the end of each trial, approximately 54 fMRI scans have be taken with an interval of 500ms.
+The entire dataset consist of voxel data and metadata providing information about the dataset.
+Included in this information is not only the trial and scan each voxel belongs to, but also the type of the trial (picture first or sentence first).
+This makes the dataset suitable for classification using supervised learning.
+
+In total, we have data available for approximately 4000 voxels per scan, 54 scans per trial, 40 trials per subject, 6 subjects.
+
+\section{Experimental Results}
+\section{Discussion / Conclusion}
 

### 2012-10-16 src/multivariate.py
changed data and removed d/2log etc

@@ -6,19 +6,37 @@ import math as m
 import scipy.stats as sis
 
 # load the matlab files
-file = [
-	'C:\Users\Robin\Documents\Machine Learning\data1-norm-roi'
-	,'C:\Users\Robin\Documents\Machine Learning\data2-norm-roi'
-	,'C:\Users\Robin\Documents\Machine Learning\data3-norm-roi'
-	,'C:\Users\Robin\Documents\Machine Learning\data4-norm-roi'
-	,'C:\Users\Robin\Documents\Machine Learning\data5-norm-roi'
-	,'C:\Users\Robin\Documents\Machine Learning\data6-norm-roi'
+# file = [
+# 	'C:\Users\Robin\Documents\Machine Learning\data1-norm-roi'
+# 	,'C:\Users\Robin\Documents\Machine Learning\data2-norm-roi'
+# 	,'C:\Users\Robin\Documents\Machine Learning\data3-norm-roi'
+# 	,'C:\Users\Robin\Documents\Machine Learning\data4-norm-roi'
+# 	,'C:\Users\Robin\Documents\Machine Learning\data5-norm-roi'
+# 	,'C:\Users\Robin\Documents\Machine Learning\data6-norm-roi'
+# ]
+
+# data = [
+# 		sio.loadmat('../matlab/data1-avg-roi-norm'),
+# 	    sio.loadmat('../matlab/data2-avg-roi-norm'),
+# 	    sio.loadmat('../matlab/data3-avg-roi-norm'),
+# 	    sio.loadmat('../matlab/data4-avg-roi-norm'),
+# 	    sio.loadmat('../matlab/data5-avg-roi-norm'),
+# 	    sio.loadmat('../matlab/data6-avg-roi-norm'),
+# ]
+data = [
+        sio.loadmat('../matlab/data1-avg-roi-norm'),
+        sio.loadmat('../matlab/data2-avg-roi-norm'),
+        sio.loadmat('../matlab/data3-avg-roi-norm'),
+        sio.loadmat('../matlab/data4-avg-roi-norm'),
+        sio.loadmat('../matlab/data5-avg-roi-norm'),
+        sio.loadmat('../matlab/data6-avg-roi-norm'),
 ]
+	
 
 #place in scipy format
-data = range(0,6)
-for n in range(0,6):
-	data[n] = sio.loadmat(file[n])
[…]

### 2012-10-16 src/multivariate.py
Separation of PS/SP and scan

@@ -27,26 +27,30 @@ correcttotal3 = 0
 correcttotal4 = 0
 
 #repeat the process for 6 subjects
-for iteration in range(0,6):
-	# arrr = data array for n-1 subjects
-	arrr = range(0,20)
-	# excl = data for the excluded (test) subject
-	excl = range(0,20)
-	for n in range(0,20):
-		arrr[n] = [],[]
-		excl[n] = [],[]
-	#determine the scans
-	scans = range(10,20) + range(27,37)
+for iteration in range(0, 6):
+    # arrr = data array for n-1 subjects
+    arrr = range(0, 20)
+    # excl = data for the excluded (test) subject
+    excl = range(0, 20)
+    for n in range(0, 20):
+		arrr[n] = [], []
+		excl[n] = [], []
 
-	#import data into arrr and excl
-	exclude = iteration
-	for subject in range(len(data)):
-		counter=-1
-		for scan in scans:
-			counter = counter + 1
+	#determine the scans
+    firstRange = range(10, 20)
+    secondRange = range(27, 37)
+    scans = firstRange + secondRange
+
+    #import data into arrr and excl
+    exclude = iteration
+    for subject in range(len(data)):
+        counter=-1
+        for scan in scans:
+            counter = counter + 1
 			for trial in range(0,53):
 				if data[subject]['info'][0,trial]['cond'][0]>1:
-					if data[subject]['info'][0,trial]['firstStimulus'][0]=='P':	
+					if (data[subject]['info'][0,trial]['firstStimulus'][0]=='P' and scan in firstRange) or
+                       (data[subject]['info'][0,trial]['firstStimulus'][0]=='S' and scan in secondRange):
 						if not subject == exclude:
 							arrr[counter][0].append(data[subject]['data'][trial][0][scan])
 						else:
@@ -56,13 +60,13 @@ for iteration in range(0,6):
 							arrr[counter][1].append(data[subject]['data'][trial][0][scan])
 						else:
 							excl[counter][1].append(data[subject]['data'][trial][0][scan])
-		
+
 	#set %correct per subject to 0
 	correct_per_subject1 = 0
[…]

### 2012-10-16 src/multivariate.py
F*ing tabs

@@ -62,22 +62,22 @@ for iteration in range(0, 6):
     #import data into arrr and excl
     exclude = iteration
     for subject in range(len(data)):
-        counter=-1
+        counter = -1
         for scan in scans:
             counter = counter + 1
-			for trial in range(0,53):
-				if data[subject]['info'][0,trial]['cond'][0]>1:
-					if (data[subject]['info'][0,trial]['firstStimulus'][0]=='P' and scan in firstRange) or
+            for trial in range(0, 53):
+                if data[subject]['info'][0,trial]['cond'][0]>1:
+                    if (data[subject]['info'][0,trial]['firstStimulus'][0]=='P' and scan in firstRange) or
                        (data[subject]['info'][0,trial]['firstStimulus'][0]=='S' and scan in secondRange):
-						if not subject == exclude:
-							arrr[counter][0].append(data[subject]['data'][trial][0][scan])
-						else:
-							excl[counter][0].append(data[subject]['data'][trial][0][scan])
-					else:
-						if not subject == exclude:
-							arrr[counter][1].append(data[subject]['data'][trial][0][scan])
-						else:
-							excl[counter][1].append(data[subject]['data'][trial][0][scan])
+                        if not subject == exclude:
+                            arrr[counter][0].append(data[subject]['data'][trial][0][scan])
+                        else:
+                            excl[counter][0].append(data[subject]['data'][trial][0][scan])
+                    else:
+                        if not subject == exclude:
+                            arrr[counter][1].append(data[subject]['data'][trial][0][scan])
+                        else:
+                            excl[counter][1].append(data[subject]['data'][trial][0][scan])
 
 	#set %correct per subject to 0
 	correct_per_subject1 = 0

### 2012-10-16 src/multivariate.py
F*ing tabs

@@ -7,21 +7,21 @@ import scipy.stats as sis
 
 # load the matlab files
 # file = [
-# 	'C:\Users\Robin\Documents\Machine Learning\data1-norm-roi'
-# 	,'C:\Users\Robin\Documents\Machine Learning\data2-norm-roi'
-# 	,'C:\Users\Robin\Documents\Machine Learning\data3-norm-roi'
-# 	,'C:\Users\Robin\Documents\Machine Learning\data4-norm-roi'
-# 	,'C:\Users\Robin\Documents\Machine Learning\data5-norm-roi'
-# 	,'C:\Users\Robin\Documents\Machine Learning\data6-norm-roi'
+#   'C:\Users\Robin\Documents\Machine Learning\data1-norm-roi'
+#   ,'C:\Users\Robin\Documents\Machine Learning\data2-norm-roi'
+#   ,'C:\Users\Robin\Documents\Machine Learning\data3-norm-roi'
+#   ,'C:\Users\Robin\Documents\Machine Learning\data4-norm-roi'
+#   ,'C:\Users\Robin\Documents\Machine Learning\data5-norm-roi'
+#   ,'C:\Users\Robin\Documents\Machine Learning\data6-norm-roi'
 # ]
 
 # data = [
-# 		sio.loadmat('../matlab/data1-avg-roi-norm'),
-# 	    sio.loadmat('../matlab/data2-avg-roi-norm'),
-# 	    sio.loadmat('../matlab/data3-avg-roi-norm'),
-# 	    sio.loadmat('../matlab/data4-avg-roi-norm'),
-# 	    sio.loadmat('../matlab/data5-avg-roi-norm'),
-# 	    sio.loadmat('../matlab/data6-avg-roi-norm'),
+#       sio.loadmat('../matlab/data1-avg-roi-norm'),
+#       sio.loadmat('../matlab/data2-avg-roi-norm'),
+#       sio.loadmat('../matlab/data3-avg-roi-norm'),
+#       sio.loadmat('../matlab/data4-avg-roi-norm'),
+#       sio.loadmat('../matlab/data5-avg-roi-norm'),
+#       sio.loadmat('../matlab/data6-avg-roi-norm'),
 # ]
 data = [
         sio.loadmat('../matlab/data1-avg-roi-norm'),
@@ -36,7 +36,7 @@ data = [
 #place in scipy format
 #data = range(0,6)
 #for n in range(0,6):
[…]

### 2012-10-16 src/multivariate.py
F*ing tabs

@@ -66,8 +66,8 @@ for iteration in range(0, 6):
         for scan in scans:
             counter = counter + 1
             for trial in range(0, 53):
-                if data[subject]['info'][0,trial]['cond'][0]>1:
-                    if (data[subject]['info'][0,trial]['firstStimulus'][0]=='P' and scan in firstRange) or
+                if data[subject]['info'][0,trial]['cond'][0] > 1:
+                    if (data[subject]['info'][0,trial]['firstStimulus'][0]=='P' and scan in firstRange) or \
                        (data[subject]['info'][0,trial]['firstStimulus'][0]=='S' and scan in secondRange):
                         if not subject == exclude:
                             arrr[counter][0].append(data[subject]['data'][trial][0][scan])

### 2012-10-16 src/multivariate.py
SP/PS madness

@@ -66,9 +66,8 @@ for iteration in range(0, 6):
         for scan in scans:
             counter = counter + 1
             for trial in range(0, 53):
-                if data[subject]['info'][0,trial]['cond'][0] > 1:
-                    if (data[subject]['info'][0,trial]['firstStimulus'][0]=='P' and scan in firstRange) or \
-                       (data[subject]['info'][0,trial]['firstStimulus'][0]=='S' and scan in secondRange):
+                if data[subject]['info'][0, trial]['cond'][0] > 1:
+                    if data[subject]['info'][0, trial]['firstStimulus'][0] == 'P':
                         if not subject == exclude:
                             arrr[counter][0].append(data[subject]['data'][trial][0][scan])
                         else:
@@ -88,9 +87,9 @@ for iteration in range(0, 6):
     # test phase
     for ps in range(0, 2):
         for trial in range(0, 20):
-            logprob_perscan_pic = [0]*20
-            logprob_perscan_sent = [0]*20
-            for scan in range(0,20):
+            logprob_perscan_pic = [0] * 20
+            logprob_perscan_sent = [0] * 20
+            for scan in range(0, 20):
                 # calculate covariance matrix
 
                     if scan < 10:

### 2012-10-16 src/multivariate.py
Fixed ep8

@@ -94,39 +94,39 @@ for iteration in range(0, 6):
 
                     if scan < 10:
                         #for picture. mat_cov = covariance matrix, mat_det = determinant, mat_inv = inverse matrix, mat_x vector of test input, mat_xminmu = (x-mu)
-                        mat_mean=np.subtract(np.matrix(arrr[scan][0]), np.mean(np.matrix(arrr[scan][0]), axis=0))
-                        mat_cov=np.dot(np.transpose(mat_mean),mat_mean)/len((arrr[scan][0]))
-                        mat_det= np.linalg.det(mat_cov)
-                        mat_inv= mat_cov.I
-                        mat_x= np.matrix(excl[scan][ps])[trial,0:25]
-                        mat_xminmu= np.matrix(excl[scan][ps])[trial,0:25]-np.mean(np.matrix(arrr[scan][0]), axis=0)
+                        mat_mean = np.subtract(np.matrix(arrr[scan][0]), np.mean(np.matrix(arrr[scan][0]), axis=0))
+                        mat_cov = np.dot(np.transpose(mat_mean), mat_mean) / len((arrr[scan][0]))
+                        mat_det = np.linalg.det(mat_cov)
+                        mat_inv = mat_cov.I
+                        mat_x = np.matrix(excl[scan][ps])[trial, 0:25]
+                        mat_xminmu = np.matrix(excl[scan][ps])[trial, 0:25] - np.mean(np.matrix(arrr[scan][0]), axis=0)
                         #multivariate equation
-                        logprob_perscan_pic[scan] = m.log(2*m.pi)-0.5*m.log(mat_det)-0.5*mat_xminmu*mat_inv*np.transpose(mat_xminmu)
+                        logprob_perscan_pic[scan] = m.log(2 * m.pi) - 0.5 * m.log(mat_det) - 0.5 * mat_xminmu * mat_inv * np.transpose(mat_xminmu)
                         #for sentence
-                        mat_mean=np.subtract(np.matrix(arrr[scan][1]), np.mean(np.matrix(arrr[scan][1]), axis=0))
-                        mat_cov=np.dot(np.transpose(mat_mean),mat_mean)/len((arrr[scan][1]))
-                        mat_det= np.linalg.det(mat_cov)
-                        mat_inv= mat_cov.I
[…]

### 2012-10-16 src/multivariate.py
updated article & refs

@@ -3,7 +3,6 @@
 import scipy.io as sio
 import numpy as np
 import math as m
-import scipy.stats as sis
 
 # load the matlab files
 # file = [

### 2012-10-16 src/multivariate.py
data file changed

@@ -24,12 +24,12 @@ import scipy.stats as sis
 #       sio.loadmat('../matlab/data6-avg-roi-norm'),
 # ]
 data = [
-        sio.loadmat('../matlab/data1-avg-roi-norm'),
-        sio.loadmat('../matlab/data2-avg-roi-norm'),
-        sio.loadmat('../matlab/data3-avg-roi-norm'),
-        sio.loadmat('../matlab/data4-avg-roi-norm'),
-        sio.loadmat('../matlab/data5-avg-roi-norm'),
-        sio.loadmat('../matlab/data6-avg-roi-norm'),
+        sio.loadmat('../matlab/data1-selectroi-norm'),
+        sio.loadmat('../matlab/data2-selectroi-norm'),
+        sio.loadmat('../matlab/data3-selectroi-norm'),
+        sio.loadmat('../matlab/data4-selectroi-norm'),
+        sio.loadmat('../matlab/data5-selectroi-norm'),
+        sio.loadmat('../matlab/data6-selectroi-norm'),
 ]
 
 
@@ -56,7 +56,7 @@ for iteration in range(0, 6):
 
     #determine the scans
     firstRange = range(10, 20)
-    secondRange = range(27, 37)
+    secondRange = range(22, 32)
     scans = firstRange + secondRange
 
     #import data into arrr and excl

### 2012-10-16 article/article.tex

@@
 In this paper, we will discuss the use of machine learning algorithms for the classification of fMRI data.
-By using classifiers we will try to predict what stimulus a subject is looking at, based on the brain activity during that time.
+By using classifiers we will try to predict the cognitive state of a subject, based on its brain activity during that time.
 } % end of abstract
@@
 \firstsection{Introduction}
+\label{sec:introduction}
 \maketitle
 %% Now the actual contents of first section
+Traditionally, it has been a challenge to relate between a simple cognitive state (such as `looking at a triangle`) and the corresponding activity of the brain \cite{Pone0001394}.
+This correspondence may even vary from subject to subject, making the challenge even bigger.
+Since the introduction of fMRI scans, researchers have been able to acquire detailed measurements of brain activity, allowing for machine learning methods to be applied for the detection of patterns in fMRI data.
+
 - basic terms introduced (fmri, classification, ...)
 - related work and results
+- briefly build up to state of the art, mention methods we tried?
 - state of the art: multivariate pattern analysis (\cite{Haxby2012852})
-- structure of rest of the paper
+
+In the rest of this paper, we will discuss a number of machine learning techniques and their ability to classify our test data.
+In section~\ref{sec:methods} we will discuss the setup of our experiments and the analysis performed on our data.
+After that, in sections~\ref{sec:results} and ~\ref{sec:discussion}, we will present the results from these experiments and discuss them.
+We conclude this paper with a discussion on possible improvements and suggest some additional experiments that could be done.
+
 \section{Application}
+\label{sec:application}
+
+What can we do with it now?
+
+- classification problem
 - prediction
@@
 \section{Methods}
+\label{sec:methods}
+- comparison
+- classifying subtrials vs trials (ie 54 vs 27)
+- analysis / tools (python etc)
+
+- same for each algorithm:
     - supervised learning
+        - feature selection
+        - classification / discrimination / scoring function
     - naive bayes
@@
     - pca
-- comparison
-- classifying subtrials vs trials (ie 54 vs 27)
 
@@
 
+Here we will need a nice picture, perhaps a pagewide one with labels of data for 1 trial?
+
 \section{Experimental Results}
+\label{sec:results}
+- raw results per algorithm
+
 \section{Discussion / Conclusion}
+\label{sec:discussion}
 

### 2012-10-16 src/demo/linear_svc.py
added demo files

@@ -1,0 +1,114 @@
+#!/usr/bin/env python
+
+from __future__ import division
+import scipy.io as sio
+import mdp
+import numpy as np
+from array import array
+from sklearn.svm import LinearSVC
+
+ROIS = ['CALC', 'LIPL', 'LT', 'LTRIA', 'LOPER', 'LIPS', 'LDLPFC']
+FIRST_STIMULUS_SCAN_INDEXES = range(10, 20)
+SECOND_STIMULUS_SCAN_INDEXES = range(22, 32)
+
+def get_number_of_trials(subject):
+    ''''''
+    return subject['meta']['ntrials']
+
+def get_number_of_voxels(subject):
+    ''''''
+    return subject['meta']['nvoxels']
+
+def get_valid_trial_indexes(subject, ntrials):
+    '''Returns the valid indexes of trials for a subject accoring to condition.
+    We care only about conditions with value 2 or 3.'''
+    return [index for index in range(ntrials)
+            if subject['info'][0]['cond'][index] > 1]
+
+def get_valid_voxel_indexes(subject, nvoxels):
+    '''Returns the valid indexes of voxels for a subject according to rois.'''
+    return [index for index in range(nvoxels)
+            if subject['meta']['colToROI'][0][0][index] in ROIS]
+
+def get_voxels_of_same_scan(subject, trial_index, scan_index):
+        '''Returns the voxel vector data for the trial with index trial_index
+        and scan with index scan_index.'''
+        return subject['data'][trial_index][0][scan_index]
+
+def get_first_stimulus_class(subject, trial_index):
+    '''Returns the first stimulus class of a trial with index trial_index. This will
+    be either `P' if the subject saw a picture first and then a sentence, or `S' if
+    the subject saw a sentence first and then a picture.'''
+    return subject['info'][0][trial_index]['firstStimulus'][0]
+
+def extract_features(subject, valid_trial_indexes, valid_voxel_indexes):
+    ''''''
+    features_p = []
+    features_s = []
+    for trial_index in valid_trial_indexes:
+        klass = get_first_stimulus_class(subject, trial_index)
+        if klass == 'P':
+            for scan_index in FIRST_STIMULUS_SCAN_INDEXES:
+                voxels = get_voxels_of_same_scan(subject, trial_index, scan_index)
+                features_p.append(voxels)
[…]

### 2012-10-16 src/demo/linear_svc.py
gaussian and linearsvc for demo

@@ -100,15 +100,15 @@ def main(subjects, c_subject):
         prediction = nb.predict(X[i])
         if prediction == Y[i]:
             correct += 1
-    print (correct / len(X)) * 100, "%"
+    return ((correct / len(X)) * 100)
 
 if __name__ == "__main__":
-    subjects = [
-        sio.loadmat('../../matlab/data1-norm-roi.mat'),
-        sio.loadmat('../../matlab/data2-norm-roi.mat'),
-        sio.loadmat('../../matlab/data3-norm-roi.mat'),
-        sio.loadmat('../../matlab/data4-norm-roi.mat'),
-        sio.loadmat('../../matlab/data5-norm-roi.mat'),
-    ]
-    c_subject = sio.loadmat('../../matlab/data6-norm-roi.mat')
-    main(subjects, c_subject)
+    total_result = 0
+    for excl_index in range(6):
+        subjects = [(sio.loadmat('../../matlab/demo/data%d-select-norm-avgroi.mat' % (index + 1))) for index in range(6) if index is not excl_index]
+        print "CLassification excluding subject %d" % (excl_index + 1)
+        c_subject = sio.loadmat('../../matlab/demo/data%d-select-norm-avgroi.mat' % (excl_index + 1))
+        result = main(subjects, c_subject)
+        total_result += result
+        print "Result: %s %%" % result
+    print "Total result: %s %%" % (total_result / 6)

### 2012-10-16 src/demo/multivariate.py
multivariate for demo

@@ -24,7 +24,7 @@ for iteration in range(0, 6):
 
     #determine the scans
     firstRange = range(10, 20)
-    secondRange = range(27, 37)
+    secondRange = range(22, 32)
     scans = firstRange + secondRange
 
     #import data into arrr and excl

### 2012-10-16 article/article.tex
Article update (on feature selection)

@@
 \item
-  Tijn Schouten is University of Groningen, E-mail: t.m.schouten.2@student.rug.nl.
+  Tijn Schouten is with University of Groningen, E-mail: t.m.schouten.2@student.rug.nl.
 \item
@@
 
-- same for each algorithm:
-    - supervised learning
-        - feature selection
-        - classification / discrimination / scoring function
-    - naive bayes
-    - multivariate
-    - pca
-
 \subsection{Data acquisition and dimensions}
@@
 In total, we have data available for approximately 4000 voxels per scan, 54 scans per trial, 40 trials per subject, 6 subjects.
+Voxels have been grouped into 25 regions of interest (see \textit{Feature selection}) that may be of use for analysis.
 
 Here we will need a nice picture, perhaps a pagewide one with labels of data for 1 trial?
+
+\subsection{Feature selection}
+%Inspired by: http://www.brainvoyager.com/bvqx/doc/UsersGuide/WebHelp/Content/MVPATools/MVPA_Basic_Concepts.htm
+Typical input for the machine learning algorithms discussed in this paper consists of a number of feature vectors, one for each fMRI scan.
+Such a feature vector $\vec{x}$ contains measures for the $N$ features we use to represent that scan and is derived from our input data.
+In the most basic case, we could use the entire set of voxels belonging to a scan, resulting in a feature vector containing potentially (hundreds of) thousands of features.
+In the case that we have a very large dimension of features combined with only a limited amount of scans, our classifier may have difficulty learning the right classification function.
+This problem is formally known as the \textit{Hughes effect}\cite{Hughes1054102} which states that the predictive power of a machine learning algorithm decreases as the dimensionality increases with a fixed number of training examples.
+
+So how do we cope with this problem?
+On common solution is to reduce $|\vec{x}|$ by only using voxels belonging to a specific region of interest (ROI).
+However, this requires us to known which regions are actually of interest \textit{a priori}.
+- ROIs, means, stddev, skew, kurtosis?
+
+- same for each algorithm:
+    - supervised learning
+        - feature selection
+            - ROI vs Whole Brain
+        - classification / discrimination / scoring function
+    - naive bayes
+    - multivariate
+    - multi-voxel pattern analysis
+    - pca
 

### 2012-10-16 src/demo/gaussian_nb.py
svc, nb for demo

@@ -1,106 +1,153 @@
-#!/usr/bin/env python
-
 from __future__ import division
 import scipy.io as sio
-import mdp
-import numpy as np
-from array import array
+from scipy import std as sstd
+from scipy import mean as smean
+from functions import *
 from sklearn.naive_bayes import GaussianNB
+from math import fsum
 
+# Data
+print "Loading subjects..."
+subjects = [sio.loadmat('../../matlab/demo/data%d-select-norm-avgroi.mat' % (index + 1)) for index in range(6)]
+print "Loaded subjects!"
+
+# Configuration
+FIRST_STIMULUS_SCAN_INDICES = range(10, 20)
+SECOND_STIMULUS_SCAN_INDICES = range(22, 32)
+COMPONENTS = 200
+
+# Rois are ignored if filter coords is false
 ROIS = ['CALC', 'LIPL', 'LT', 'LTRIA', 'LOPER', 'LIPS', 'LDLPFC']
-FIRST_STIMULUS_SCAN_INDEXES = range(10, 20)
-SECOND_STIMULUS_SCAN_INDEXES = range(22, 32)
 
-def get_number_of_trials(subject):
-    ''''''
-    return subject['meta']['ntrials']
+FLAG_PCA = False
+FLAG_NORM = False
+FLAG_FILTER_COORDS = False
 
-def get_number_of_voxels(subject):
-    ''''''
-    return subject['meta']['nvoxels']
+# If true, average the prediction per set of scans
+FLAG_PER_SCAN = False
 
-def get_valid_trial_indexes(subject, ntrials):
-    '''Returns the valid indexes of trials for a subject accoring to condition.
-    We care only about conditions with value 2 or 3.'''
-    return [index for index in range(ntrials)
-            if subject['info'][0]['cond'][index] > 1]
+'''
+PCA ALGORITHM
+1) Apply filtering on voxels (for each subject)
+2) Construct data matrix for PCA:
+- each row is a scan from the range of applicable scans (split for type of trial - PS, SP, PS+SP?)
+- each column is a value for a certain coordinate
+3) PCA
+4) Split matrix based on subject
+5) Use one part for training
+6) Use the other part for testing
+'''
+# 1: Get list of coordinates
+validCoords = []
+trials = getValidTrialIndices(subjects[0])
+if FLAG_FILTER_COORDS:
+    print "Constructing nonempty coordinates across subjects"
+    coords = getCoordinatesForSubject(subjects[0], ROIS)
[…]

### 2012-10-16 src/demo/linear_svc.py

@@ -1,114 +1,150 @@
-#!/usr/bin/env python
-
 from __future__ import division
 import scipy.io as sio
-import mdp
-import numpy as np
-from array import array
+from scipy import std as sstd
+from scipy import mean as smean
+from functions import *
 from sklearn.svm import LinearSVC
+from math import fsum
 
+# Data
+print "Loading subjects..."
+subjects = [sio.loadmat('../../matlab/demo/data%d-select-norm-avgroi.mat' % (index + 1)) for index in range(6)]
+print "Loaded subjects!"
+
+# Configuration
+FIRST_STIMULUS_SCAN_INDICES = range(10, 20)
+SECOND_STIMULUS_SCAN_INDICES = range(22, 32)
+COMPONENTS = 200
+
+# Rois are ignored if filter coords is false
 ROIS = ['CALC', 'LIPL', 'LT', 'LTRIA', 'LOPER', 'LIPS', 'LDLPFC']
-FIRST_STIMULUS_SCAN_INDEXES = range(10, 20)
-SECOND_STIMULUS_SCAN_INDEXES = range(22, 32)
 
-def get_number_of_trials(subject):
-    ''''''
-    return subject['meta']['ntrials']
+FLAG_PCA = False
+FLAG_NORM = False
+FLAG_FILTER_COORDS = False
 
-def get_number_of_voxels(subject):
-    ''''''
-    return subject['meta']['nvoxels']
+# If true, average the prediction per set of scans
+FLAG_PER_SCAN = False
 
-def get_valid_trial_indexes(subject, ntrials):
-    '''Returns the valid indexes of trials for a subject accoring to condition.
-    We care only about conditions with value 2 or 3.'''
-    return [index for index in range(ntrials)
-            if subject['info'][0]['cond'][index] > 1]
+'''
+PCA ALGORITHM
+1) Apply filtering on voxels (for each subject)
+2) Construct data matrix for PCA:
+- each row is a scan from the range of applicable scans (split for type of trial - PS, SP, PS+SP?)
+- each column is a value for a certain coordinate
+3) PCA
+4) Split matrix based on subject
+5) Use one part for training
+6) Use the other part for testing
+'''
+# 1: Get list of coordinates
+validCoords = []
+trials = getValidTrialIndices(subjects[0])
+if FLAG_FILTER_COORDS:
+    print "Constructing nonempty coordinates across subjects"
+    coords = getCoordinatesForSubject(subjects[0], ROIS)
[…]

### 2012-10-16 article/article.tex

@@
 - briefly build up to state of the art, mention methods we tried?
-- state of the art: multivariate pattern analysis (\cite{Haxby2012852})
+- state of the art: probably multivariate pattern analysis! (\cite{Haxby2012852})
 
@@
 \label{sec:methods}
+General approach consisting of three steps, plus basic data acquisition and preprocessing.
+\begin{enumerate}
+  \item Feature selection
+  \item Training
+  \item Testing
+\end{enumerate}
+
 - comparison
@@
 \subsection{Feature selection}
-%Inspired by: http://www.brainvoyager.com/bvqx/doc/UsersGuide/WebHelp/Content/MVPATools/MVPA_Basic_Concepts.htm
+%Useful background info: http://www.brainvoyager.com/bvqx/doc/UsersGuide/WebHelp/Content/MVPATools/MVPA_Basic_Concepts.htm
 Typical input for the machine learning algorithms discussed in this paper consists of a number of feature vectors, one for each fMRI scan.
@@
 - same for each algorithm:
-    - supervised learning
-        - feature selection
-            - ROI vs Whole Brain
-        - classification / discrimination / scoring function
-    - naive bayes
+    - What features did we select?
+    - How did we train?
+    - How did we test?
+        - How did we calculate performance?
+- algorithms:
+    - gaussian naive-bayes (gnb)
+        - pca + gnb
     - multivariate
-    - multi-voxel pattern analysis
-    - pca
+    - svm
 

### 2012-10-16 article/article.tex
article update

--- 2012-10-16 article/article.tex
@@
 In this paper, we will discuss the use of machine learning algorithms for the classification of fMRI data.
-By using classifiers we will try to predict the cognitive state of a subject, based on its brain activity during that time.
+By using classifiers we will try to predict the cognitive state of subjects, based on their brain activity during that time.
 } % end of abstract
@@
 Traditionally, it has been a challenge to relate between a simple cognitive state (such as `looking at a triangle`) and the corresponding activity of the brain \cite{Pone0001394}.
-This correspondence may even vary from subject to subject, making the challenge even bigger.
-Since the introduction of fMRI scans, researchers have been able to acquire detailed measurements of brain activity, allowing for machine learning methods to be applied for the detection of patterns in fMRI data.
+This correspondence may even vary from subject to subject, making the challenge even harder.
+Since the introduction of fMRI, researchers have been able to acquire detailed measurements of brain activity, allowing for machine learning methods to be applied for the detection of patterns in fMRI data.
 
@@
 
-In total, we have data available for approximately 4000 voxels per scan, 54 scans per trial, 40 trials per subject, 6 subjects.
+In total, we have data available for approximately 4800 voxels per scan, 54 scans per trial, 40 trials per subject, 6 subjects.
 

### 2012-10-16 article/article.tex

--- 2012-10-16 article/article.tex
@@
 
-In total, we have data available for approximately 4000 voxels per scan, 54 scans per trial, 40 trials per subject, 6 subjects.
+In total, we have data available for approximately 4800 voxels per scan, 54 scans per trial, 40 trials per subject, 6 subjects.
 Voxels have been grouped into 25 regions of interest (see \textit{Feature selection}) that may be of use for analysis.

### 2012-10-22 src/functions.py
Modified PCA and added component iteration

@@ -73,7 +73,7 @@ def getVoxelsForScan(subject, trial_index, scan_index, voxel_indices):
         if voxel_indices:
             voxels = []
             for voxel_index in voxel_indices:
-                voxels.append(subject['data'][trial_index][0][scan_index][voxel_index])
+                voxels.append(subject['data'][trial_index][0][scan_index][voxel_index - 1])
             return voxels
 
         return subject['data'][trial_index][0][scan_index]
@@ -104,5 +104,5 @@ def applyPCA(features, components):
     pcanode.train(features)
     pcanode.stop_training()
 
-    print "Explained variance: %s" % pcanode.explained_variance
+    #print "Explained variance: %s" % pcanode.explained_variance
     return pcanode(features)

### 2012-10-22 src/Naive_Bayes_per_scan.py
Naive

@@ -1,0 +1,116 @@
+#!/usr/bin/env python
+# C:\Users\Robin\Documents\Machine Learning\stuff\src
+import scipy.io as sio
+import numpy as np
+import math as m
+import scipy.stats as sis
+
+file = [
+	'C:\Users\Robin\Documents\Machine Learning\data1-norm-roi'
+	,'C:\Users\Robin\Documents\Machine Learning\data2-norm-roi'
+	,'C:\Users\Robin\Documents\Machine Learning\data3-norm-roi'
+	,'C:\Users\Robin\Documents\Machine Learning\data4-norm-roi'
+	,'C:\Users\Robin\Documents\Machine Learning\data5-norm-roi'
+	,'C:\Users\Robin\Documents\Machine Learning\data6-norm-roi'
+]
+
+data = range(0,6)
+for n in range(0,6):
+	data[n] = sio.loadmat(file[n])
+
+# make variables
+correcttotal1 = 0
+correcttotal2 = 0
+correcttotal3 = 0
+correcttotal4 = 0
+for iteration in range(0,6):
+	arrr = range(0,20)
+	excl = range(0,20)
+	means = range(0,20)	
+	sds = range(0,20)
+	for n in range(0,20):
+		arrr[n] = [],[]
+		excl[n] = [],[]
+		means[n] = [],[]
+		sds[n] = [],[]
+	means = arrr
+
+
+	scans = range(10,20) + range(22,32)
+
+	#import data
+	exclude = iteration
+	for subject in range(len(data)):
+		counter=-1
+		for scan in scans:
+			counter = counter + 1
+			for trial in range(0,53):
+				if data[subject]['info'][0,trial]['cond'][0]>1:
+					if data[subject]['info'][0,trial]['firstStimulus'][0]=='P':	
+						if not subject == exclude:
+							arrr[counter][0].append(data[subject]['data'][trial][0][scan])
+						else:
+							excl[counter][0].append(data[subject]['data'][trial][0][scan])
+					else:
+						if not subject == exclude:
+							arrr[counter][1].append(data[subject]['data'][trial][0][scan])
+						else:
+							excl[counter][1].append(data[subject]['data'][trial][0][scan])
+				# arrr[counter][0]=np.matrix(arrr[counter][0])
[…]

### 2012-10-22 src/demo/settings.py
changes

@@ -1,4 +1,6 @@
 FILES = {
+	'Orig': 'data%d-original.mat',
+	'AvgROI': 'data%d_average_rois.mat',
     'NormAvgROI': './matlab/data%d-select-norm-avgroi.mat',
     'NormAvgSDROI': './matlab/data%d-select-norm-avgsdroi.mat',
     'NormPerVoxAvgROI': './matlab/data%d-select-normpervox-avgroi.mat',

### 2012-10-22 src/demo/multivariate.py
minor changes

@@ -5,7 +5,7 @@ import math as m
 from settings import FILES
 
 # load the matlab files
-data = [sio.loadmat(FILES['NormAvgROI'] % (index + 1)) for index in range(6)]
+data = [sio.loadmat(FILES['NormPerVoxAvgROI'] % (index + 1)) for index in range(6)]
 
 # make variables
 correcttotal1 = 0

### 2012-10-22 src/demo/settings.py

@@ -1,6 +1,6 @@
 FILES = {
-	'Orig': 'data%d-original.mat',
-	'AvgROI': 'data%d_average_rois.mat',
+    'Orig': './matlab/data%d-original.mat',
+    'AvgROI': './matlab/data%d_average_rois.mat',
     'NormAvgROI': './matlab/data%d-select-norm-avgroi.mat',
     'NormAvgSDROI': './matlab/data%d-select-norm-avgsdroi.mat',
     'NormPerVoxAvgROI': './matlab/data%d-select-normpervox-avgroi.mat',

### 2012-10-22 src/demo/Naive_Bayes_per_scan.py
results table

@@ -1,0 +1,100 @@
+#!/usr/bin/env python
+import scipy.io as sio
+import numpy as np
+import math as m
+import scipy.stats as sis
+from settings import FILES
+
+# load the matlab files
+data = [sio.loadmat(FILES['NormAvgSDROI'] % (index + 1)) for index in range(6)]
+
+# make variables
+correcttotal1 = 0
+correcttotal2 = 0
+correcttotal3 = 0
+correcttotal4 = 0
+for iteration in range(0,6):
+    arrr = range(0,20)
+    excl = range(0,20)
+    means = range(0,20) 
+    sds = range(0,20)
+    for n in range(0,20):
+        arrr[n] = [],[]
+        excl[n] = [],[]
+        means[n] = [],[]
+        sds[n] = [],[]
+    means = arrr
+
+
+    scans = range(10,20) + range(22,32)
+
+    #import data
+    exclude = iteration
+    for subject in range(len(data)):
+        counter=-1
+        for scan in scans:
+            counter = counter + 1
+            for trial in range(0,53):
+                if data[subject]['info'][0,trial]['cond'][0]>1:
+                    if data[subject]['info'][0,trial]['firstStimulus'][0]=='P': 
+                        if not subject == exclude:
+                            arrr[counter][0].append(data[subject]['data'][trial][0][scan])
+                        else:
+                            excl[counter][0].append(data[subject]['data'][trial][0][scan])
+                    else:
+                        if not subject == exclude:
+                            arrr[counter][1].append(data[subject]['data'][trial][0][scan])
+                        else:
+                            excl[counter][1].append(data[subject]['data'][trial][0][scan])
+                # arrr[counter][0]=np.matrix(arrr[counter][0])
+                # arrr[counter][1]=np.matrix(arrr[counter][1])
+        
+    #probabilities
+    correct_per_subject1 = 0
+    correct_per_subject2 = 0
+    correct_per_subject3 = 0
+    correct_per_subject4 = 0
+    for ps in range(0,2):
+        for trial in range(0,20):
+            logprob_perscan_pic = [0]*20
+            logprob_perscan_sent = [0]*20
[…]

### 2012-10-22 src/demo/linear_svc.py

@@ -9,7 +9,7 @@ from settings import FILES
 
 # Data
 print "Loading subjects..."
-subjects = [sio.loadmat(FILES['NormAvgROI'] % (index + 1)) for index in range(6)]
+subjects = [sio.loadmat(FILES['NormPerVoxAvgROI'] % (index + 1)) for index in range(6)]
 print "Loaded subjects!"
 
 # Configuration

### 2012-10-22 src/demo/multivariate.py

@@ -5,7 +5,7 @@ import math as m
 from settings import FILES
 
 # load the matlab files
-data = [sio.loadmat(FILES['NormPerVoxAvgROI'] % (index + 1)) for index in range(6)]
+data = [sio.loadmat(FILES['NormAvgROI'] % (index + 1)) for index in range(6)]
 
 # make variables
 correcttotal1 = 0

### 2012-10-22 src/pca_all_subs_by_index.py
Added matlab commands for movie

@@ -32,15 +32,16 @@ print "Loaded subjects!"
 # Configuration
 FIRST_STIMULUS_SCAN_INDICES = range(10, 20)
 SECOND_STIMULUS_SCAN_INDICES = range(27, 37)
-COMPONENTS = 500
+COMPONENTS = 250
 
 # Rois are ignored if filter coords is false
-#ROIS = ['CALC', 'LIPL', 'LT', 'LTRIA', 'LOPER', 'LIPS', 'LDLPFC']
-ROIS = []
+ROIS = ['CALC', 'LIPL', 'LT', 'LTRIA', 'LOPER', 'LIPS', 'LDLPFC']
+#ROIS = []
 
 FLAG_PCA = True
 FLAG_NORM = False
-FLAG_FILTER_COORDS = True
+FLAG_FILTER_COORDS = False
+FLAG_FILTER_INDICES = True
 
 # If true, average the prediction per set of scans
 FLAG_PER_SCAN = False
@@ -63,7 +64,7 @@ PCA ALGORITHM
 5) Use one part for training
 6) Use the other part for testing
 '''
-# 1: Get list of coordinates
+# 1: Get list of all indices
 validCoords = []
 trials = getValidTrialIndices(subjects[0])
 if FLAG_FILTER_COORDS:
@@ -78,6 +79,22 @@ if FLAG_FILTER_COORDS:
         print "No valid coordinates present"
         quit()
 
+if FLAG_FILTER_INDICES:
+    '''
+    Take all voxels that are present for each subject
+    '''
+    lengths = []
+    for subject in subjects:
+        lengths.append(len(subject['meta']['colToROI'][0][0]))
+    voxel_indices = range(1, min(lengths))
+
+    # Filter ROIS
+    if len(ROIS) > 0:
+        roi_indices = []
+        for index in voxel_indices:
+            if subjects[0]['meta']['colToROI'][0][0][index - 1] in ROIS:
+                roi_indices.append(index)
+        voxel_indices = roi_indices
 
 # 2: Construct Matrix based on those coordinates
 scans = []
@@ -87,7 +104,6 @@ for subject_index, subject in enumerate(subjects):
     if not FLAG_FILTER_COORDS:
         validCoords = getCoordinatesForSubject(subject, [])
 
-    voxel_indices = getVoxelsForCoordinates(subject, validCoords)
     print "Appending scans for subject %s" % subject_index
     for trial_index in trials:
[…]

### 2012-10-22 src/demo/pca_by_index.py
Moved pca by index to demo folder

@@ -20,12 +20,13 @@ SECOND_STIMULUS_SCAN_INDICES = range(27, 37)
 COMPONENTS = 148
 
 # Rois are ignored if filter coords is false
-#ROIS = ['CALC', 'LIPL', 'LT', 'LTRIA', 'LOPER', 'LIPS', 'LDLPFC']
-ROIS = []
+ROIS = ['CALC', 'LIPL', 'LT', 'LTRIA', 'LOPER', 'LIPS', 'LDLPFC']
+#ROIS = []
 
 FLAG_PCA = True
 FLAG_NORM = False
-FLAG_FILTER_COORDS = True
+FLAG_FILTER_COORDS = False
+FLAG_FILTER_INDICES = True
 
 # If true, average the prediction per set of scans
 FLAG_PER_SCAN = False
@@ -48,7 +49,7 @@ PCA ALGORITHM
 5) Use one part for training
 6) Use the other part for testing
 '''
-# 1: Get list of coordinates
+# 1: Get list of all indices
 validCoords = []
 trials = getValidTrialIndices(subjects[0])
 if FLAG_FILTER_COORDS:
@@ -63,7 +64,24 @@ if FLAG_FILTER_COORDS:
         print "No valid coordinates present"
         quit()
 
+if FLAG_FILTER_INDICES:
+    '''
+    Take all voxels that are present for each subject
+    '''
+    lengths = []
+    for subject in subjects:
+        lengths.append(len(subject['meta']['colToROI'][0][0]))
+    voxel_indices = range(1, min(lengths))
+    print lengths
+    # Filter ROIS
+    if len(ROIS) > 0:
+        roi_indices = []
+        for index in voxel_indices:
+            if subjects[0]['meta']['colToROI'][0][0][index - 1] in ROIS:
+                roi_indices.append(index)
+        voxel_indices = roi_indices
 
+print len(voxel_indices)
 # 2: Construct Matrix based on those coordinates
 scans = []
 labels = []
@@ -72,7 +90,6 @@ for subject_index, subject in enumerate(subjects):
     if not FLAG_FILTER_COORDS:
         validCoords = getCoordinatesForSubject(subject, [])
 
-    voxel_indices = getVoxelsForCoordinates(subject, validCoords)
     print "Appending scans for subject %s" % subject_index
     for trial_index in trials:
         for scan_index in FIRST_STIMULUS_SCAN_INDICES:

### 2012-10-22 src/pca_all_subs.py

@@ -24,8 +24,8 @@ print "Loaded subjects!"
 
 # Configuration
 FIRST_STIMULUS_SCAN_INDICES = range(10, 20)
-SECOND_STIMULUS_SCAN_INDICES = range(27, 37)
-COMPONENTS = 500
+SECOND_STIMULUS_SCAN_INDICES = range(22, 32)
+COMPONENTS = 148
 
 # Rois are ignored if filter coords is false
 #ROIS = ['CALC', 'LIPL', 'LT', 'LTRIA', 'LOPER', 'LIPS', 'LDLPFC']
@@ -106,7 +106,7 @@ original_scans = scans
 vec_x = []
 vec_y = []
 
-for n in range(COMPONENTS - 50, COMPONENTS + 50):
+for n in range(COMPONENTS, COMPONENTS):
     scans = applyPCA(original_scans, n + 1)
     #print "Reduced input data to %s components using PCA" % (len(scans[0]))
     #print "Explained variance is %s" % scans_reduced.explained_variance

### 2012-10-22 src/pca_all_subs_by_index.py

@@ -12,12 +12,12 @@ from math import fsum
 print "Loading subjects..."
 
 subjects = [
-    sio.loadmat('../matlab/data-starplus-04799-v7.mat'),
-    sio.loadmat('../matlab/data-starplus-04820-v7.mat'),
-    sio.loadmat('../matlab/data-starplus-04847-v7.mat'),
-    sio.loadmat('../matlab/data-starplus-05675-v7.mat'),
-    sio.loadmat('../matlab/data-starplus-05680-v7.mat'),
-    sio.loadmat('../matlab/data-starplus-05710-v7.mat')
+    sio.loadmat('demo/matlab/data1-original.mat'),
+    sio.loadmat('demo/matlab/data2-original.mat'),
+    sio.loadmat('demo/matlab/data3-original.mat'),
+    sio.loadmat('demo/matlab/data4-original.mat'),
+    sio.loadmat('demo/matlab/data5-original.mat'),
+    sio.loadmat('demo/matlab/data6-original.mat')
 ]
 # subjects = [
 #     sio.loadmat('../matlab/matlab_avergage_rois_04799.mat'),
@@ -32,11 +32,11 @@ print "Loaded subjects!"
 # Configuration
 FIRST_STIMULUS_SCAN_INDICES = range(10, 20)
 SECOND_STIMULUS_SCAN_INDICES = range(27, 37)
-COMPONENTS = 250
+COMPONENTS = 148
 
 # Rois are ignored if filter coords is false
-ROIS = ['CALC', 'LIPL', 'LT', 'LTRIA', 'LOPER', 'LIPS', 'LDLPFC']
-#ROIS = []
+#ROIS = ['CALC', 'LIPL', 'LT', 'LTRIA', 'LOPER', 'LIPS', 'LDLPFC']
+ROIS = []
 
 FLAG_PCA = True
 FLAG_NORM = False
@@ -87,7 +87,7 @@ if FLAG_FILTER_INDICES:
     for subject in subjects:
         lengths.append(len(subject['meta']['colToROI'][0][0]))
     voxel_indices = range(1, min(lengths))
-
+    print lengths
     # Filter ROIS
     if len(ROIS) > 0:
         roi_indices = []
@@ -96,6 +96,7 @@ if FLAG_FILTER_INDICES:
                 roi_indices.append(index)
         voxel_indices = roi_indices
 
+print len(voxel_indices)
 # 2: Construct Matrix based on those coordinates
 scans = []
 labels = []
[…]

### 2012-10-22 src/demo/pca_by_index.py
Moved images, added some comments to pca_by_index

@@ -1,3 +1,23 @@
+'''
+Classification of trials using PCA (principal component analysis) as a
+dimension reduction technique.
+
+Input:      Original voxel data, with a dimensionality of ~4900 voxels per scans_per_subject
+Training:   PCA reduction, followed by GaussianNB
+Scoring:    Per group of scans (belonging to the same subtrial), the prediction for each scan
+            is combined for an overall judgement.
+
+Notes:      ROI filter can be applied if you do not like the computational intensity of PCA
+            on the whole set of voxels. This will however reduce the generalization performance
+            of the classifier.
+
+Algorithm:  - get voxel indices
+            - construct voxel matrices (containing both training data and test data)
+            - apply PCA to voxel matrix
+            - split matrix into training and testing data
+            - train classifier
+            - test classifier
+'''
 from __future__ import division
 import scipy.io as sio
 import pylab
@@ -6,50 +26,27 @@ from scipy import mean as smean
 from functions import *
 from settings import FILES
 from sklearn.naive_bayes import GaussianNB
-#from sklearn.preprocessing import normalize
-from math import fsum
 
-# Data
-print "Loading subjects..."
+# Load the dataset
 subjects = [sio.loadmat(FILES['Orig'] % (index + 1)) for index in range(6)]
-print "Loaded subjects!"
 
-# Configuration
+# Configuration of the classifier. Please indicate the scangroups and the range
+# of PCA components to try.
+# If desired, the list ROIS can contain strings to filter the voxels on, or
+# leave empty for all ROIS.
 FIRST_STIMULUS_SCAN_INDICES = range(10, 20)
 SECOND_STIMULUS_SCAN_INDICES = range(27, 37)
 COMPONENTS = 148
-
-# Rois are ignored if filter coords is false
 ROIS = ['CALC', 'LIPL', 'LT', 'LTRIA', 'LOPER', 'LIPS', 'LDLPFC']
-#ROIS = []
 
+# Some flags for testing, probably no need to change this
 FLAG_PCA = True
 FLAG_NORM = False
 FLAG_FILTER_COORDS = False
 FLAG_FILTER_INDICES = True
-
-# If true, average the prediction per set of scans
 FLAG_PER_SCAN = False
 
-# Just some debugging output
-# for coord in validCoords:
-#     print "Regions for coord (%s,%s,%s)" % (coord[0], coord[1], coord[2])
[…]

### 2012-10-23 src/demo/multivariate.py
printing

@@ -5,7 +5,7 @@ import math as m
 from settings import FILES
 
 # load the matlab files
-data = [sio.loadmat(FILES['NormAvgROI'] % (index + 1)) for index in range(6)]
+data = [sio.loadmat(FILES['AvgROI'] % (index + 1)) for index in range(6)]
 
 # make variables
 correcttotal1 = 0
@@ -106,6 +106,7 @@ for iteration in range(0, 6):
                 correct_per_subject3 = correct_per_subject3 + 1
             if (np.sum(logprob_perscan_pic[0:10]) < np.sum(logprob_perscan_sent[0:10])and ps == 1):
                 correct_per_subject4 = correct_per_subject4 + 1
+
     # add to total
     correcttotal1 = correcttotal1 + correct_per_subject1
     correcttotal2 = correcttotal2 + correct_per_subject2
@@ -113,5 +114,8 @@ for iteration in range(0, 6):
     correcttotal4 = correcttotal4 + correct_per_subject4
 
 # print accuracy
-print 'picture % (1-10) = ', correcttotal1 / (6.0 * 20), 'sentence % (1-10) = ', correcttotal4 / (6.0 * 20), 'picture % (10-20) = ', correcttotal2 / (6.0 * 20), 'sentence % (10-20) = ', correcttotal3 / (6.0 * 20)
+total_picture = (((correcttotal1 / (6.0 * 20)) + (correcttotal2 / (6.0 * 20))) / 2) * 100
+total_sentence = (((correcttotal3 / (6.0 * 20)) + (correcttotal4 / (6.0 * 20))) / 2) * 100
+print 'Picture: %s %%' % total_picture
+print 'Sentence: %s %%' % total_sentence 
 print 'Total accuracy: %s %%' % (((correcttotal1 + correcttotal2 + correcttotal3 + correcttotal4) / (6.0 * 80)) * 100)

### 2012-10-23 src/demo/Naive_Bayes_per_scan.py
imporoved printing

@@ -6,95 +6,100 @@ import scipy.stats as sis
 from settings import FILES
 
 # load the matlab files
-data = [sio.loadmat(FILES['NormAvgSDROI'] % (index + 1)) for index in range(6)]
+for datafile in FILES:
+    data = [sio.loadmat(FILES[datafile] % (index + 1)) for index in range(6)]
 
-# make variables
-correcttotal1 = 0
-correcttotal2 = 0
-correcttotal3 = 0
-correcttotal4 = 0
-for iteration in range(0,6):
-    arrr = range(0,20)
-    excl = range(0,20)
-    means = range(0,20) 
-    sds = range(0,20)
-    for n in range(0,20):
-        arrr[n] = [],[]
-        excl[n] = [],[]
-        means[n] = [],[]
-        sds[n] = [],[]
-    means = arrr
+    # make variables
+    correcttotal1 = 0
+    correcttotal2 = 0
+    correcttotal3 = 0
+    correcttotal4 = 0
+    for iteration in range(0,6):
+        arrr = range(0,20)
+        excl = range(0,20)
+        means = range(0,20) 
+        sds = range(0,20)
+        for n in range(0,20):
+            arrr[n] = [],[]
+            excl[n] = [],[]
+            means[n] = [],[]
+            sds[n] = [],[]
+        means = arrr
 
 
-    scans = range(10,20) + range(22,32)
+        scans = range(10,20) + range(22,32)
 
-    #import data
-    exclude = iteration
-    for subject in range(len(data)):
-        counter=-1
-        for scan in scans:
-            counter = counter + 1
-            for trial in range(0,53):
-                if data[subject]['info'][0,trial]['cond'][0]>1:
-                    if data[subject]['info'][0,trial]['firstStimulus'][0]=='P': 
-                        if not subject == exclude:
-                            arrr[counter][0].append(data[subject]['data'][trial][0][scan])
+        #import data
+        exclude = iteration
+        for subject in range(len(data)):
+            counter=-1
+            for scan in scans:
+                counter = counter + 1
+                for trial in range(0,53):
+                    if data[subject]['info'][0,trial]['cond'][0]>1:
+                        if data[subject]['info'][0,trial]['firstStimulus'][0]=='P': 
[…]

### 2012-10-23 src/demo/multivariate.py

@@ -5,117 +5,120 @@ import math as m
 from settings import FILES
 
 # load the matlab files
-data = [sio.loadmat(FILES['AvgROI'] % (index + 1)) for index in range(6)]
+for datafile in FILES:
+    data = [sio.loadmat(FILES[datafile] % (index + 1)) for index in range(6)]
 
-# make variables
-correcttotal1 = 0
-correcttotal2 = 0
-correcttotal3 = 0
-correcttotal4 = 0
+    # make variables
+    correcttotal1 = 0
+    correcttotal2 = 0
+    correcttotal3 = 0
+    correcttotal4 = 0
 
-#repeat the process for 6 subjects
-for iteration in range(0, 6):
-    # arrr = data array for n-1 subjects
-    arrr = range(0, 20)
-    # excl = data for the excluded (test) subject
-    excl = range(0, 20)
-    for n in range(0, 20):
-        arrr[n] = [], []
-        excl[n] = [], []
+    #repeat the process for 6 subjects
+    for iteration in range(0, 6):
+        # arrr = data array for n-1 subjects
+        arrr = range(0, 20)
+        # excl = data for the excluded (test) subject
+        excl = range(0, 20)
+        for n in range(0, 20):
+            arrr[n] = [], []
+            excl[n] = [], []
 
-    #determine the scans
-    firstRange = range(10, 20)
-    secondRange = range(22, 32)
-    scans = firstRange + secondRange
+        #determine the scans
+        firstRange = range(10, 20)
+        secondRange = range(22, 32)
+        scans = firstRange + secondRange
 
-    #import data into arrr and excl
-    exclude = iteration
-    for subject in range(len(data)):
-        counter = -1
-        for scan in scans:
-            counter = counter + 1
-            for trial in range(0, 53):
-                if data[subject]['info'][0, trial]['cond'][0] > 1:
-                    if data[subject]['info'][0, trial]['firstStimulus'][0] == 'P':
-                        if not subject == exclude:
-                            arrr[counter][0].append(data[subject]['data'][trial][0][scan])
+        #import data into arrr and excl
+        exclude = iteration
+        for subject in range(len(data)):
+            counter = -1
+            for scan in scans:
+                counter = counter + 1
+                for trial in range(0, 53):
+                    if data[subject]['info'][0, trial]['cond'][0] > 1:
[…]

### 2012-10-23 src/demo/pca_by_coord.py

@@ -1,6 +1,6 @@
 from __future__ import division
 import scipy.io as sio
-import pylab
+# import pylab
 from scipy import std as sstd
 from scipy import mean as smean
 from functions import *
@@ -11,7 +11,7 @@ from math import fsum
 
 # Data
 print "Loading subjects..."
-subjects = [sio.loadmat(FILES['Orig'] % (index + 1)) for index in range(6)]
+subjects = [sio.loadmat(ORIG_FILES['Orig'] % (index + 1)) for index in range(6)]
 print "Loaded subjects!"
 
 # Configuration
@@ -177,10 +177,10 @@ for n in range(COMPONENTS, COMPONENTS + 1):
 
 
 # Plot?
-pylab.plot(vec_x, vec_y)
-pylab.xlabel('Number of components')
-pylab.ylabel('Accuracy (%)')
-pylab.title('PCA training results for different amounts of components')
-pylab.grid(True)
-pylab.savefig('pca_plot')
-pylab.show()
+# pylab.plot(vec_x, vec_y)
+# pylab.xlabel('Number of components')
+# pylab.ylabel('Accuracy (%)')
+# pylab.title('PCA training results for different amounts of components')
+# pylab.grid(True)
+# pylab.savefig('pca_plot')
+# pylab.show()

### 2012-10-23 src/demo/pca_by_index.py

@@ -20,7 +20,7 @@ Algorithm:  - get voxel indices
 '''
 from __future__ import division
 import scipy.io as sio
-import pylab
+# import pylab
 from scipy import std as sstd
 from scipy import mean as smean
 from functions import *
@@ -28,7 +28,7 @@ from settings import FILES
 from sklearn.naive_bayes import GaussianNB
 
 # Load the dataset
-subjects = [sio.loadmat(FILES['Orig'] % (index + 1)) for index in range(6)]
+subjects = [sio.loadmat(ORIG_FILES['Orig'] % (index + 1)) for index in range(6)]
 
 # Configuration of the classifier. Please indicate the scangroups and the range
 # of PCA components to try.
@@ -190,10 +190,10 @@ for n in range(COMPONENTS, COMPONENTS + 1):
 
 
 # Plot?
-pylab.plot(vec_x, vec_y)
-pylab.xlabel('Number of components')
-pylab.ylabel('Accuracy (%)')
-pylab.title('PCA training results for different amounts of components')
-pylab.grid(True)
-pylab.savefig('pca_plot')
-pylab.show()
+# pylab.plot(vec_x, vec_y)
+# pylab.xlabel('Number of components')
+# pylab.ylabel('Accuracy (%)')
+# pylab.title('PCA training results for different amounts of components')
+# pylab.grid(True)
+# pylab.savefig('pca_plot')
+# pylab.show()

### 2012-10-23 src/demo/settings.py

@@ -1,7 +1,10 @@
 FILES = {
-    'Orig': './matlab/data%d-original.mat',
     'AvgROI': './matlab/data%d_average_rois.mat',
     'NormAvgROI': './matlab/data%d-select-norm-avgroi.mat',
     'NormAvgSDROI': './matlab/data%d-select-norm-avgsdroi.mat',
     'NormPerVoxAvgROI': './matlab/data%d-select-normpervox-avgroi.mat',
+}
+
+ORIG_FILES = {
+	'Orig': './matlab/data%d-original.mat',
 }

### 2012-10-23 src/demo/functions.py
fixed some indentation

@@ -63,29 +63,29 @@ def getCoordinatesForSubject(subject, rois):
 
 
 def getVoxelsForScan(subject, trial_index, scan_index, voxel_indices):
-        '''
-        Returns the voxel vector data for the trial with index trial_index
-        and scan with index scan_index.
+    '''
+    Returns the voxel vector data for the trial with index trial_index
+    and scan with index scan_index.
 
-        Optionally, a list of voxel_indices can be used to further filter the
-        result.
-        '''
-        if voxel_indices:
-            voxels = []
-            for voxel_index in voxel_indices:
-                voxels.append(subject['data'][trial_index][0][scan_index][voxel_index - 1])
-            return voxels
+    Optionally, a list of voxel_indices can be used to further filter the
+    result.
+    '''
+    if voxel_indices:
+        voxels = []
+        for voxel_index in voxel_indices:
+            voxels.append(subject['data'][trial_index][0][scan_index][voxel_index - 1])
+        return voxels
 
-        return subject['data'][trial_index][0][scan_index]
+    return subject['data'][trial_index][0][scan_index]
 
 
 def getValidTrialIndices(subject):
-        '''
-        Returns the valid indices of trials for a subject according to condition.
-        We care only about conditions with value 2 or 3.
-        '''
-        return [index for index in range(subject['meta']['ntrials'])
-                if subject['info'][0]['cond'][index] > 1]
+    '''
+    Returns the valid indices of trials for a subject according to condition.
+    We care only about conditions with value 2 or 3.
+    '''
+    return [index for index in range(subject['meta']['ntrials'])
+            if subject['info'][0]['cond'][index] > 1]
 
 
 def getVoxelsForCoordinates(subject, coords):

### 2012-10-23 src/demo/pca_by_coord.py
pca

@@ -4,7 +4,7 @@ import scipy.io as sio
 from scipy import std as sstd
 from scipy import mean as smean
 from functions import *
-from settings import FILES
+from settings import ORIG_FILES
 from sklearn.naive_bayes import GaussianNB
 #from sklearn.preprocessing import normalize
 from math import fsum

### 2012-10-23 src/demo/pca_by_index.py

@@ -24,7 +24,7 @@ import scipy.io as sio
 from scipy import std as sstd
 from scipy import mean as smean
 from functions import *
-from settings import FILES
+from settings import ORIG_FILES
 from sklearn.naive_bayes import GaussianNB
 
 # Load the dataset

### 2012-10-23 src/demo/gaussian_nb.py
printing for all kind of datafiles

@@ -7,148 +7,151 @@ from sklearn.naive_bayes import GaussianNB
 from math import fsum
 from settings import FILES
 
-# Data
-print "Loading subjects..."
-subjects = [sio.loadmat(FILES['NormPerVoxAvgROI'] % (index + 1)) for index in range(6)]
-print "Loaded subjects!"
+for datafile in FILES:
+    # Data
+    print "ACCURACY FOR DATAFILE %s" % datafile
+    print "Loading subjects..."
+    subjects = [sio.loadmat(FILES[datafile] % (index + 1)) for index in range(6)]
+    print "Loaded subjects!"
 
-# Configuration
-FIRST_STIMULUS_SCAN_INDICES = range(10, 20)
-SECOND_STIMULUS_SCAN_INDICES = range(22, 32)
-COMPONENTS = 200
+    # Configuration
+    FIRST_STIMULUS_SCAN_INDICES = range(10, 20)
+    SECOND_STIMULUS_SCAN_INDICES = range(22, 32)
+    COMPONENTS = 200
 
-# Rois are ignored if filter coords is false
-ROIS = ['CALC', 'LIPL', 'LT', 'LTRIA', 'LOPER', 'LIPS', 'LDLPFC']
+    # Rois are ignored if filter coords is false
+    ROIS = ['CALC', 'LIPL', 'LT', 'LTRIA', 'LOPER', 'LIPS', 'LDLPFC']
 
-FLAG_PCA = False
-FLAG_NORM = False
-FLAG_FILTER_COORDS = False
+    FLAG_PCA = False
+    FLAG_NORM = False
+    FLAG_FILTER_COORDS = False
 
-# If true, average the prediction per set of scans
-FLAG_PER_SCAN = False
+    # If true, average the prediction per set of scans
+    FLAG_PER_SCAN = False
 
-'''
-PCA ALGORITHM
-1) Apply filtering on voxels (for each subject)
-2) Construct data matrix for PCA:
-- each row is a scan from the range of applicable scans (split for type of trial - PS, SP, PS+SP?)
-- each column is a value for a certain coordinate
-3) PCA
-4) Split matrix based on subject
-5) Use one part for training
-6) Use the other part for testing
-'''
-# 1: Get list of coordinates
-validCoords = []
-trials = getValidTrialIndices(subjects[0])
-if FLAG_FILTER_COORDS:
-    print "Constructing nonempty coordinates across subjects"
-    coords = getCoordinatesForSubject(subjects[0], ROIS)
[…]

### 2012-10-23 src/demo/linear_svc.py

@@ -7,145 +7,148 @@ from sklearn.svm import LinearSVC
 from math import fsum
 from settings import FILES
 
-# Data
-print "Loading subjects..."
-subjects = [sio.loadmat(FILES['NormPerVoxAvgROI'] % (index + 1)) for index in range(6)]
-print "Loaded subjects!"
+for datafile in FILES:
+    # Data
+    print "ACCURACY FOR DATAFILE %s" % datafile
+    print "Loading subjects..."
+    subjects = [sio.loadmat(FILES[datafile] % (index + 1)) for index in range(6)]
+    print "Loaded subjects!"
 
-# Configuration
-FIRST_STIMULUS_SCAN_INDICES = range(10, 20)
-SECOND_STIMULUS_SCAN_INDICES = range(22, 32)
-COMPONENTS = 200
+    # Configuration
+    FIRST_STIMULUS_SCAN_INDICES = range(10, 20)
+    SECOND_STIMULUS_SCAN_INDICES = range(22, 32)
+    COMPONENTS = 200
 
-# Rois are ignored if filter coords is false
-ROIS = ['CALC', 'LIPL', 'LT', 'LTRIA', 'LOPER', 'LIPS', 'LDLPFC']
+    # Rois are ignored if filter coords is false
+    ROIS = ['CALC', 'LIPL', 'LT', 'LTRIA', 'LOPER', 'LIPS', 'LDLPFC']
 
-FLAG_PCA = False
-FLAG_NORM = False
-FLAG_FILTER_COORDS = False
+    FLAG_PCA = False
+    FLAG_NORM = False
+    FLAG_FILTER_COORDS = False
 
-# If true, average the prediction per set of scans
-FLAG_PER_SCAN = False
+    # If true, average the prediction per set of scans
+    FLAG_PER_SCAN = False
 
-'''
-PCA ALGORITHM
-1) Apply filtering on voxels (for each subject)
-2) Construct data matrix for PCA:
-- each row is a scan from the range of applicable scans (split for type of trial - PS, SP, PS+SP?)
-- each column is a value for a certain coordinate
-3) PCA
-4) Split matrix based on subject
-5) Use one part for training
-6) Use the other part for testing
-'''
-# 1: Get list of coordinates
-validCoords = []
-trials = getValidTrialIndices(subjects[0])
-if FLAG_FILTER_COORDS:
-    print "Constructing nonempty coordinates across subjects"
-    coords = getCoordinatesForSubject(subjects[0], ROIS)
-    for coord in coords:
[…]

### 2012-11-06 article/article.tex
general introduction

@@
 %% Now the actual contents of first section
-Traditionally, it has been a challenge to relate between a simple cognitive state (such as `looking at a triangle`) and the corresponding activity of the brain \cite{Pone0001394}.
+Human functional brain mapping is a field of study that has grown strongly in the past decades.
+This is not a surprise, because it is one of the few ways to look at brain activity of a living animal.
+There are several techniques available to scan for brain activity in a living human.
+Of these techniques, functional magnetic resonance imaging (fMRI) has become more and more popular because it has a very good spacial resolution (contrary to MEG/EEG), and it is non-invasive (contrary to PET).
+Brain activity is inferred from differences in the ratio between oxygenated and deoxygenated blood; when brain activity occurs, more oxygenated blood will flow to a brain region, and this change is detected by the machine.
+This is called the blood oxygen level dependant (BOLD) response \cite{sm:2012fMRI}.\\
+\indent The fMRI scanner measures the activity in three-dimensional pixels called voxels, which are about 3 x 3 x 3 mm in size, and a brain scan is made somewhere between every 500ms and every few seconds.
+However, before analysing the data is it usually pre-processed to account for head movements by participants in the scanner, and to make between-subjects comparisons possible \cite{sl:2009rl}.\\
+\indent Traditionally, data from each voxel, or from a group of voxels, are analysed with analysis of variance.
+Voxels are usually convoluted with a haemodynamic response function in an attempt to infer the spiking neural activity from the slowly building and dropping BOLD response \cite{he:2002na}.
+Experimental conditions are the independent variables in this setup, and a general linear model is fitted to each voxel, or group of voxels.
+By using this technique voxels or brain areas can be identified that differ in average BOLD response across conditions.\\
+\indent Alternatively, Machine learning algorithms have been employed to learn the patterns of activity associated with some experimental condition.
+Subsequently, these learned patterns can be used on unseen subjects in order to predict experimental conditions from their brain activity patterns.
+When these predictions are accurate then it means that the identified patterns hold information about the brain activity associated with some experimental task.\\
+\indent These machine learning techniques have been successfully used for decoding the orientation of a stimulus from fMRI scans \cite{ka:2005decoding}, identifying different patterns of activity for animate versus inanimate objects \cite{kr:2008RSA}, and even reconstructing visual input based on information in the fMRI scans \cite{mi:2008re}.
+The reason that these machine learning techniques work so well is that patterns of voxel activity can be identified rather than a significant deviation from some baseline.
+Specifically, information can be present in the pattern of voxels that belong to a certain brain region, even when the average signal from that region is zero (i.e., on average it does not differ from the baseline condition) \cite{to:2012rev}.
+\\
+
+%Traditionally, it has been a challenge to relate between a simple cognitive state (such as `looking at a triangle`) and the corresponding activity of the brain \cite{Pone0001394}.
 This correspondence may even vary from subject to subject, making the challenge even harder.

### 2012-11-08 article/article.tex
ins alinea line

@@
 There are several techniques available to scan for brain activity in a living human.
-Of these techniques, functional magnetic resonance imaging (fMRI) has become more and more popular because it has a very good spacial resolution (contrary to MEG/EEG), and it is non-invasive (contrary to PET).
+Of these techniques, functional magnetic resonance imaging (fMRI) has become more and more popular because it has a very good spatial resolution (contrary to MEG/EEG), and it is non-invasive (contrary to PET).
 Brain activity is inferred from differences in the ratio between oxygenated and deoxygenated blood; when brain activity occurs, more oxygenated blood will flow to a brain region, and this change is detected by the machine.
@@
 By using this technique voxels or brain areas can be identified that differ in average BOLD response across conditions.\\
+\indent Insert Alinea about large scale fMRI studies indicating that "everything is significant", thus identifying patterns makes more sense than identifying areas that differ from baseline. \\
 \indent Alternatively, Machine learning algorithms have been employed to learn the patterns of activity associated with some experimental condition.

### 2012-11-08 article/article.tex
ins alinea line

@@
 \\
+\indent Insert alinea discussing Starplus data. - related work and results\\
+\indent Insert alinea to explain shortly what our goal and used classifiers are (explained further in application?)\\
 
@@
 
-- basic terms introduced (fmri, classification, ...)
-- related work and results
 - briefly build up to state of the art, mention methods we tried?

### 2012-11-08 article/article.tex
ins alinea line

@@
 By using this technique voxels or brain areas can be identified that differ in average BOLD response across conditions.\\
-\indent Insert Alinea about large scale fMRI studies indicating that "everything is significant", thus identifying patterns makes more sense than identifying areas that differ from baseline. \\
+\indent Recent studies indicate that this approach may have some serious limitations.
+One study used a dataset of 1326 subjects, which is much higher than the usual 15 to 25 subjects found in fMRI studies \cite{thy:2012very}.
+This study explored what happened to a simple contrast between BOLD signal while watching faces and baseline activity, when increasing the number of subjects included in the analysis.
+They found that brain areas that extended well beyond the expected visual- and face recognition areas showed significant deviation from baseline.
+Another study, using only three subjects, repeated a simple letter/number discrimination task and visual stimulation.
+The notable part is that this very simple experiment was run 100 times for each subject, again resulting in exceptionally high signal-to-noise ratio compared to conventional fMRI experiments.
+In this study the authors again found widespread significant BOLD-responses \cite{go:2012whole}.
+These findings suggest that significant deviations from baseline activity do not necessarily indicate any special involvement in the performed task, because almost the entire brain will show significant deviations given a high enough signal-to-noise ratio.
+These findings are perhaps not surprising, because everything we know about the brain indicates that it best seen as a network with countless connections to other parts of the brain.
+Therefore, it makes sense to also analyse signals from the brain as a pattern of information coming from the entire brain, rather attempting to link specific brain areas to experimental conditions.\\
 \indent Alternatively, Machine learning algorithms have been employed to learn the patterns of activity associated with some experimental condition.

### 2012-11-08 article/article.tex
added alineas intro

@@
 Therefore, it makes sense to also analyse signals from the brain as a pattern of information coming from the entire brain, rather attempting to link specific brain areas to experimental conditions.\\
-\indent Alternatively, Machine learning algorithms have been employed to learn the patterns of activity associated with some experimental condition.
+\indent In fact, Machine learning algorithms have been employed to learn the patterns of activity associated with some experimental condition.
 Subsequently, these learned patterns can be used on unseen subjects in order to predict experimental conditions from their brain activity patterns.
@@
 \\
-\indent Insert alinea discussing Starplus data. - related work and results\\
+\indent  The general goal of these machine learning studies is to identify what experimental condition a subject is in at a certain time.
+The state that a person is in during some trial (e.g., looking at a face) is referred to as emph{cognitive state}, because the cognitive processes that are associated with being in that state are of interest.
+Specifically, the aforementioned studies try to infer the cognitive state of a subject from the brain activity that is evoked in a subject by performing that trial.
+For example, in an experiment where subjects look at a set of pictures (e.g., a star and a plus), they would be in a certain cognitive state that is the result of looking at these pictures.
+One could then attempt to use machine learning on the fMRI signal that is evoked by the the brain activity associated with these cognitive states, and use this to learn for example what patterns of activity in the brain is associated with that cognitive state.
+This \emph{StarPlus} experiment has been performed, where in addition to viewing pictures, subjects also read sentences that contained some true or false statement about the pictures.
+These data have then been used to successfully infer the cognitive states of looking at either sentences or pictures within single subjects\cite{mi:2003within}, and between subjects \cite{wa:2003betw}.
+\\Insert alinea discussing Starplus data.
+- related work and results\\
 \indent Insert alinea to explain shortly what our goal and used classifiers are (explained further in application?)\\

### 2012-11-08 article/article.tex
intro

@@
 This \emph{StarPlus} experiment has been performed, where in addition to viewing pictures, subjects also read sentences that contained some true or false statement about the pictures.
-These data have then been used to successfully infer the cognitive states of looking at either sentences or pictures within single subjects\cite{mi:2003within}, and between subjects \cite{wa:2003betw}.
-\\Insert alinea discussing Starplus data.
-- related work and results\\
+These data have then been used to successfully infer the cognitive states of looking at either sentences or pictures within single subjects \cite{mi:2003within}, and between subjects \cite{wa:2003betw}.
+\\
 \indent Insert alinea to explain shortly what our goal and used classifiers are (explained further in application?)\\

### 2012-11-09 article/article.tex
minor layout, abstract left to fix
genious alinea added

@@
 %% References
-\documentclass[preprint,journal]{vgtc}
+\documentclass[preprint,journal,11pt]{vgtc}
 \let\ifpdf\relax
@@
 \item
-  Spyros Ioakeimidis  is with University of Groningen, E-mail: spyrosikmd@gmail.com.
+  Spyros Ioakeimidis  is with University of Groningen, E-mail: s.ioakeimidis@studenr.rug.nl.
 }
@@
 \\
-\indent Insert alinea to explain shortly what our goal and used classifiers are (explained further in application?)\\
+\indent In this study we use the same StarPlus data, and use machine learning algorithms to decode the cognitive states of subjects when they are looking at Pictures or when they are looking at sentences.
+An interesting difference between many applications of machine learning, and this study is that most machine learning algorithms have the goal of achieving the maximum possible discrimination between conditions or classes.
+However, for fMRI studies it is important that the pattern of activity that is identified by the algorithm is a pattern associated with the cognitive state.
+When using non-linear classifiers, it is possible to decode features predictive of the cognitive state, where the brain does not actually respond to these features.
+For example, when decoding the primary visual cortex with high enough spatial resolution, it would be possible to combine simple features with a non-linear neural network to decode the presence or absence of, for example, faces.
+However, this information is not actually processed in that part of the brain \cite{to:2012rev}.
+To avoid this problem it is recommended to only use linear classifiers when decoding cognitive states from fMRI data.
+The classifiers that we compare in this paper are Gaussian Naive Bayesian (GNB) classifiers, Linear Support Vector (LSV) machines, and a Multivariate Gaussian (MV) classifier.
+       Insert alinea to explain shortly what our goal and used classifiers are (explained further in application?)\\
 

### 2012-11-09 article/article.tex
spelling

@@
 \\
-\indent In this study we use the same StarPlus data, and use machine learning algorithms to decode the cognitive states of subjects when they are looking at Pictures or when they are looking at sentences.
+\indent In this study we use the same StarPlus data from the previously mentioned experiment, and use machine learning algorithms to decode the cognitive states of subjects when they are looking at pictures or when they are looking at sentences.
 An interesting difference between many applications of machine learning, and this study is that most machine learning algorithms have the goal of achieving the maximum possible discrimination between conditions or classes.

### 2012-11-09 article/article.tex
hi there
dunno what

--- 2012-11-09 article/article.tex
@@
 These findings suggest that significant deviations from baseline activity do not necessarily indicate any special involvement in the performed task, because almost the entire brain will show significant deviations given a high enough signal-to-noise ratio.
-These findings are perhaps not surprising, because everything we know about the brain indicates that it best seen as a network with countless connections to other parts of the brain.
+These findings are perhaps not surprising, because everything we know about the brain indicates that it is best seen as a network with countless connections to other parts of the brain.
 Therefore, it makes sense to also analyse signals from the brain as a pattern of information coming from the entire brain, rather attempting to link specific brain areas to experimental conditions.\\
@@
 \\
-\indent In this study we use the same StarPlus data, and use machine learning algorithms to decode the cognitive states of subjects when they are looking at Pictures or when they are looking at sentences.
+\indent In this study we use the same StarPlus data as in the previously mentioned experiment, and use machine learning algorithms to decode the cognitive states of subjects when they are looking at pictures or when they are looking at sentences.
 An interesting difference between many applications of machine learning, and this study is that most machine learning algorithms have the goal of achieving the maximum possible discrimination between conditions or classes.
@@
 To avoid this problem it is recommended to only use linear classifiers when decoding cognitive states from fMRI data.
-The classifiers that we compare in this paper are Gaussian Naive Bayesian (GNB) classifiers, Linear Support Vector (LSV) machines, and a Multivariate Gaussian (MV) classifier.
+The classifiers that we compare in this paper are Gaussian Naive Bayesian (GNB) classifiers, Linear Support Vector (LSV) machines, and a Multivariate Gaussian (MV) classifier.\\
        Insert alinea to explain shortly what our goal and used classifiers are (explained further in application?)\\

### 2012-11-09 article/article.tex
yo
spelling

@@
 %% indicate IEEE Member or Student Member in form indicated below
-\author{Mattijs Meiboom, Robin Mills, Tijn Schouten, Spyros Ioakeimidis}
+\author{Mattijs Meiboom, 
+, Tijn Schouten, Spyros Ioakeimidis}
 \authorfooter{
@@
 \item
-  Robin Mills is with University of Groningen, E-mail: .
+  Robin Mills is with University of Groningen, E-mail: r.mills@student.rug.nl.
 \item
@@
 To avoid this problem it is recommended to only use linear classifiers when decoding cognitive states from fMRI data.
-The classifiers that we compare in this paper are Gaussian Naive Bayesian (GNB) classifiers, Linear Support Vector (LSV) machines, and a Multivariate Gaussian (MV) classifier.\\
+The classifiers that we compare in this paper are Gaussian Naive Bayesian (GNB) classifiers, Linear Support Vector Machine (LSVM), and a Multivariate Gaussian (MVG) classifier.\\
        Insert alinea to explain shortly what our goal and used classifiers are (explained further in application?)\\

### 2012-11-09 article/article.tex
results

--- 2012-11-09 article/article.tex
@@
 %% Abstract section.
-\abstract{
-Functional magnetic resonance imaging (fMRI) is a brain mapping technique that allows researchers to measure activity in a subject's brain.
+\abstract{Functional magnetic resonance imaging (fMRI) is a brain mapping technique that allows researchers to measure activity in a subject's brain.
 This technique is typically used to determine which areas of the brain are involved in performing a certain task.
@@
 One could then attempt to use machine learning on the fMRI signal that is evoked by the the brain activity associated with these cognitive states, and use this to learn for example what patterns of activity in the brain is associated with that cognitive state.
-This \emph{StarPlus} experiment has been performed, where in addition to viewing pictures, subjects also read sentences that contained some true or false statement about the pictures.
+This \emph{StarPlus} \cite{} experiment has been performed, where in addition to viewing pictures, subjects also read sentences that contained some true or false statement about the pictures.
 These data have then been used to successfully infer the cognitive states of looking at either sentences or pictures within single subjects \cite{mi:2003within}, and between subjects \cite{wa:2003betw}.
 \\
-\indent In this study we use the same StarPlus data from the previously mentioned experiment, and use machine learning algorithms to decode the cognitive states of subjects when they are looking at pictures or when they are looking at sentences.
-An interesting difference between many applications of machine learning, and this study is that most machine learning algorithms have the goal of achieving the maximum possible discrimination between conditions or classes.
-However, for fMRI studies it is important that the pattern of activity that is identified by the algorithm is a pattern associated with the cognitive state.
-When using non-linear classifiers, it is possible to decode features predictive of the cognitive state, where the brain does not actually respond to these features.
-For example, when decoding the primary visual cortex with high enough spatial resolution, it would be possible to combine simple features with a non-linear neural network to decode the presence or absence of, for example, faces.
-However, this information is not actually processed in that part of the brain \cite{to:2012rev}.
-To avoid this problem it is recommended to only use linear classifiers when decoding cognitive states from fMRI data.
-The classifiers that we compare in this paper are Gaussian Naive Bayesian (GNB) classifiers, Linear Support Vector (LSV) machines, and a Multivariate Gaussian (MV) classifier.
-       Insert alinea to explain shortly what our goal and used classifiers are (explained further in application?)\\
+\indent Insert alinea to explain shortly what our goal and used classifiers are (explained further in application?)\\
 
@@
 \label{sec:results}
-- raw results per algorithm
+
+Table~\ref{tab:results}
+
+explain the symbols on the table, explain what rows (feature selection methods) are and what columns are (different classifiers and some of them with different structure on the data)
+
+first column naive bayes with pca 148 components, others common feature selection methods as described in Section~\ref{sec:featureSelection}
+
+* means different structure of the data
+
+classifiers performance do not differ significantly
+
+because even PCA has accuracy 61.5\% we can be sure that all the classifiers differ significantly from 50\%
+
+mention confidence interval (how much percent sure we are that the actual accuracy of the classifier is between these regions)
+
+\begin{table*}[htdp]
+\centering
+{\small
+\begin{tabular}{c|c|c|c|c|c}
+& \textbf{Naive Bayes} & \textbf{Naive Bayes} & \textbf{Linear SVC} & \textbf{Naive Bayes*} & \textbf{Multivariate*} \\ \hline
+PCA & 61.5 \% &  - & - & - & - \\
+M-ROIs & - & 70.2 \% & 72.5 \% & 74.8 \% & 76.7 \% \\
+Norm-M-ROIs & - & 71.7 \% & 71.3 \% & 75.8 \% & 77.1 \% \\
+Norm-M-SD-ROIs & - & 65.6 \% & 69.8 \% & 69.4 \% & 65.0 \%\\
+Norm/Vox-M-ROIs & - & 73.1 \% & 76.7 \% & 76.9 \% & 79.6 \% \\
+\end{tabular}}
+\caption{Performance results of three classifiers using different methods for feature selection and structure.}
+\label{tab:results}
+\end{table*}
 

### 2012-11-09 article/article.tex

--- 2012-11-09 article/article.tex
@@
 %% indicate IEEE Member or Student Member in form indicated below
-\author{Mattijs Meiboom, 
+\author{Mattijs Meiboom, Robin Mills
 , Tijn Schouten, Spyros Ioakeimidis}

### 2012-11-09 article/article.tex
title
hi

--- 2012-11-09 article/article.tex
@@
 %% indicate IEEE Member or Student Member in form indicated below
-\author{Mattijs Meiboom, 
-, Tijn Schouten, Spyros Ioakeimidis}
+\author{Mattijs Meiboom, Robin Mills, Tijn Schouten, Spyros Ioakeimidis}
 \authorfooter{
@@
 To avoid this problem it is recommended to only use linear classifiers when decoding cognitive states from fMRI data.
-The classifiers that we compare in this paper are Gaussian Naive Bayesian (GNB) classifiers, Linear Support Vector Machine (LSVM), and a Multivariate Gaussian (MVG) classifier.\\
+The classifiers that we compare in this paper are Gaussian Naive Bayesian (GNB) classifiers, Linear Support Vector (LSV) machines, and a Multivariate Gaussian (MV) classifier.\\
        Insert alinea to explain shortly what our goal and used classifiers are (explained further in application?)\\
@@
     - svm
+    - why multivariate is not a priori better than naive: less assumption, higher generalizability, better estimation parameters.
+
+\section{Data Normalization}
+\label{sec:normalization}
+- how did we normalize?
 

### 2012-11-09 article/article.tex

@@
 
-\title{Bayes network classifiers for fMRI}
+\title{Comparing Machine Learning algorithms for classifying cognitive states from fMRI data}
 

### 2012-11-09 article/article.tex
results

--- 2012-11-09 article/article.tex
@@
 
-Table~\ref{tab:results}
+As a comparison measure, we used the performance of the classifiers on predicting the cognitive state that the subject was in.
+Table~\ref{tab:results} illustrates the final results.
+Rows indicate various feature selection methods.
+The letter \emph{M} means Columns depict the used classifiers.
+For the classifiers depicted with (*), we used different structure for the data that we fit to them (take example from section where the selection is introduced).
 
@@
 {\small
-\begin{tabular}{c|c|c|c|c|c}
-& \textbf{Naive Bayes} & \textbf{Naive Bayes} & \textbf{Linear SVC} & \textbf{Naive Bayes*} & \textbf{Multivariate*} \\ \hline
-PCA & 61.5 \% &  - & - & - & - \\
-M-ROIs & - & 70.2 \% & 72.5 \% & 74.8 \% & 76.7 \% \\
-Norm-M-ROIs & - & 71.7 \% & 71.3 \% & 75.8 \% & 77.1 \% \\
-Norm-M-SD-ROIs & - & 65.6 \% & 69.8 \% & 69.4 \% & 65.0 \%\\
-Norm/Vox-M-ROIs & - & 73.1 \% & 76.7 \% & 76.9 \% & 79.6 \% \\
+\begin{tabular}{c|c|c|c|c}
+& \textbf{Naive Bayes} & \textbf{Linear SVC} & \textbf{Naive Bayes*} & \textbf{Multivariate*} \\ \hline
+PCA & 61.5 \% &  - & - & - \\
+M-ROIs & 70.2 \% & 72.5 \% & 74.8 \% & 76.7 \% \\
+Norm-M-ROIs & 71.7 \% & 71.3 \% & 75.8 \% & 77.1 \% \\
+Norm-M-SD-ROIs & 65.6 \% & 69.8 \% & 69.4 \% & 65.0 \%\\
+Norm/Vox-M-ROIs & 73.1 \% & 76.7 \% & 76.9 \% & 79.6 \% \\
 \end{tabular}}

### 2012-11-09 article/article.tex

@@
 %% indicate IEEE Member or Student Member in form indicated below
-\author{Mattijs Meiboom, Robin Mills
-, Tijn Schouten, Spyros Ioakeimidis}
+\author{Mattijs Meiboom, Robin Mills, Tijn Schouten, Spyros Ioakeimidis}
 \authorfooter{

### 2012-11-09 article/article.tex
hi
inserted ROI paragraph

--- 2012-11-09 article/article.tex
@@
 %% indicate IEEE Member or Student Member in form indicated below
-\author{Mattijs Meiboom, Robin Mills, Tijn Schouten, Spyros Ioakeimidis}
+\author{Mattijs Meiboom, Robin Mills
+, Tijn Schouten, Spyros Ioakeimidis}
 \authorfooter{
@@
 The classifiers that we compare in this paper are Gaussian Naive Bayesian (GNB) classifiers, Linear Support Vector Machine (LSVM), and a Multivariate Gaussian (MVG) classifier.\\
-       Insert alinea to explain shortly what our goal and used classifiers are (explained further in application?)\\
+\indent	The goal of this study is to compare the classification performance of the cognitive states associated with looking at a picture or looking at a sentence between subject.
+This objective introduces an important issue, namely, how the data from the fMRI scans can be compared between subjects.
+There are two mains ways this could be achieved.
+First, groups of voxels that are anatomically defined as belonging to a certain brain region, a region of interest (ROI), can be averaged into a single \emph{supervoxel}, which can be compared between subjects.
+Second, an anatomically defined talarach coordinate system can be used that maps the voxels of each subject onto an anatomically defined coordinate.
+The ROI averaging it is a rather crude method, and a lot of information of the individual voxels is lost, however, it also doubles as a dimensionality reduction method.
+By using the talarach coordinate system the data of each individual voxel can be used, however, to use this system the anatomical information for each subject has to be known and mapped onto the fMRI data in great detail.
+Because we lack the detailed data to precisely use the coordinate system we will use averaged ROI's in order to compare the classifiers.<insert citation in this paragraph> \\
 
@@
     - svm
-    - why multivariate is not a priori better than naive: less assumption, higher generalizability, better estimation parameters.
-
-\section{Data Normalization}
-\label{sec:normalization}
-- how did we normalize?
 

### 2012-11-09 article/article.tex

@@
     - svm
+    - why multivariate is not a priori better than naive: less assumption, higher generalizability, better estimation parameters.
 

### 2012-11-09 article/article.tex
results

--- 2012-11-09 article/article.tex
@@
 As a comparison measure, we used the performance of the classifiers on predicting the cognitive state that the subject was in.
-Table~\ref{tab:results} illustrates the final results.
+Table~\ref{tab:results} illustrates the overall performance of .
 Rows indicate various feature selection methods.
@@
 & \textbf{Naive Bayes} & \textbf{Linear SVC} & \textbf{Naive Bayes*} & \textbf{Multivariate*} \\ \hline
-PCA & 61.5 \% &  - & - & - \\
-M-ROIs & 70.2 \% & 72.5 \% & 74.8 \% & 76.7 \% \\
-Norm-M-ROIs & 71.7 \% & 71.3 \% & 75.8 \% & 77.1 \% \\
-Norm-M-SD-ROIs & 65.6 \% & 69.8 \% & 69.4 \% & 65.0 \%\\
-Norm/Vox-M-ROIs & 73.1 \% & 76.7 \% & 76.9 \% & 79.6 \% \\
+PCA & 61.46 &  - & - & - \\
+M-ROIs & 70.21 & 72.5 & 74.79 & 76.67 \\
+Norm-M-ROIs & 71.67 & 71.25 & 75.83 & 77.08 \\
+Norm-M-SD-ROIs & 65.63 & 69.79 & 69.38 & 65.0 \\
+Norm/Vox-M-ROIs & 73.13 & 76.67 & 76.88 & 79.58 \\
 \end{tabular}}
-\caption{Performance results of three classifiers using different methods for feature selection and structure.}
+\caption{Performance results (in \%) of the three classifiers for different methods for feature selection and structure.}
 \label{tab:results}

### 2012-11-09 article/article.tex

@@
 
-\section{Data Normalization}
-\label{sec:normalization}
-- how did we normalize?
-
 \section{Experimental Results}

### 2012-11-09 article/article.tex
hoi

--- 2012-11-09 article/article.tex
@@
 The classifiers that we compare in this paper are Gaussian Naive Bayesian (GNB) classifiers, Linear Support Vector Machine (LSVM), and a Multivariate Gaussian (MVG) classifier.\\
-\indent	The goal of this study is to compare the classification performance of the cognitive states associated with looking at a picture or looking at a sentence between subject.
-This objective introduces an important issue, namely, how the data from the fMRI scans can be compared between subjects.
-There are two mains ways this could be achieved.
-First, groups of voxels that are anatomically defined as belonging to a certain brain region, a region of interest (ROI), can be averaged into a single \emph{supervoxel}, which can be compared between subjects.
-Second, an anatomically defined talarach coordinate system can be used that maps the voxels of each subject onto an anatomically defined coordinate.
-The ROI averaging it is a rather crude method, and a lot of information of the individual voxels is lost, however, it also doubles as a dimensionality reduction method.
-By using the talarach coordinate system the data of each individual voxel can be used, however, to use this system the anatomical information for each subject has to be known and mapped onto the fMRI data in great detail.
-Because we lack the detailed data to precisely use the coordinate system we will use averaged ROI's in order to compare the classifiers.<insert citation in this paragraph> \\
+       Insert alinea to explain shortly what our goal and used classifiers are (explained further in application?)\\
 
@@
 
+To reducde the dimensionality of the input, two different methods are used.
+ROI mapping. 
 So how do we cope with this problem?

### 2012-11-09 article/article.tex
results

--- 2012-11-09 article/article.tex
@@
 As a comparison measure, we used the performance of the classifiers on predicting the cognitive state that the subject was in.
-Table~\ref{tab:results} illustrates the overall performance of .
+Table~\ref{tab:results} illustrates the overall performance of the three classifiers.
 Rows indicate various feature selection methods.
-The letter \emph{M} means Columns depict the used classifiers.
+The letter \emph{M} denotes Columns depict the used classifiers.
 For the classifiers depicted with (*), we used different structure for the data that we fit to them (take example from section where the selection is introduced).
 
-explain the symbols on the table, explain what rows (feature selection methods) are and what columns are (different classifiers and some of them with different structure on the data)
+%explain the symbols on the table, explain what rows (feature selection methods) are and what columns are (different classifiers and some of them with different structure on the data)
 
-first column naive bayes with pca 148 components, others common feature selection methods as described in Section~\ref{sec:featureSelection}
+%first column naive bayes with pca 148 components, others common feature selection methods as described in Section~\ref{sec:featureSelection}
 
-* means different structure of the data
+%* means different structure of the data
 
-classifiers performance do not differ significantly
+%classifiers performance do not differ significantly
 
-because even PCA has accuracy 61.5\% we can be sure that all the classifiers differ significantly from 50\%
+%because even PCA has accuracy 61.5\% we can be sure that all the classifiers differ significantly from 50\%
 
-mention confidence interval (how much percent sure we are that the actual accuracy of the classifier is between these regions)
+%mention confidence interval (how much percent sure we are that the actual accuracy of the classifier is between these regions)
 

### 2012-11-09 article/article.tex
introduction

--- 2012-11-09 article/article.tex
@@
 The classifiers that we compare in this paper are Gaussian Naive Bayesian (GNB) classifiers, Linear Support Vector Machine (LSVM), and a Multivariate Gaussian (MVG) classifier.\\
-       Insert alinea to explain shortly what our goal and used classifiers are (explained further in application?)\\
+\indent	The goal of this study is to compare the classification performance of the cognitive states associated with looking at a picture or looking at a sentence between subject.
+This objective introduces an important issue, namely, how the data from the fMRI scans can be compared between subjects.
+There are two mains ways this could be achieved.
+First, groups of voxels that are anatomically defined as belonging to a certain brain region, a region of interest (ROI), can be averaged into a single \emph{supervoxel}, which can be compared between subjects.
+Second, an anatomically defined talairach coordinate system can be used that maps the voxels of each subject onto an anatomically defined coordinate.
+The ROI averaging it is a rather crude method, and a lot of information of the individual voxels is lost, however, it also doubles as a dimensionality reduction method.
+By using the talairach coordinate system the data of each individual voxel can be used, however, to use this system the anatomical information for each subject has to be known and mapped onto the fMRI data in great detail.
+Because we lack the detailed data to precisely use the coordinate system we will use averaged ROI's in order to compare the classifiers.<insert citation in this paragraph> \\
 
@@
 
-To reducde the dimensionality of the input, two different methods are used.
-ROI mapping. 
 So how do we cope with this problem?

### 2012-11-09 article/article.tex

--- 2012-11-09 article/article.tex
@@
 First, groups of voxels that are anatomically defined as belonging to a certain brain region, a region of interest (ROI), can be averaged into a single \emph{supervoxel}, which can be compared between subjects.
-Second, an anatomically defined talarach coordinate system can be used that maps the voxels of each subject onto an anatomically defined coordinate.
+Second, an anatomically defined talairach coordinate system can be used that maps the voxels of each subject onto an anatomically defined coordinate.
 The ROI averaging it is a rather crude method, and a lot of information of the individual voxels is lost, however, it also doubles as a dimensionality reduction method.
-By using the talarach coordinate system the data of each individual voxel can be used, however, to use this system the anatomical information for each subject has to be known and mapped onto the fMRI data in great detail.
+By using the talairach coordinate system the data of each individual voxel can be used, however, to use this system the anatomical information for each subject has to be known and mapped onto the fMRI data in great detail.
 Because we lack the detailed data to precisely use the coordinate system we will use averaged ROI's in order to compare the classifiers.<insert citation in this paragraph> \\

### 2012-11-09 article/article.tex
yo nigger

--- 2012-11-09 article/article.tex
@@
 The classifiers that we compare in this paper are Gaussian Naive Bayesian (GNB) classifiers, Linear Support Vector Machine (LSVM), and a Multivariate Gaussian (MVG) classifier.\\
-\indent	The goal of this study is to compare the classification performance of the cognitive states associated with looking at a picture or looking at a sentence between subject.
-This objective introduces an important issue, namely, how the data from the fMRI scans can be compared between subjects.
-There are two mains ways this could be achieved.
-First, groups of voxels that are anatomically defined as belonging to a certain brain region, a region of interest (ROI), can be averaged into a single \emph{supervoxel}, which can be compared between subjects.
-Second, an anatomically defined talairach coordinate system can be used that maps the voxels of each subject onto an anatomically defined coordinate.
-The ROI averaging it is a rather crude method, and a lot of information of the individual voxels is lost, however, it also doubles as a dimensionality reduction method.
-By using the talairach coordinate system the data of each individual voxel can be used, however, to use this system the anatomical information for each subject has to be known and mapped onto the fMRI data in great detail.
-Because we lack the detailed data to precisely use the coordinate system we will use averaged ROI's in order to compare the classifiers.<insert citation in this paragraph> \\
+       Insert alinea to explain shortly what our goal and used classifiers are (explained further in application?)\\
 
@@
 
+To reducde the dimensionality of the input, two different methods are used.
+
+ROI mapping. 
 So how do we cope with this problem?
@@
 - ROIs, means, stddev, skew, kurtosis?
+
+\subsection{Classifiers}
 

### 2012-11-09 article/article.tex
results

--- 2012-11-09 article/article.tex
@@
 
-As a comparison measure, we used the performance of the classifiers on predicting the cognitive state that the subject was in.
+As a comparison measure, we used the performance of the classifiers for predicting the cognitive state that the subject was in.
 Table~\ref{tab:results} illustrates the overall performance of the three classifiers.
-Rows indicate various feature selection methods.
-The letter \emph{M} denotes Columns depict the used classifiers.
-For the classifiers depicted with (*), we used different structure for the data that we fit to them (take example from section where the selection is introduced).
+Rows indicate various feature selection methods, where \emph{M} stands for \emph{Mean}, \emph{Norm} for \emph{Normalisation}, and \emph{SD} for \emph{Standard Deviation}.
+For instance, \emph{Norm-M-SD ROIs} denotes that normalisation was used in the original data and then mean and standard deviation was applied to them as a selection method.
+
+Columns depict the used classifiers. For the classifiers depicted with (*), we used different structure for the data that we fit to them (take example from section where the selection is introduced).
 
@@
 PCA & 61.46 &  - & - & - \\
-M-ROIs & 70.21 & 72.5 & 74.79 & 76.67 \\
-Norm-M-ROIs & 71.67 & 71.25 & 75.83 & 77.08 \\
-Norm-M-SD-ROIs & 65.63 & 69.79 & 69.38 & 65.0 \\
-Norm/Vox-M-ROIs & 73.13 & 76.67 & 76.88 & 79.58 \\
+M ROIs & 70.21 & 72.5 & 74.79 & 76.67 \\
+Norm-M ROIs & 71.67 & 71.25 & 75.83 & 77.08 \\
+Norm-M-SD ROIs & 65.63 & 69.79 & 69.38 & 65.0 \\
+Norm/Vox-M ROIs & 73.13 & 76.67 & 76.88 & 79.58 \\
 \end{tabular}}
-\caption{Performance results (in \%) of the three classifiers for different methods for feature selection and structure.}
+\caption{Performance results (in \%) of the three classifiers on different methods for feature selection and structure.}
 \label{tab:results}

### 2012-11-09 article/article.tex
results

@@
 Rows indicate various feature selection methods, where \emph{M} stands for \emph{Mean}, \emph{Norm} for \emph{Normalisation}, and \emph{SD} for \emph{Standard Deviation}.
-For instance, \emph{Norm-M-SD ROIs} denotes that normalisation was used in the original data and then mean and standard deviation was applied to them as a selection method.
+For instance, \emph{Norm-M-SD ROIs} denotes that normalisation was used in the voxels belonging to the selected ROIs over the mean and standard deviation.
 
-Columns depict the used classifiers. For the classifiers depicted with (*), we used different structure for the data that we fit to them (take example from section where the selection is introduced).
+Columns depict the used classifiers---Naive Bayes, Linear SVM, and Multivariate.
+For the classifiers depicted with *, we used different structure for the data that we fit to them, as described in Section~\ref{}.
+The table reveals the following general results:
 
-%explain the symbols on the table, explain what rows (feature selection methods) are and what columns are (different classifiers and some of them with different structure on the data)
+\begin{itemize}
+	\item Classifiers performance do not differ significantly between them.
+	\item Classifiers performance differ significantly from 50\%.
+\end{itemize}
 
 %first column naive bayes with pca 148 components, others common feature selection methods as described in Section~\ref{sec:featureSelection}
-
-%* means different structure of the data
 
@@
 \begin{tabular}{c|c|c|c|c}
-& \textbf{Naive Bayes} & \textbf{Linear SVC} & \textbf{Naive Bayes*} & \textbf{Multivariate*} \\ \hline
+& \textbf{Naive Bayes} & \textbf{Linear SVM} & \textbf{Naive Bayes*} & \textbf{Multivariate*} \\ \hline
 PCA & 61.46 &  - & - & - \\

### 2012-11-09 article/article.tex
Peter jackson

--- 2012-11-09 article/article.tex
@@
 
+\subsection{Preprocessing}
+
 \subsection{Feature selection}
@@
 ROI mapping.
+
+Principal Component Analysis (PCA).
+
 So how do we cope with this problem?

### 2012-11-09 article/article.tex
improved roi paragraph

@@
 Brain activity is inferred from differences in the ratio between oxygenated and deoxygenated blood; when brain activity occurs, more oxygenated blood will flow to a brain region, and this change is detected by the machine.
-This is called the blood oxygen level dependant (BOLD) response \cite{sm:2012fMRI}.\\
-\indent The fMRI scanner measures the activity in three-dimensional pixels called voxels, which are about 3 x 3 x 3 mm in size, and a brain scan is made somewhere between every 500ms and every few seconds.
-However, before analysing the data is it usually pre-processed to account for head movements by participants in the scanner, and to make between-subjects comparisons possible \cite{sl:2009rl}.\\
-\indent Traditionally, data from each voxel, or from a group of voxels, are analysed with analysis of variance.
+This is called the blood oxygen level dependant (BOLD) response \cite{sm:2012fMRI}.
+
+The fMRI scanner measures the activity in three-dimensional pixels called voxels, which are about 3 x 3 x 3 mm in size, and a brain scan is made somewhere between every 500ms and every few seconds.
+However, before analysing the data is it usually pre-processed to account for head movements by participants in the scanner, and to make between-subjects comparisons possible \cite{sl:2009rl}.
+
+Traditionally, data from each voxel, or from a group of voxels, are analysed with analysis of variance.
 Voxels are usually convoluted with a haemodynamic response function in an attempt to infer the spiking neural activity from the slowly building and dropping BOLD response \cite{he:2002na}.
@@
 These findings are perhaps not surprising, because everything we know about the brain indicates that it is best seen as a network with countless connections to other parts of the brain.
-Therefore, it makes sense to also analyse signals from the brain as a pattern of information coming from the entire brain, rather attempting to link specific brain areas to experimental conditions.\\
-\indent In fact, Machine learning algorithms have been employed to learn the patterns of activity associated with some experimental condition.
+Therefore, it makes sense to also analyse signals from the brain as a pattern of information coming from the entire brain, rather attempting to link specific brain areas to experimental conditions.
+
+In fact, Machine learning algorithms have been employed to learn the patterns of activity associated with some experimental condition.
 Subsequently, these learned patterns can be used on unseen subjects in order to predict experimental conditions from their brain activity patterns.
-When these predictions are accurate then it means that the identified patterns hold information about the brain activity associated with some experimental task.\\
-\indent These machine learning techniques have been successfully used for decoding the orientation of a stimulus from fMRI scans \cite{ka:2005decoding}, identifying different patterns of activity for animate versus inanimate objects \cite{kr:2008RSA}, and even reconstructing visual input based on information in the fMRI scans \cite{mi:2008re}.
+When these predictions are accurate then it means that the identified patterns hold information about the brain activity associated with some experimental task.
+
+These machine learning techniques have been successfully used for decoding the orientation of a stimulus from fMRI scans \cite{ka:2005decoding}, identifying different patterns of activity for animate versus inanimate objects \cite{kr:2008RSA}, and even reconstructing visual input based on information in the fMRI scans \cite{mi:2008re}.
 The reason that these machine learning techniques work so well is that patterns of voxel activity can be identified rather than a significant deviation from some baseline.
 Specifically, information can be present in the pattern of voxels that belong to a certain brain region, even when the average signal from that region is zero (i.e., on average it does not differ from the baseline condition) \cite{to:2012rev}.
-\\
-\indent  The general goal of these machine learning studies is to identify what experimental condition a subject is in at a certain time.
+The general goal of these machine learning studies is to identify what experimental condition a subject is in at a certain time.
 The state that a person is in during some trial (e.g., looking at a face) is referred to as emph{cognitive state}, because the cognitive processes that are associated with being in that state are of interest.
@@
 These data have then been used to successfully infer the cognitive states of looking at either sentences or pictures within single subjects \cite{mi:2003within}, and between subjects \cite{wa:2003betw}.
-\\
-\indent In this study we use the same StarPlus data as in the previously mentioned experiment, and use machine learning algorithms to decode the cognitive states of subjects when they are looking at pictures or when they are looking at sentences.
+
+In this study we use the same StarPlus data as in the previously mentioned experiment, and use machine learning algorithms to decode the cognitive states of subjects when they are looking at pictures or when they are looking at sentences.
 An interesting difference between many applications of machine learning, and this study is that most machine learning algorithms have the goal of achieving the maximum possible discrimination between conditions or classes.
@@
 To avoid this problem it is recommended to only use linear classifiers when decoding cognitive states from fMRI data.
-The classifiers that we compare in this paper are Gaussian Naive Bayesian (GNB) classifiers, Linear Support Vector Machine (LSVM), and a Multivariate Gaussian (MVG) classifier.\\
-\indent	The goal of this study is to compare the classification performance of the cognitive states associated with looking at a picture or looking at a sentence between subject.
+The classifiers that we compare in this paper are Gaussian Naive Bayesian (GNB) classifiers, Linear Support Vector Machine (LSVM), and a Multivariate Gaussian (MVG) classifier.
+
+The goal of this study is to compare the classification performance of the cognitive states associated with looking at a picture or looking at a sentence between subject.
 This objective introduces an important issue, namely, how the data from the fMRI scans can be compared between subjects.
@@
 Second, an anatomically defined talairach coordinate system can be used that maps the voxels of each subject onto an anatomically defined coordinate.
-The ROI averaging it is a rather crude method, and a lot of information of the individual voxels is lost, however, it also doubles as a dimensionality reduction method.
-By using the talairach coordinate system the data of each individual voxel can be used, however, to use this system the anatomical information for each subject has to be known and mapped onto the fMRI data in great detail.
-Because we lack the detailed data to precisely use the coordinate system we will use averaged ROI's in order to compare the classifiers.<insert citation in this paragraph> \\
+The ROI averaging it is a rather crude method, and a lot of information of the individual voxels is lost.
+However, because the ROI's are relatively easy to identify for different subjects they can be compared easily between subjects.
+By using the talairach coordinate system the data of each individual voxel can be used.
+However, because the error of doing this is usually in the order of a centimetre, while the size of the voxels are in the order of millimetres, the coordinates are difficult to generalise.
+Because we aim to do between-subjects comparisons the generalisation performance is important, thus we will use averaged ROI's in order to compare the classifiers \cite{mi:2004coord}.\\
 
@@
 
-\subsection{Preprocessing}
-
 \subsection{Feature selection}
@@
 To reducde the dimensionality of the input, two different methods are used.
-
 ROI mapping. 
-
-Principal Component Analysis (PCA).
-
 So how do we cope with this problem?
@@
 - ROIs, means, stddev, skew, kurtosis?
-
-\subsection{Classifiers}
 

### 2012-11-09 article/article.tex
removed placeholders

@@
 
-%Traditionally, it has been a challenge to relate between a simple cognitive state (such as `looking at a triangle`) and the corresponding activity of the brain \cite{Pone0001394}.
-This correspondence may even vary from subject to subject, making the challenge even harder.
-Since the introduction of fMRI, researchers have been able to acquire detailed measurements of brain activity, allowing for machine learning methods to be applied for the detection of patterns in fMRI data.
-
-- briefly build up to state of the art, mention methods we tried?
-- state of the art: probably multivariate pattern analysis! (\cite{Haxby2012852})
-
-In the rest of this paper, we will discuss a number of machine learning techniques and their ability to classify our test data.
+%In the rest of this paper, we will discuss a number of machine learning techniques and their ability to classify our test data.
 In section~\ref{sec:methods} we will discuss the setup of our experiments and the analysis performed on our data.
@@
 To reducde the dimensionality of the input, two different methods are used.
+
 ROI mapping.
+
+Principal Component Analysis (PCA).
+
 So how do we cope with this problem?

### 2012-11-09 article/article.tex
removed placeholders
George Clinton
Monster Truck

@@
 The ROI averaging it is a rather crude method, and a lot of information of the individual voxels is lost.
-However, because the ROI's are relatively easy to identify for different subjects they can be compared easily between subjects.
+However, because the ROI's are relatively easy to identify by an expert they can be compared easily between subjects.
 By using the talairach coordinate system the data of each individual voxel can be used.
@@
 
-\subsection{Feature selection}
+\subsection{Dimensionality Reduction}
 %Useful background info: http://www.brainvoyager.com/bvqx/doc/UsersGuide/WebHelp/Content/MVPATools/MVPA_Basic_Concepts.htm
@@
 ROI mapping.
+As described in the introduction, ROIs are anatomically defined brain areas classified as relevant by an expert in the field.
+To reduce the number of features, average voxel values within this regions are taken as input features.
+As other higher-order features  such as the standard deviation, kurtosis and skweness within ROIs did not increase predictability, these were not utilized.
 

### 2012-11-09 article/article.tex
Diving Bell

@@
 To reduce the number of features, average voxel values within this regions are taken as input features.
-As other higher-order features  such as the standard deviation, kurtosis and skweness within ROIs did not increase predictability, these were not utilized.
+As other higher-order features such as the standard deviation, kurtosis and skewness within ROIs did not increase predictability, these were not utilized.
 

### 2012-11-09 article/article.tex
dunno what changed

--- 2012-11-09 article/article.tex
@@
 This study explored what happened to a simple contrast between BOLD signal while watching faces and baseline activity, when increasing the number of subjects included in the analysis.
-They found that brain areas that extended well beyond the expected visual- and face recognition areas showed significant deviation from baseline.
+They found that, as the number of analysed subjects increased, an increasing number of brain areas that extended well beyond the expected visual- and face recognition areas showed significant deviation from baseline.
 Another study, using only three subjects, repeated a simple letter/number discrimination task and visual stimulation.

### 2012-11-09 article/article.tex
results

--- 2012-11-09 article/article.tex
@@
 Brain activity is inferred from differences in the ratio between oxygenated and deoxygenated blood; when brain activity occurs, more oxygenated blood will flow to a brain region, and this change is detected by the machine.
-This is called the blood oxygen level dependant (BOLD) response \cite{sm:2012fMRI}.
-
-The fMRI scanner measures the activity in three-dimensional pixels called voxels, which are about 3 x 3 x 3 mm in size, and a brain scan is made somewhere between every 500ms and every few seconds.
-However, before analysing the data is it usually pre-processed to account for head movements by participants in the scanner, and to make between-subjects comparisons possible \cite{sl:2009rl}.
-
-Traditionally, data from each voxel, or from a group of voxels, are analysed with analysis of variance.
+This is called the blood oxygen level dependant (BOLD) response \cite{sm:2012fMRI}.\\
+\indent The fMRI scanner measures the activity in three-dimensional pixels called voxels, which are about 3 x 3 x 3 mm in size, and a brain scan is made somewhere between every 500ms and every few seconds.
+However, before analysing the data is it usually pre-processed to account for head movements by participants in the scanner, and to make between-subjects comparisons possible \cite{sl:2009rl}.\\
+\indent Traditionally, data from each voxel, or from a group of voxels, are analysed with analysis of variance.
 Voxels are usually convoluted with a haemodynamic response function in an attempt to infer the spiking neural activity from the slowly building and dropping BOLD response \cite{he:2002na}.
@@
 This study explored what happened to a simple contrast between BOLD signal while watching faces and baseline activity, when increasing the number of subjects included in the analysis.
-They found that, as the number of analysed subjects increased, an increasing number of brain areas that extended well beyond the expected visual- and face recognition areas showed significant deviation from baseline.
+They found that brain areas that extended well beyond the expected visual- and face recognition areas showed significant deviation from baseline.
 Another study, using only three subjects, repeated a simple letter/number discrimination task and visual stimulation.
@@
 These findings are perhaps not surprising, because everything we know about the brain indicates that it is best seen as a network with countless connections to other parts of the brain.
-Therefore, it makes sense to also analyse signals from the brain as a pattern of information coming from the entire brain, rather attempting to link specific brain areas to experimental conditions.
-
-In fact, Machine learning algorithms have been employed to learn the patterns of activity associated with some experimental condition.
+Therefore, it makes sense to also analyse signals from the brain as a pattern of information coming from the entire brain, rather attempting to link specific brain areas to experimental conditions.\\
+\indent In fact, Machine learning algorithms have been employed to learn the patterns of activity associated with some experimental condition.
 Subsequently, these learned patterns can be used on unseen subjects in order to predict experimental conditions from their brain activity patterns.
-When these predictions are accurate then it means that the identified patterns hold information about the brain activity associated with some experimental task.
-
-These machine learning techniques have been successfully used for decoding the orientation of a stimulus from fMRI scans \cite{ka:2005decoding}, identifying different patterns of activity for animate versus inanimate objects \cite{kr:2008RSA}, and even reconstructing visual input based on information in the fMRI scans \cite{mi:2008re}.
+When these predictions are accurate then it means that the identified patterns hold information about the brain activity associated with some experimental task.\\
+\indent These machine learning techniques have been successfully used for decoding the orientation of a stimulus from fMRI scans \cite{ka:2005decoding}, identifying different patterns of activity for animate versus inanimate objects \cite{kr:2008RSA}, and even reconstructing visual input based on information in the fMRI scans \cite{mi:2008re}.
 The reason that these machine learning techniques work so well is that patterns of voxel activity can be identified rather than a significant deviation from some baseline.
 Specifically, information can be present in the pattern of voxels that belong to a certain brain region, even when the average signal from that region is zero (i.e., on average it does not differ from the baseline condition) \cite{to:2012rev}.
-The general goal of these machine learning studies is to identify what experimental condition a subject is in at a certain time.
+\\
+\indent  The general goal of these machine learning studies is to identify what experimental condition a subject is in at a certain time.
 The state that a person is in during some trial (e.g., looking at a face) is referred to as emph{cognitive state}, because the cognitive processes that are associated with being in that state are of interest.
@@
 These data have then been used to successfully infer the cognitive states of looking at either sentences or pictures within single subjects \cite{mi:2003within}, and between subjects \cite{wa:2003betw}.
-
-In this study we use the same StarPlus data as in the previously mentioned experiment, and use machine learning algorithms to decode the cognitive states of subjects when they are looking at pictures or when they are looking at sentences.
+\\
+\indent In this study we use the same StarPlus data as in the previously mentioned experiment, and use machine learning algorithms to decode the cognitive states of subjects when they are looking at pictures or when they are looking at sentences.
 An interesting difference between many applications of machine learning, and this study is that most machine learning algorithms have the goal of achieving the maximum possible discrimination between conditions or classes.
@@
 To avoid this problem it is recommended to only use linear classifiers when decoding cognitive states from fMRI data.
-The classifiers that we compare in this paper are Gaussian Naive Bayesian (GNB) classifiers, Linear Support Vector Machine (LSVM), and a Multivariate Gaussian (MVG) classifier.
-
-The goal of this study is to compare the classification performance of the cognitive states associated with looking at a picture or looking at a sentence between subject.
+The classifiers that we compare in this paper are Gaussian Naive Bayesian (GNB) classifiers, Linear Support Vector Machine (LSVM), and a Multivariate Gaussian (MVG) classifier.\\
+\indent	The goal of this study is to compare the classification performance of the cognitive states associated with looking at a picture or looking at a sentence between subject.
 This objective introduces an important issue, namely, how the data from the fMRI scans can be compared between subjects.
@@
 Second, an anatomically defined talairach coordinate system can be used that maps the voxels of each subject onto an anatomically defined coordinate.
-The ROI averaging it is a rather crude method, and a lot of information of the individual voxels is lost.
-However, because the ROI's are relatively easy to identify by an expert they can be compared easily between subjects.
-By using the talairach coordinate system the data of each individual voxel can be used.
-However, because the error of doing this is usually in the order of a centimetre, while the size of the voxels are in the order of millimetres, the coordinates are difficult to generalise.
-Because we aim to do between-subjects comparisons the generalisation performance is important, thus we will use averaged ROI's in order to compare the classifiers \cite{mi:2004coord}.\\
+The ROI averaging it is a rather crude method, and a lot of information of the individual voxels is lost, however, it also doubles as a dimensionality reduction method.
+By using the talairach coordinate system the data of each individual voxel can be used, however, to use this system the anatomical information for each subject has to be known and mapped onto the fMRI data in great detail.
+Because we lack the detailed data to precisely use the coordinate system we will use averaged ROI's in order to compare the classifiers.<insert citation in this paragraph> \\
 
-%In the rest of this paper, we will discuss a number of machine learning techniques and their ability to classify our test data.
+%Traditionally, it has been a challenge to relate between a simple cognitive state (such as `looking at a triangle`) and the corresponding activity of the brain \cite{Pone0001394}.
+This correspondence may even vary from subject to subject, making the challenge even harder.
+Since the introduction of fMRI, researchers have been able to acquire detailed measurements of brain activity, allowing for machine learning methods to be applied for the detection of patterns in fMRI data.
+
+- briefly build up to state of the art, mention methods we tried?
+- state of the art: probably multivariate pattern analysis! (\cite{Haxby2012852})
+
+In the rest of this paper, we will discuss a number of machine learning techniques and their ability to classify our test data.
 In section~\ref{sec:methods} we will discuss the setup of our experiments and the analysis performed on our data.
@@
 \label{sec:results}
-- raw results per algorithm
+
+As a comparison measure, we used the performance of the classifiers for predicting the cognitive state that the subject was in.
+Table~\ref{tab:results} illustrates the overall performance of the three classifiers.
+Rows indicate various feature selection methods, where \emph{M} stands for \emph{Mean}, \emph{Norm} for \emph{Normalisation}, and \emph{SD} for \emph{Standard Deviation}.
+For instance, \emph{Norm-M-SD ROIs} denotes that normalisation was used in the voxels belonging to the selected ROIs over the mean and standard deviation.
+
+\begin{table*}[htpb]
+\centering
+{\small
+\begin{tabular}{c|c|c|c|c}
+& \textbf{Naive Bayes} & \textbf{Linear SVM} & \textbf{Naive Bayes*} & \textbf{Multivariate*} \\ \hline
+PCA & 61.46 &  - & - & - \\
+M ROIs & 70.21 & 72.5 & 74.79 & 76.67 \\
+Norm-M ROIs & 71.67 & 71.25 & 75.83 & 77.08 \\
+Norm-M-SD ROIs & 65.63 & 69.79 & 69.38 & 65.0 \\
+Norm/Vox-M ROIs & 73.13 & 76.67 & 76.88 & 79.58 \\
+\end{tabular}}
+\caption{Performance results (in \%) of the three classifiers on different methods for feature selection and structure.}
+\label{tab:results}
+\end{table*}
+
+Columns depict the used classifiers---Naive Bayes, Linear SVM, and Multivariate.
+For the classifiers depicted with *, we used different structure for the data that we fit to them, as described in Section~\ref{}.
+The table reveals the following general results:
+
+\begin{itemize}
+	\item Classifiers performance do not differ significantly between them. (explain)
+	\item Classifiers performance differ significantly from 50\%.
+\end{itemize}
+
+Moreover, it is observed that the Multivariate case has the best performance between all classifiers (why?).
+When PCA is used for feature extraction, the performance of Naive Bayes is lower than when other feature selection methods are used (why?).
+
+We can also distinguish the fact that when different structure is used for the fitting data, the performance increases, though not significantly.
+This is observed in the case of the Naive Bayes classifier.
+When different structure is used, its performance increases approximately 4\%.
+
+%mention confidence interval (how much percent sure we are that the actual accuracy of the classifier is between these regions)
 

### 2012-11-09 article/article.tex
results

@@
 \begin{itemize}
-	\item Classifiers performance do not differ significantly between them. (explain)
+	\item Classifiers performance do not differ significantly between them.
+Except from the case that PCA is used, the difference in performance between the classifiers for the rest of the cases is not more than 10\%.
 	\item Classifiers performance differ significantly from 50\%.
@@
 
-Moreover, it is observed that the Multivariate case has the best performance between all classifiers (why?).
-When PCA is used for feature extraction, the performance of Naive Bayes is lower than when other feature selection methods are used (why?).
+Moreover, it is observed that the Multivariate case has the best performance between.
+When PCA is used for feature extraction, the performance of Naive Bayes is lower than when other feature selection methods are used.
 
 We can also distinguish the fact that when different structure is used for the fitting data, the performance increases, though not significantly.
-This is observed in the case of the Naive Bayes classifier.
-When different structure is used, its performance increases approximately 4\%.
+For example, in the case of the Naive Bayes classifier when different structure is used, its performance increases approximately 4\%.
+
+Comparing classifiers when the same structure is used, it is observed that Linear SVM performs better than Naive Bayes.
+The same remains true when comparing the Multivariate case with Naive Bayes.
+The former one has better performance.
 

### 2012-11-09 article/article.tex
barry

--- 2012-11-09 article/article.tex
@@
 %% Abstract section.
-\abstract{Functional magnetic resonance imaging (fMRI) is a brain mapping technique that allows researchers to measure activity in a subject's brain.
+\abstract{
+Functional magnetic resonance imaging (fMRI) is a brain mapping technique that allows researchers to measure activity in a subject's brain.
 This technique is typically used to determine which areas of the brain are involved in performing a certain task.
@@
 ROI mapping.
+ROI mapping is a form of feature selection, wherein only a subset of the d-dimensional input space is used.
 As described in the introduction, ROIs are anatomically defined brain areas classified as relevant by an expert in the field.
 To reduce the number of features, average voxel values within this regions are taken as input features.
-As other higher-order features such as the standard deviation, kurtosis and skewness within ROIs did not increase predictability, these were not utilized.
+As other higher-order features such as the standard deviation, kurtosis and skewness within ROIs did not increase predictability, these were not included in the subsequent analysis.
 
 Principal Component Analysis (PCA).
+As feature extraction method, we applied PCA.
+In between subject classification, the input vectors of the PCA should ideally be the standardized Talairach coordinates.
+As this coordinate system appeared to be inaccurate for this dataset, we were forced to use the voxel indices.
+Because brain sizes and shapes differ per subject, it is uncertain whether the voxel indices map well onto each other.
+As a result, the accuracy of the PCA is reduced.
 
@@
 \label{sec:results}
-
-As a comparison measure, we used the performance of the classifiers for predicting the cognitive state that the subject was in.
-Table~\ref{tab:results} illustrates the overall performance of the three classifiers.
-Rows indicate various feature selection methods, where \emph{M} stands for \emph{Mean}, \emph{Norm} for \emph{Normalisation}, and \emph{SD} for \emph{Standard Deviation}.
-For instance, \emph{Norm-M-SD ROIs} denotes that normalisation was used in the voxels belonging to the selected ROIs over the mean and standard deviation.
-
-\begin{table*}[htpb]
-\centering
-{\small
-\begin{tabular}{c|c|c|c|c}
-& \textbf{Naive Bayes} & \textbf{Linear SVM} & \textbf{Naive Bayes*} & \textbf{Multivariate*} \\ \hline
-PCA & 61.46 &  - & - & - \\
-M ROIs & 70.21 & 72.5 & 74.79 & 76.67 \\
-Norm-M ROIs & 71.67 & 71.25 & 75.83 & 77.08 \\
-Norm-M-SD ROIs & 65.63 & 69.79 & 69.38 & 65.0 \\
-Norm/Vox-M ROIs & 73.13 & 76.67 & 76.88 & 79.58 \\
-\end{tabular}}
-\caption{Performance results (in \%) of the three classifiers on different methods for feature selection and structure.}
-\label{tab:results}
-\end{table*}
-
-Columns depict the used classifiers---Naive Bayes, Linear SVM, and Multivariate.
-For the classifiers depicted with *, we used different structure for the data that we fit to them, as described in Section~\ref{}.
-The table reveals the following general results:
-
-\begin{itemize}
-	\item Classifiers performance do not differ significantly between them. (explain)
-	\item Classifiers performance differ significantly from 50\%.
-\end{itemize}
-
-Moreover, it is observed that the Multivariate case has the best performance between all classifiers (why?).
-When PCA is used for feature extraction, the performance of Naive Bayes is lower than when other feature selection methods are used (why?).
-
-We can also distinguish the fact that when different structure is used for the fitting data, the performance increases, though not significantly.
-This is observed in the case of the Naive Bayes classifier.
-When different structure is used, its performance increases approximately 4\%.
-
-%mention confidence interval (how much percent sure we are that the actual accuracy of the classifier is between these regions)
+- raw results per algorithm
 

### 2012-11-09 article/article.tex
abstract

@@
 \abstract{
-Functional magnetic resonance imaging (fMRI) is a brain mapping technique that allows researchers to measure activity in a subject's brain.
+Functional magnetic resonance imaging (fMRI) is a brain mapping technique that allows researchers to infer activity in a subject's brain.
 This technique is typically used to determine which areas of the brain are involved in performing a certain task.
 Using this knowledge, we may also be able to determine what task was performed by looking at fMRI data.
-In this paper, we will discuss the use of machine learning algorithms for the classification of fMRI data.
-By using classifiers we will try to predict the cognitive state of subjects, based on their brain activity during that time.
+In this paper, we will compare machine learning algorithms in between-subjects classification performance of cognitive states using fMRI data.
+We found that averaging across ROI gave the best classification performance.
+All classifiers that we used performed above-chance, and we found that a Multivariate Gaussian classifier outperformed a Naive Bayesian classifier and a Support Vector Machine.
+Overall we can say that machine learning algorithms provide a feasible method for decoding cognitive states from fMRI data.
 } % end of abstract
@@
 To avoid this problem it is recommended to only use linear classifiers when decoding cognitive states from fMRI data.
-The classifiers that we compare in this paper are Gaussian Naive Bayesian (GNB) classifiers, Linear Support Vector Machine (LSVM), and a Multivariate Gaussian (MVG) classifier.
-
-The goal of this study is to compare the classification performance of the cognitive states associated with looking at a picture or looking at a sentence between subject.
+The classifiers that we compare in this paper are Gaussian Naive Bayesian (GNB) classifiers, Linear Support Vector Machine (LSVM), and a Multivariate Gaussian (MVG) classifier.\\
+\indent The goal of this study is to compare the classification performance of the cognitive states associated with looking at a picture or looking at a sentence between subject.
 This objective introduces an important issue, namely, how the data from the fMRI scans can be compared between subjects.
@@
 ROI mapping.
-ROI mapping is a form of feature selection, wherein only a subset of the d-dimensional input space is used.
 As described in the introduction, ROIs are anatomically defined brain areas classified as relevant by an expert in the field.
 To reduce the number of features, average voxel values within this regions are taken as input features.
-As other higher-order features such as the standard deviation, kurtosis and skewness within ROIs did not increase predictability, these were not included in the subsequent analysis.
+As other higher-order features such as the standard deviation, kurtosis and skewness within ROIs did not increase predictability, these were not utilized.
 
 Principal Component Analysis (PCA).
-As feature extraction method, we applied PCA.
-In between subject classification, the input vectors of the PCA should ideally be the standardized Talairach coordinates.
-As this coordinate system appeared to be inaccurate for this dataset, we were forced to use the voxel indices.
-Because brain sizes and shapes differ per subject, it is uncertain whether the voxel indices map well onto each other.
-As a result, the accuracy of the PCA is reduced.
 

### 2012-11-09 article/article.tex

--- 2012-11-09 article/article.tex
@@
 \label{sec:results}
-- raw results per algorithm
+
+As a comparison measure, we used the performance of the classifiers for predicting the cognitive state that the subject was in.
+Table~\ref{tab:results} illustrates the overall performance of the three classifiers.
+Rows indicate various feature selection methods, where \emph{M} stands for \emph{Mean}, \emph{Norm} for \emph{Normalisation}, and \emph{SD} for \emph{Standard Deviation}.
+For instance, \emph{Norm-M-SD ROIs} denotes that normalisation was used in the voxels belonging to the selected ROIs over the mean and standard deviation.
+
+\begin{table*}[htpb]
+\centering
+{\small
+\begin{tabular}{c|c|c|c|c}
+& \textbf{Naive Bayes} & \textbf{Linear SVM} & \textbf{Naive Bayes*} & \textbf{Multivariate*} \\ \hline
+PCA & 61.46 &  - & - & - \\
+M ROIs & 70.21 & 72.5 & 74.79 & 76.67 \\
+Norm-M ROIs & 71.67 & 71.25 & 75.83 & 77.08 \\
+Norm-M-SD ROIs & 65.63 & 69.79 & 69.38 & 65.0 \\
+Norm/Vox-M ROIs & 73.13 & 76.67 & 76.88 & 79.58 \\
+\end{tabular}}
+\caption{Performance results (in \%) of the three classifiers on different methods for feature selection and structure.}
+\label{tab:results}
+\end{table*}
+
+Columns depict the used classifiers---Naive Bayes, Linear SVM, and Multivariate.
+For the classifiers depicted with *, we used different structure for the data that we fit to them, as described in Section~\ref{}.
+The table reveals the following general results:
+
+\begin{itemize}
+	\item Classifiers performance do not differ significantly between them.
+Except from the case that PCA is used, the difference in performance between the classifiers for the rest of the cases is not more than 10\%.
+	\item Classifiers performance differ significantly from 50\%.
+\end{itemize}
+
+Moreover, it is observed that the Multivariate case has the best performance between.
+When PCA is used for feature extraction, the performance of Naive Bayes is lower than when other feature selection methods are used.
+
+We can also distinguish the fact that when different structure is used for the fitting data, the performance increases, though not significantly.
+For example, in the case of the Naive Bayes classifier when different structure is used, its performance increases approximately 4\%.
+
+Comparing classifiers when the same structure is used, it is observed that Linear SVM performs better than Naive Bayes.
+The same remains true when comparing the Multivariate case with Naive Bayes.
+The former one has better performance.
+
+%mention confidence interval (how much percent sure we are that the actual accuracy of the classifier is between these regions)
 

### 2012-11-09 article/article.tex
results
Peter Mozes Micky Palindrome Steve Bone-dry Watson

--- 2012-11-09 article/article.tex
@@
 As other higher-order features such as the standard deviation, kurtosis and skewness within ROIs did not increase predictability, these were not included in the subsequent analysis.
+ROIs require \textit{a priori} knowledge.
+When this knowledge is inaccurate, it results a inductive bias.
 
@@
 As feature extraction method, we applied PCA.
-In between subject classification, the input vectors of the PCA should ideally be the standardized Talairach coordinates.
+Ideally, the input vectors of the PCA should be the standardized Talairach coordinates, because we are conducting between-subject classification.
 As this coordinate system appeared to be inaccurate for this dataset, we were forced to use the voxel indices.
-Because brain sizes and shapes differ per subject, it is uncertain whether the voxel indices map well onto each other.
+The problem is that indices of different subjects may not map well onto each other, due to differences in brain size and shape.
 As a result, the accuracy of the PCA is reduced.

### 2012-11-09 article/article.tex

--- 2012-11-09 article/article.tex
@@
 
-As a comparison measure, we used the performance of the classifiers for predicting the cognitive state that the subject was in.
-Table~\ref{tab:results} illustrates the overall performance of the three classifiers.
-Rows indicate various feature selection methods, where \emph{M} stands for \emph{Mean}, \emph{Norm} for \emph{Normalisation}, and \emph{SD} for \emph{Standard Deviation}.
-For instance, \emph{Norm-M-SD ROIs} denotes that normalisation was used in the voxels belonging to the selected ROIs over the mean and standard deviation.
-
 \begin{table*}[htpb]
@@
 \end{table*}
+
+As a comparison measure, we used the performance of the classifiers for predicting the cognitive state that the subject was in.
+Table~\ref{tab:results} illustrates the overall performance of the three classifiers.
+Rows indicate various feature selection methods, where \emph{M} stands for \emph{Mean}, \emph{Norm} for \emph{Normalisation}, and \emph{SD} for \emph{Standard Deviation}.
+For instance, \emph{Norm-M-SD ROIs} denotes that normalisation was used in the voxels belonging to the selected ROIs over the mean and standard deviation.
 

### 2012-11-09 article/article.tex
bleh

@@
 ROIs require \textit{a priori} knowledge.
-When this knowledge is inaccurate, it results a inductive bias.
+When this knowledge is inaccurate, it results in a inductive bias.
 
@@
 As a result, the accuracy of the PCA is reduced.
-
-So how do we cope with this problem?
-On common solution is to reduce $|\vec{x}|$ by only using voxels belonging to a specific region of interest (ROI).
-However, this requires us to known which regions are actually of interest \textit{a priori}.
-- ROIs, means, stddev, skew, kurtosis?
 

### 2012-11-09 article/article.tex
results

--- 2012-11-09 article/article.tex
@@
 As other higher-order features such as the standard deviation, kurtosis and skewness within ROIs did not increase predictability, these were not included in the subsequent analysis.
-ROIs require \textit{a priori} knowledge.
-When this knowledge is inaccurate, it results a inductive bias.
 
@@
 As feature extraction method, we applied PCA.
-Ideally, the input vectors of the PCA should be the standardized Talairach coordinates, because we are conducting between-subject classification.
+In between subject classification, the input vectors of the PCA should ideally be the standardized Talairach coordinates.
 As this coordinate system appeared to be inaccurate for this dataset, we were forced to use the voxel indices.
-The problem is that indices of different subjects may not map well onto each other, due to differences in brain size and shape.
+Because brain sizes and shapes differ per subject, it is uncertain whether the voxel indices map well onto each other.
 As a result, the accuracy of the PCA is reduced.
@@
 \end{tabular}}
-\caption{Performance results (in \%) of the three classifiers on different methods for feature selection and structure.}
+\caption{Performance results (in \%) of the three classifiers on different methods for feature selection and structure.
+\emph{M} stands for \emph{Mean}, \emph{Norm} for \emph{Normalisation}, and \emph{SD} for \emph{Standard Deviation}}
 \label{tab:results}
@@
 Table~\ref{tab:results} illustrates the overall performance of the three classifiers.
-Rows indicate various feature selection methods, where \emph{M} stands for \emph{Mean}, \emph{Norm} for \emph{Normalisation}, and \emph{SD} for \emph{Standard Deviation}.
+Rows indicate various feature selection methods.
 For instance, \emph{Norm-M-SD ROIs} denotes that normalisation was used in the voxels belonging to the selected ROIs over the mean and standard deviation.
@@
 For the classifiers depicted with *, we used different structure for the data that we fit to them, as described in Section~\ref{}.
-The table reveals the following general results:
 
-\begin{itemize}
-	\item Classifiers performance do not differ significantly between them.
-Except from the case that PCA is used, the difference in performance between the classifiers for the rest of the cases is not more than 10\%.
-	\item Classifiers performance differ significantly from 50\%.
-\end{itemize}
-
-Moreover, it is observed that the Multivariate case has the best performance between.
 When PCA is used for feature extraction, the performance of Naive Bayes is lower than when other feature selection methods are used.
 
-We can also distinguish the fact that when different structure is used for the fitting data, the performance increases, though not significantly.
+We can also distinguish the fact that when different structure is used for the data, the performance increases, though not significantly.
 For example, in the case of the Naive Bayes classifier when different structure is used, its performance increases approximately 4\%.
@@
 
-%mention confidence interval (how much percent sure we are that the actual accuracy of the classifier is between these regions)
+%mention confidence interval (how much percent sure we are that the actual accuracy of the classifier is between these regions) exact binomial test
 

### 2012-11-09 article/article.tex

@@
 As other higher-order features such as the standard deviation, kurtosis and skewness within ROIs did not increase predictability, these were not included in the subsequent analysis.
+ROIs require \textit{a priori} knowledge.
+When this knowledge is inaccurate, it results in a inductive bias.
 
@@
 As feature extraction method, we applied PCA.
-In between subject classification, the input vectors of the PCA should ideally be the standardized Talairach coordinates.
+Ideally, the input vectors of the PCA should be the standardized Talairach coordinates, because we are conducting between-subject classification.
 As this coordinate system appeared to be inaccurate for this dataset, we were forced to use the voxel indices.
-Because brain sizes and shapes differ per subject, it is uncertain whether the voxel indices map well onto each other.
+The problem is that indices of different subjects may not map well onto each other, due to differences in brain size and shape.
 As a result, the accuracy of the PCA is reduced.
-
-So how do we cope with this problem?
-On common solution is to reduce $|\vec{x}|$ by only using voxels belonging to a specific region of interest (ROI).
-However, this requires us to known which regions are actually of interest \textit{a priori}.
-- ROIs, means, stddev, skew, kurtosis?
 

### 2012-11-09 article/article.tex
pe
Added talairach reference, read-through of abstract and chapter 1 for readability and grammar

@@
 Using this knowledge, we may also be able to determine what task was performed by looking at fMRI data.
-In this paper, we will compare machine learning algorithms in between-subjects classification performance of cognitive states using fMRI data.
-We found that averaging across ROI gave the best classification performance.
+In this paper, we present a comparison of the performance of machine learning algorithms in between-subjects classification of cognitive states, using fMRI data.
+We found that averaging across ROIs gave the best classification performance.
 All classifiers that we used performed above-chance, and we found that a Multivariate Gaussian classifier outperformed a Naive Bayesian classifier and a Support Vector Machine.
@@
 Of these techniques, functional magnetic resonance imaging (fMRI) has become more and more popular because it has a very good spatial resolution (contrary to MEG/EEG), and it is non-invasive (contrary to PET).
-Brain activity is inferred from differences in the ratio between oxygenated and deoxygenated blood; when brain activity occurs, more oxygenated blood will flow to a brain region, and this change is detected by the machine.
-This is called the blood oxygen level dependant (BOLD) response \cite{sm:2012fMRI}.
+Brain activity is inferred from differences in the ratio between oxygenated and deoxygenated blood; when brain activity occurs, more oxygenated blood will flow to a brain region, and this change is detected by a machine.
+This is called the blood oxygen level dependent (BOLD) response \cite{sm:2012fMRI}.
 
 The fMRI scanner measures the activity in three-dimensional pixels called voxels, which are about 3 x 3 x 3 mm in size, and a brain scan is made somewhere between every 500ms and every few seconds.
-However, before analysing the data is it usually pre-processed to account for head movements by participants in the scanner, and to make between-subjects comparisons possible \cite{sl:2009rl}.
+Before analysing the acquired data, it is usually pre-processed to account for head movements by participants in the scanner and to make between-subjects comparisons possible \cite{sl:2009rl}.
 
@@
 Experimental conditions are the independent variables in this setup, and a general linear model is fitted to each voxel, or group of voxels.
-By using this technique voxels or brain areas can be identified that differ in average BOLD response across conditions.\\
+By using this technique, voxels or brain areas can be identified that differ in their average BOLD response across conditions.\\
 \indent Recent studies indicate that this approach may have some serious limitations.
@@
 These findings are perhaps not surprising, because everything we know about the brain indicates that it is best seen as a network with countless connections to other parts of the brain.
-Therefore, it makes sense to also analyse signals from the brain as a pattern of information coming from the entire brain, rather attempting to link specific brain areas to experimental conditions.
+Therefore, it makes sense to also analyse signals from the brain as a pattern of information coming from the entire brain, rather than attempting to link specific brain areas to experimental conditions.
 
@@
 Subsequently, these learned patterns can be used on unseen subjects in order to predict experimental conditions from their brain activity patterns.
-When these predictions are accurate then it means that the identified patterns hold information about the brain activity associated with some experimental task.
+If these predictions are accurate then it means that the identified patterns hold information about the brain activity associated with some experimental task.
 
-These machine learning techniques have been successfully used for decoding the orientation of a stimulus from fMRI scans \cite{ka:2005decoding}, identifying different patterns of activity for animate versus inanimate objects \cite{kr:2008RSA}, and even reconstructing visual input based on information in the fMRI scans \cite{mi:2008re}.
+Machine learning techniques have been successfully used for decoding the orientation of a stimulus from fMRI scans \cite{ka:2005decoding}, identifying different patterns of activity for animate versus inanimate objects \cite{kr:2008RSA}, and even reconstructing visual input based on information in the fMRI scans \cite{mi:2008re}.
 The reason that these machine learning techniques work so well is that patterns of voxel activity can be identified rather than a significant deviation from some baseline.
@@
 The general goal of these machine learning studies is to identify what experimental condition a subject is in at a certain time.
-The state that a person is in during some trial (e.g., looking at a face) is referred to as emph{cognitive state}, because the cognitive processes that are associated with being in that state are of interest.
+The state that a person is in during some trial (e.g., looking at a face) is referred to as \emph{cognitive state}, because the cognitive processes that are associated with being in that state are of interest.
 Specifically, the aforementioned studies try to infer the cognitive state of a subject from the brain activity that is evoked in a subject by performing that trial.
-For example, in an experiment where subjects look at a set of pictures (e.g., a star and a plus), they would be in a certain cognitive state that is the result of looking at these pictures.
-One could then attempt to use machine learning on the fMRI signal that is evoked by the the brain activity associated with these cognitive states, and use this to learn for example what patterns of activity in the brain is associated with that cognitive state.
-This \emph{StarPlus} experiment has been performed, where in addition to viewing pictures, subjects also read sentences that contained some true or false statement about the pictures.
+For example, in an experiment where subjects look at a set of pictures (e.g.
+a star and a plus), they would be in a certain cognitive state that is the result of looking at these pictures.
+One could then attempt to use machine learning on the fMRI signal that is evoked by the the brain activity associated with these cognitive states, and use this to learn what patterns of activity in the brain are associated with that cognitive state.
+This \emph{StarPlus} experiment has been performed by researches where, in addition to viewing pictures, subjects also were tasked to read sentences containing some true or false statement about the pictures.
 These data have then been used to successfully infer the cognitive states of looking at either sentences or pictures within single subjects \cite{mi:2003within}, and between subjects \cite{wa:2003betw}.
 
-In this study we use the same StarPlus data as in the previously mentioned experiment, and use machine learning algorithms to decode the cognitive states of subjects when they are looking at pictures or when they are looking at sentences.
+In this study we use the same StarPlus data as in the aforementioned experiment, and use machine learning algorithms to decode the cognitive states of subjects when they are looking at pictures or when they are looking at sentences.
 An interesting difference between many applications of machine learning, and this study is that most machine learning algorithms have the goal of achieving the maximum possible discrimination between conditions or classes.
@@
 The classifiers that we compare in this paper are Gaussian Naive Bayesian (GNB) classifiers, Linear Support Vector Machine (LSVM), and a Multivariate Gaussian (MVG) classifier.\\
-\indent The goal of this study is to compare the classification performance of the cognitive states associated with looking at a picture or looking at a sentence between subject.
+\indent The goal of this study is to compare the classification performance of the cognitive states associated with looking at a picture or looking at a sentence between subjects.
 This objective introduces an important issue, namely, how the data from the fMRI scans can be compared between subjects.
-There are two mains ways this could be achieved.
-First, groups of voxels that are anatomically defined as belonging to a certain brain region, a region of interest (ROI), can be averaged into a single \emph{supervoxel}, which can be compared between subjects.
+There are two main ways this could be achieved.
+First, groups of voxels that are anatomically defined as belonging to a certain brain region, a so-called region of interest (ROI), can be averaged into a single \emph{supervoxel}, which can be compared between subjects.
 Second, an anatomically defined talairach coordinate system can be used that maps the voxels of each subject onto an anatomically defined coordinate.
@@
 However, because the ROI's are relatively easy to identify by an expert they can be compared easily between subjects.
-By using the talairach coordinate system the data of each individual voxel can be used.
+By using the Talairach coordinate system the data of each individual voxel can be used~\cite{talairach1988co}.
 However, because the error of doing this is usually in the order of a centimetre, while the size of the voxels are in the order of millimetres, the coordinates are difficult to generalise.
-Because we aim to do between-subjects comparisons the generalisation performance is important, thus we will use averaged ROI's in order to compare the classifiers \cite{mi:2004coord}.\\
+Because we aim to do between-subjects comparisons the generalisation performance is important, thus we will use averaged ROIs in order to compare the classifiers \cite{mi:2004coord}.\\
 

### 2012-11-09 article/article.tex
More references, initial Application section

@@
 
-In fact, Machine learning algorithms have been employed to learn the patterns of activity associated with some experimental condition.
+In fact, Machine learning algorithms have been employed in studies to learn the patterns of activity associated with some experimental condition.
 Subsequently, these learned patterns can be used on unseen subjects in order to predict experimental conditions from their brain activity patterns.
-If these predictions are accurate then it means that the identified patterns hold information about the brain activity associated with some experimental task.
+If these predictions are accurate then it means that the identified patterns hold information about the brain activity associated with some experimental task.\\
 
-Machine learning techniques have been successfully used for decoding the orientation of a stimulus from fMRI scans \cite{ka:2005decoding}, identifying different patterns of activity for animate versus inanimate objects \cite{kr:2008RSA}, and even reconstructing visual input based on information in the fMRI scans \cite{mi:2008re}.
-The reason that these machine learning techniques work so well is that patterns of voxel activity can be identified rather than a significant deviation from some baseline.
-Specifically, information can be present in the pattern of voxels that belong to a certain brain region, even when the average signal from that region is zero (i.e., on average it does not differ from the baseline condition) \cite{to:2012rev}.
 The general goal of these machine learning studies is to identify what experimental condition a subject is in at a certain time.
@@
 This \emph{StarPlus} experiment has been performed by researches where, in addition to viewing pictures, subjects also were tasked to read sentences containing some true or false statement about the pictures.
-These data have then been used to successfully infer the cognitive states of looking at either sentences or pictures within single subjects \cite{mi:2003within}, and between subjects \cite{wa:2003betw}.
+These data have then been used to successfully infer the cognitive states of looking at either sentences or pictures within single subjects \cite{mi:2003within}, and between subjects \cite{wa:2003betw}.\\
 
-In this study we use the same StarPlus data as in the aforementioned experiment, and use machine learning algorithms to decode the cognitive states of subjects when they are looking at pictures or when they are looking at sentences.
+In this study we use the same StarPlus data as in the aforementioned experiment, and use machine learning algorithms to decode the cognitive states of subjects while looking at pictures or while looking at sentences.
 An interesting difference between many applications of machine learning, and this study is that most machine learning algorithms have the goal of achieving the maximum possible discrimination between conditions or classes.
@@
 However, because the error of doing this is usually in the order of a centimetre, while the size of the voxels are in the order of millimetres, the coordinates are difficult to generalise.
-Because we aim to do between-subjects comparisons the generalisation performance is important, thus we will use averaged ROIs in order to compare the classifiers \cite{mi:2004coord}.\\
+Because we aim to do between-subjects comparisons the generalisation performance is important, thus we will use averaged ROIs in order to compare the classifiers \cite{mi:2004coord}.
 
@@
 \label{sec:application}
+Thus far, machine learning techniques have been successfully applied for decoding the orientation of a stimulus from fMRI scans \cite{ka:2005decoding}, identifying different patterns of activity for animate versus inanimate objects \cite{kr:2008RSA}, and even reconstructing visual input based on information in the fMRI scans \cite{mi:2008re}.
+The reason that these machine learning techniques work so well is that patterns of voxel activity can be identified rather than a significant deviation from some baseline.
+Specifically, information can be present in the pattern of voxels that belong to a certain brain region, even when the average signal from that region is zero (i.e., on average it does not differ from the baseline condition) \cite{to:2012rev}.
 
-What can we do with it now?
+The studies mentioned before all depend on the question of whether fMRI data carry class
+information (pattern discrimination) and exploiting this information to perform classification.
+However, researchers have indicated at least two other basic questions to which classifiers can be applied, being ``where or when
+is class information represented?'' (pattern localization) and ``how is
+class information encoded'' or ``how does its encoding relate to known
+relationships between stimuli'' (pattern characterization)~\cite{pereira2009machine}.
+While the main focus of this paper is on the application of machine learning techniques in pattern discrimination, we will briefly discuss these two other applications.
 
-- classification problem
-- prediction
-- cognitive state
 \section{Methods}

### 2012-11-09 article/article.tex
Added BR figure, some more application

@@
 
+\begin{figure}
+	\centering
+	\includegraphics[width=60mm]{figures/bold_response.png}
+  	\caption{Sample output of the BOLD response function based on a short, spiked stimulus}
+  	\label{fig:bold_response}
+\end{figure}
+
 Traditionally, data from each voxel, or from a group of voxels, are analysed with analysis of variance.
 Voxels are usually convoluted with a haemodynamic response function in an attempt to infer the spiking neural activity from the slowly building and dropping BOLD response \cite{he:2002na}.
-Experimental conditions are the independent variables in this setup, and a general linear model is fitted to each voxel, or group of voxels.
+Figure~\ref{fig:bold_response} depicts such a response function.
+In this setup, experimental conditions are the independent variables and a general linear model is fitted to each voxel, or group of voxels.
 By using this technique, voxels or brain areas can be identified that differ in their average BOLD response across conditions.\\

### 2012-11-09 article/article.tex
sorry for pix

@@
 
-\begin{figure}
-	\centering
-	\includegraphics[width=60mm]{figures/bold_response.png}
-  	\caption{Sample output of the BOLD response function based on a short, spiked stimulus}
-  	\label{fig:bold_response}
-\end{figure}
-
 Traditionally, data from each voxel, or from a group of voxels, are analysed with analysis of variance.
@@
 By using this technique, voxels or brain areas can be identified that differ in their average BOLD response across conditions.\\
+
+\begin{figure}
+	\centering
+	\includegraphics[width=60mm]{figures/f1_bold_response}
+  	\caption{Above is the hypothetical neural response to some stimulus.
+Below is the BOLD response that is evoked by the short spike of neural activity.
+The neural activity shown above is attempted to be inferred from the BOLD response that is measured by the fMRI scanner.
+Figure is adopted from Heeger and Ress \cite{he:2002na}.}
+  	\label{fig:bold_response}
+\end{figure}
+
 \indent Recent studies indicate that this approach may have some serious limitations.

### 2012-11-09 article/article.tex
spelling

@@
 These findings are perhaps not surprising, because everything we know about the brain indicates that it is best seen as a network with countless connections to other parts of the brain.
-Therefore, it makes sense to also analyse signals from the brain as a pattern of information coming from the entire brain, rather than attempting to link specific brain areas to experimental conditions.
-
-In fact, Machine learning algorithms have been employed in studies to learn the patterns of activity associated with some experimental condition.
+Therefore, it makes sense to also analyse signals from the brain as a pattern of information coming from the entire brain, rather than attempting to link specific brain areas to experimental conditions.\\
+\indent In fact, Machine learning algorithms have been employed in studies to learn the patterns of activity associated with some experimental condition.
 Subsequently, these learned patterns can be used on unseen subjects in order to predict experimental conditions from their brain activity patterns.
 If these predictions are accurate then it means that the identified patterns hold information about the brain activity associated with some experimental task.\\
-
-The general goal of these machine learning studies is to identify what experimental condition a subject is in at a certain time.
+\indent The general goal of these machine learning studies is to identify what experimental condition a subject is in at a certain time.
 The state that a person is in during some trial (e.g., looking at a face) is referred to as \emph{cognitive state}, because the cognitive processes that are associated with being in that state are of interest.
@@
 These data have then been used to successfully infer the cognitive states of looking at either sentences or pictures within single subjects \cite{mi:2003within}, and between subjects \cite{wa:2003betw}.\\
-
-In this study we use the same StarPlus data as in the aforementioned experiment, and use machine learning algorithms to decode the cognitive states of subjects while looking at pictures or while looking at sentences.
+\indent In this study we use the same StarPlus data as in the aforementioned experiment, and use machine learning algorithms to decode the cognitive states of subjects while looking at pictures or while looking at sentences.
 An interesting difference between many applications of machine learning, and this study is that most machine learning algorithms have the goal of achieving the maximum possible discrimination between conditions or classes.

### 2012-11-09 article/article.tex
locations

@@
 Traditionally, data from each voxel, or from a group of voxels, are analysed with analysis of variance.
-Voxels are usually convoluted with a haemodynamic response function in an attempt to infer the spiking neural activity from the slowly building and dropping BOLD response \cite{he:2002na}.
-Figure~\ref{fig:bold_response} depicts such a response function.
+Voxels are usually convoluted with a haemodynamic response function in an attempt to infer the spiking neural activity from the slowly building and dropping BOLD response \cite{he:2002na} (see Figure~\ref{fig:bold_response}).
 In this setup, experimental conditions are the independent variables and a general linear model is fitted to each voxel, or group of voxels.
 By using this technique, voxels or brain areas can be identified that differ in their average BOLD response across conditions.\\
-
 \begin{figure}
@@
 \end{figure}
-
 \indent Recent studies indicate that this approach may have some serious limitations.

### 2012-11-09 article/article.tex
hi

--- 2012-11-09 article/article.tex
@@
 Using this knowledge, we may also be able to determine what task was performed by looking at fMRI data.
-In this paper, we present a comparison of the performance of machine learning algorithms in between-subjects classification of cognitive states, using fMRI data.
-We found that averaging across ROIs gave the best classification performance.
+In this paper, we will compare machine learning algorithms in between-subjects classification performance of cognitive states using fMRI data.
+We found that averaging across ROI gave the best classification performance.
 All classifiers that we used performed above-chance, and we found that a Multivariate Gaussian classifier outperformed a Naive Bayesian classifier and a Support Vector Machine.
@@
 Of these techniques, functional magnetic resonance imaging (fMRI) has become more and more popular because it has a very good spatial resolution (contrary to MEG/EEG), and it is non-invasive (contrary to PET).
-Brain activity is inferred from differences in the ratio between oxygenated and deoxygenated blood; when brain activity occurs, more oxygenated blood will flow to a brain region, and this change is detected by a machine.
-This is called the blood oxygen level dependent (BOLD) response \cite{sm:2012fMRI}.
+Brain activity is inferred from differences in the ratio between oxygenated and deoxygenated blood; when brain activity occurs, more oxygenated blood will flow to a brain region, and this change is detected by the machine.
+This is called the blood oxygen level dependant (BOLD) response \cite{sm:2012fMRI}.
 
 The fMRI scanner measures the activity in three-dimensional pixels called voxels, which are about 3 x 3 x 3 mm in size, and a brain scan is made somewhere between every 500ms and every few seconds.
-Before analysing the acquired data, it is usually pre-processed to account for head movements by participants in the scanner and to make between-subjects comparisons possible \cite{sl:2009rl}.
+However, before analysing the data is it usually pre-processed to account for head movements by participants in the scanner, and to make between-subjects comparisons possible \cite{sl:2009rl}.
 
@@
 Voxels are usually convoluted with a haemodynamic response function in an attempt to infer the spiking neural activity from the slowly building and dropping BOLD response \cite{he:2002na}.
-Figure~\ref{fig:bold_response} depicts such a response function.
-In this setup, experimental conditions are the independent variables and a general linear model is fitted to each voxel, or group of voxels.
-By using this technique, voxels or brain areas can be identified that differ in their average BOLD response across conditions.\\
-
-\begin{figure}
-	\centering
-	\includegraphics[width=60mm]{figures/f1_bold_response}
-  	\caption{Above is the hypothetical neural response to some stimulus.
-Below is the BOLD response that is evoked by the short spike of neural activity.
-The neural activity shown above is attempted to be inferred from the BOLD response that is measured by the fMRI scanner.
-Figure is adopted from Heeger and Ress \cite{he:2002na}.}
-  	\label{fig:bold_response}
-\end{figure}
-
+Experimental conditions are the independent variables in this setup, and a general linear model is fitted to each voxel, or group of voxels.
+By using this technique voxels or brain areas can be identified that differ in average BOLD response across conditions.\\
 \indent Recent studies indicate that this approach may have some serious limitations.
@@
 These findings are perhaps not surprising, because everything we know about the brain indicates that it is best seen as a network with countless connections to other parts of the brain.
-Therefore, it makes sense to also analyse signals from the brain as a pattern of information coming from the entire brain, rather than attempting to link specific brain areas to experimental conditions.
+Therefore, it makes sense to also analyse signals from the brain as a pattern of information coming from the entire brain, rather attempting to link specific brain areas to experimental conditions.
 
-In fact, Machine learning algorithms have been employed in studies to learn the patterns of activity associated with some experimental condition.
+In fact, Machine learning algorithms have been employed to learn the patterns of activity associated with some experimental condition.
 Subsequently, these learned patterns can be used on unseen subjects in order to predict experimental conditions from their brain activity patterns.
-If these predictions are accurate then it means that the identified patterns hold information about the brain activity associated with some experimental task.\\
+When these predictions are accurate then it means that the identified patterns hold information about the brain activity associated with some experimental task.
 
+These machine learning techniques have been successfully used for decoding the orientation of a stimulus from fMRI scans \cite{ka:2005decoding}, identifying different patterns of activity for animate versus inanimate objects \cite{kr:2008RSA}, and even reconstructing visual input based on information in the fMRI scans \cite{mi:2008re}.
+The reason that these machine learning techniques work so well is that patterns of voxel activity can be identified rather than a significant deviation from some baseline.
+Specifically, information can be present in the pattern of voxels that belong to a certain brain region, even when the average signal from that region is zero (i.e., on average it does not differ from the baseline condition) \cite{to:2012rev}.
 The general goal of these machine learning studies is to identify what experimental condition a subject is in at a certain time.
-The state that a person is in during some trial (e.g., looking at a face) is referred to as \emph{cognitive state}, because the cognitive processes that are associated with being in that state are of interest.
+The state that a person is in during some trial (e.g., looking at a face) is referred to as emph{cognitive state}, because the cognitive processes that are associated with being in that state are of interest.
 Specifically, the aforementioned studies try to infer the cognitive state of a subject from the brain activity that is evoked in a subject by performing that trial.
-For example, in an experiment where subjects look at a set of pictures (e.g.
-a star and a plus), they would be in a certain cognitive state that is the result of looking at these pictures.
-One could then attempt to use machine learning on the fMRI signal that is evoked by the the brain activity associated with these cognitive states, and use this to learn what patterns of activity in the brain are associated with that cognitive state.
-This \emph{StarPlus} experiment has been performed by researches where, in addition to viewing pictures, subjects also were tasked to read sentences containing some true or false statement about the pictures.
-These data have then been used to successfully infer the cognitive states of looking at either sentences or pictures within single subjects \cite{mi:2003within}, and between subjects \cite{wa:2003betw}.\\
+For example, in an experiment where subjects look at a set of pictures (e.g., a star and a plus), they would be in a certain cognitive state that is the result of looking at these pictures.
+One could then attempt to use machine learning on the fMRI signal that is evoked by the the brain activity associated with these cognitive states, and use this to learn for example what patterns of activity in the brain is associated with that cognitive state.
+This \emph{StarPlus} experiment has been performed, where in addition to viewing pictures, subjects also read sentences that contained some true or false statement about the pictures.
+These data have then been used to successfully infer the cognitive states of looking at either sentences or pictures within single subjects \cite{mi:2003within}, and between subjects \cite{wa:2003betw}.
 
-In this study we use the same StarPlus data as in the aforementioned experiment, and use machine learning algorithms to decode the cognitive states of subjects while looking at pictures or while looking at sentences.
+In this study we use the same StarPlus data as in the previously mentioned experiment, and use machine learning algorithms to decode the cognitive states of subjects when they are looking at pictures or when they are looking at sentences.
 An interesting difference between many applications of machine learning, and this study is that most machine learning algorithms have the goal of achieving the maximum possible discrimination between conditions or classes.
@@
 The classifiers that we compare in this paper are Gaussian Naive Bayesian (GNB) classifiers, Linear Support Vector Machine (LSVM), and a Multivariate Gaussian (MVG) classifier.\\
-\indent The goal of this study is to compare the classification performance of the cognitive states associated with looking at a picture or looking at a sentence between subjects.
+\indent The goal of this study is to compare the classification performance of the cognitive states associated with looking at a picture or looking at a sentence between subject.
 This objective introduces an important issue, namely, how the data from the fMRI scans can be compared between subjects.
-There are two main ways this could be achieved.
-First, groups of voxels that are anatomically defined as belonging to a certain brain region, a so-called region of interest (ROI), can be averaged into a single \emph{supervoxel}, which can be compared between subjects.
+There are two mains ways this could be achieved.
+First, groups of voxels that are anatomically defined as belonging to a certain brain region, a region of interest (ROI), can be averaged into a single \emph{supervoxel}, which can be compared between subjects.
 Second, an anatomically defined talairach coordinate system can be used that maps the voxels of each subject onto an anatomically defined coordinate.
@@
 However, because the ROI's are relatively easy to identify by an expert they can be compared easily between subjects.
-By using the Talairach coordinate system the data of each individual voxel can be used~\cite{talairach1988co}.
+By using the talairach coordinate system the data of each individual voxel can be used.
 However, because the error of doing this is usually in the order of a centimetre, while the size of the voxels are in the order of millimetres, the coordinates are difficult to generalise.
-Because we aim to do between-subjects comparisons the generalisation performance is important, thus we will use averaged ROIs in order to compare the classifiers \cite{mi:2004coord}.
+Because we aim to do between-subjects comparisons the generalisation performance is important, thus we will use averaged ROI's in order to compare the classifiers \cite{mi:2004coord}.\\
 
@@
 \label{sec:application}
-Thus far, machine learning techniques have been successfully applied for decoding the orientation of a stimulus from fMRI scans \cite{ka:2005decoding}, identifying different patterns of activity for animate versus inanimate objects \cite{kr:2008RSA}, and even reconstructing visual input based on information in the fMRI scans \cite{mi:2008re}.
-The reason that these machine learning techniques work so well is that patterns of voxel activity can be identified rather than a significant deviation from some baseline.
-Specifically, information can be present in the pattern of voxels that belong to a certain brain region, even when the average signal from that region is zero (i.e., on average it does not differ from the baseline condition) \cite{to:2012rev}.
 
-The studies mentioned before all depend on the question of whether fMRI data carry class
-information (pattern discrimination) and exploiting this information to perform classification.
-However, researchers have indicated at least two other basic questions to which classifiers can be applied, being ``where or when
-is class information represented?'' (pattern localization) and ``how is
-class information encoded'' or ``how does its encoding relate to known
-relationships between stimuli'' (pattern characterization)~\cite{pereira2009machine}.
-While the main focus of this paper is on the application of machine learning techniques in pattern discrimination, we will briefly discuss these two other applications.
+What can we do with it now?
 
+- classification problem
+- prediction
+- cognitive state
 \section{Methods}
@@
 
+We compare the following classifiers:
+
+Gaussian Naive Bayes (GNB).\cite{bk:2004aply}
+
 - same for each algorithm:

### 2012-11-10 article/article.tex
locations
peace

--- 2012-11-09 article/article.tex
@@
 Below is the BOLD response that is evoked by the short spike of neural activity.
-The neural activity shown above is attempted to be inferred from the BOLD response that is measured by the fMRI scanner.
+The neural activity shown above is attempted to be inferred from the BOLD response below that is measured by the fMRI scanner.
 Figure is adopted from Heeger and Ress \cite{he:2002na}.}
@@
 One could then attempt to use machine learning on the fMRI signal that is evoked by the the brain activity associated with these cognitive states, and use this to learn what patterns of activity in the brain are associated with that cognitive state.
-This \emph{StarPlus} experiment has been performed by researches where, in addition to viewing pictures, subjects also were tasked to read sentences containing some true or false statement about the pictures.
+This \emph{StarPlus} experiment has been performed by researchers where, in addition to viewing pictures, subjects also were tasked to read sentences containing some true or false statement about the pictures.
 These data have then been used to successfully infer the cognitive states of looking at either sentences or pictures within single subjects \cite{mi:2003within}, and between subjects \cite{wa:2003betw}.\\
@@
 The ROI averaging it is a rather crude method, and a lot of information of the individual voxels is lost.
-However, because the ROI's are relatively easy to identify by an expert they can be compared easily between subjects.
+However, because the ROIs are relatively easy to identify by an expert they can be compared easily between subjects.
 By using the Talairach coordinate system the data of each individual voxel can be used~\cite{talairach1988co}.
@@
 
-To reducde the dimensionality of the input, two different methods are used.
+To reducde the dimensionality of the input, several different methods were used.
 
@@
 As described in the introduction, ROIs are anatomically defined brain areas classified as relevant by an expert in the field.
-To reduce the number of features, average voxel values within this regions are taken as input features.
+We used 7 of the 25 defined ROIs, labeled  'CALC' 'LIPL' 'LT' 'LTRIA' 'LOPER' 'LIPS' and  'LDLPFC'.
+These ROIs have previously been found to be most relevant \cite{wa:2003betw}.
+To further reduce the number of features, average voxel values within this regions are taken as new input features.
 As other higher-order features such as the standard deviation, kurtosis and skewness within ROIs did not increase predictability, these were not included in the subsequent analysis.
-ROIs require \textit{a priori} knowledge.
-When this knowledge is inaccurate, it results in a inductive bias.
+The disadvantage of using ROIs is that they require \textit{a priori} knowledge.
+When this knowledge is inaccurate, it results in an inductive bias.
 
@@
 \subsection{Classifiers}
+
+We compare the following classifiers:
+
+Gaussian Naive Bayes (GNB).\cite{bk:2004aply}
 

### 2012-11-10 article/article.tex
boe

@@
 
-In these trials, subjects were presented with a stimulus on a computer screen.
-This stimulus was either a picture or a sentence.
-After a period of time, the stimulus would be replaced with a stimulus of the opposite type (i.e.
-picture with sentence and vice versa) after which the subject would indicate by pressing a button whether the sentence correctly described the picture.
-At the end of each trial, approximately 54 fMRI scans have be taken with an interval of 500ms.
+In some of these trials (baseline trials), no stimulus was presented and participants were instructed to remain inert.
+In other trials, subjects were presented with a sequence of two stimuli on a computer screen.
+The first stimulus was either a picture or a sentence lasting 4 seconds, followed by a blank screen of 4 seconds.
+After the blank screen, the opposite type of stimulus was presented for 4 seconds, followed by a blank screen of 15 seconds, during which the subject would indicate by pressing a button whether the sentence correctly described the picture.
+During a trial, approximately 54 fMRI scans have be taken with an interval of 500ms.
+In total, 14 trials were baseline trials, while half of the remaining trials had a picture as first stimulus and the other half a sentence.
 The entire dataset consist of voxel data and metadata providing information about the dataset.
@@
 
-In total, we have data available for approximately 4800 voxels per scan, 54 scans per trial, 40 trials per subject, 6 subjects.
+In total, we have data available for approximately 4800 voxels per scan, 54 scans per trial, 40 trials per subject (excluding baseline trials) and 6 subjects.
 Voxels have been grouped into 25 regions of interest (see \textit{Feature selection}) that may be of use for analysis.
@@
 The problem is that indices of different subjects may not map well onto each other, due to differences in brain size and shape.
-As a result, the accuracy of the PCA is reduced.
+As a result, the accuracy of the PCA may be reduced.
 

### 2012-11-10 article/article.tex
boe

@@
 The problem is that indices of different subjects may not map well onto each other, due to differences in brain size and shape.
-As a result, the accuracy of the PCA may be reduced.
+As a result, the accuracy of the PCA may be greatly reduced.
+
+Temporal dimensionality reduction.
+As described in the introduction, the peak of the BOLD response is approximately 5 seconds after a stimulus is presented.
+This means that the first scans after a stimulus may not represent activity related to this stimulus and should not be used for classification.
+To examine which scans provide maximal class information, we plotted the average BOLD activity per scan for all trials of all subjects, given a trial type and ROI.
+As can be seen in Figure, scans 10 to 20 provide maximal information for the first stimulus, while scans 23 to 33 provide maximal information for the second stimulus.
+The other scans were excluded.
+As a comparison,
 

### 2012-11-10 article/article.tex
new car
mozes

@@
 To examine which scans provide maximal class information, we plotted the average BOLD activity per scan for all trials of all subjects, given a trial type and ROI.
-As can be seen in Figure, scans 10 to 20 provide maximal information for the first stimulus, while scans 23 to 33 provide maximal information for the second stimulus.
+As can be seen in Figure~\ref{fig:boldroi}, scans 10 to 20 provide maximal information for the first stimulus, while scans 23 to 33 provide maximal information for the second stimulus.
 The other scans were excluded.
-As a comparison,
-
+\begin{figure}
+	\centering
+	\includegraphics[width=90mm]{figures/scan_hemodynamic_BOLD_roi}
+  	\caption{Haemodynamic BOLD response for ROI "CALC". Average Bold activity for all subjects is plotted against the scan number. The coloured regions depict an interval of 1 standard deviation.}
+  	\label{fig:boldroi}
+\end{figure}
 \subsection{Classifiers}

### 2012-11-10 article/article.tex
mozes

@@
 ROI mapping.
-ROI mapping is a form of feature selection, wherein only a subset of the d-dimensional input space is used.
+ROI mapping is a form of feature selection, wherein only a subset of the N-dimensional input space is used.
 As described in the introduction, ROIs are anatomically defined brain areas classified as relevant by an expert in the field.
@@
 These ROIs have previously been found to be most relevant \cite{wa:2003betw}.
-To further reduce the number of features, average voxel values within this regions are taken as new input features.
+To further reduce the number of features, average voxel values within these regions are taken as new input features.
 As other higher-order features such as the standard deviation, kurtosis and skewness within ROIs did not increase predictability, these were not included in the subsequent analysis.

### 2012-11-10 article/article.tex
mozes

@@
 After the blank screen, the opposite type of stimulus was presented for 4 seconds, followed by a blank screen of 15 seconds, during which the subject would indicate by pressing a button whether the sentence correctly described the picture.
-During a trial, approximately 54 fMRI scans have be taken with an interval of 500ms.
+During each trial, approximately 54 fMRI scans have been taken with an interval of 500ms.
 In total, 14 trials were baseline trials, while half of the remaining trials had a picture as first stimulus and the other half a sentence.
@@
 	\includegraphics[width=90mm]{figures/scan_hemodynamic_BOLD_roi}
-  	\caption{Haemodynamic BOLD response for ROI "CALC". Average Bold activity for all subjects is plotted against the scan number. The coloured regions depict an interval of 1 standard deviation.}
+  	\caption{Haemodynamic BOLD response for ROI "CALC".
+The red line depicts a sentence-first trial (S-P) and the blue line depicts a picture-first trials (P-S).
+Average Bold activity for all subjects is plotted against the scan number.
+The coloured regions depict an interval of 1 standard deviation.}
   	\label{fig:boldroi}
@@
 
-Gaussian Naive Bayes (GNB).\cite{bk:2004aply}
+Gaussian Naive Bayes (GNB).\cite{bk:2004aplay}
+
+Linear Support Vector Machine.
+
+Multivariate Gaussian.
 

### 2012-11-10 article/article.tex
mozes
results

--- 2012-11-10 article/article.tex
@@
 \subsection{Dimensionality Reduction}
+\label{sec:dimensionalityReduction}
 %Useful background info: http://www.brainvoyager.com/bvqx/doc/UsersGuide/WebHelp/Content/MVPATools/MVPA_Basic_Concepts.htm
@@
 \caption{Performance results (in \%) of the three classifiers on different methods for feature selection and structure.
-\emph{M} stands for \emph{Mean}, \emph{Norm} for \emph{Normalisation}, and \emph{SD} for \emph{Standard Deviation}}
+\emph{M} stands for \emph{Mean}, \emph{Norm} for \emph{Normalisation}, \emph{SD} for \emph{Standard Deviation}, and \emph{Norm/Vox} for \emph{Normalisation per Voxel index}.}
 \label{tab:results}
@@
 
-As a comparison measure, we used the performance of the classifiers for predicting the cognitive state that the subject was in.
+As a comparison measure, we used the performance of the classifiers for predicting the cognitive state that a subject was in.
+The predictive performance of Naive Bayes, Linear SVM, and Multivariate was compared.
+We observed that different feature selection and extraction methods, as well as different structure on the data affects the predictive performance of the classifiers.
+
  Table~\ref{tab:results} illustrates the overall performance of the three classifiers.
 Rows indicate various feature selection methods.
-For instance, \emph{Norm-M-SD ROIs} denotes that normalisation was used in the voxels belonging to the selected ROIs over the mean and standard deviation.
-
-Columns depict the used classifiers---Naive Bayes, Linear SVM, and Multivariate.
+We applied PCA to the original data, while the rest methods were applied to the data belonging to the selected ROIs (cf.~Section~\ref{sec:dimensionalityReduction}).
+For instance, \emph{Norm/Vox-M-SD ROIs} indicates that we normalised over the voxel values of the same voxel index using the mean and standard deviation, while \emph{Norm-M-SD ROIs} indicates that we normalised over the voxel values of the same scan index using the mean and standard deviation.
+Columns depict the used classifiers.
 For the classifiers depicted with *, we used different structure for the data that we fit to them, as described in Section~\ref{}.
 
-When PCA is used for feature extraction, the performance of Naive Bayes is lower than when other feature selection methods are used.
+When PCA is used for feature extraction, the performance of Naive Bayes is significantly lower than when other feature selection methods are used.
+By using an exact binomial test we found out the confidence interval of this result
 
@@
 %%use following if all content of bibtex file should be shown
-\nocite{*}
+%\nocite{*}
 \bibliography{references}

### 2012-11-10 article/article.tex
results

@@
 
-When PCA is used for feature extraction, the performance of Naive Bayes is significantly lower than when other feature selection methods are used.
-By using an exact binomial test we found out the confidence interval of this result
+From Table~\ref{tab:results} we can distinguish the fact that when PCA is used for feature extraction, the performance of Naive Bayes is lower than when other feature selection methods are used.
+By using an exact binomial test we found out that the confidence interval of the predictive performance of Naive Bayes when PCA is used is $95\% CI: 56.9 - 65.8$.
 
-We can also distinguish the fact that when different structure is used for the data, the performance increases, though not significantly.
-For example, in the case of the Naive Bayes classifier when different structure is used, its performance increases approximately 4\%.
+We can also distinguish the fact that when different structure is used for the data, the performance increases,
 

### 2012-11-10 article/article.tex
mozes

--- 2012-11-10 article/article.tex
@@
 In total, 14 trials were baseline trials, while half of the remaining trials had a picture as first stimulus and the other half a sentence.
-The entire dataset consist of voxel data and metadata providing information about the dataset.
+The entire dataset consists of voxel data and meta-data providing information about the dataset.
 Included in this information is not only the trial and scan each voxel belongs to, but also the type of the trial (picture first or sentence first).
@@
 This problem is formally known as the \textit{Hughes effect}\cite{Hughes1054102} which states that the predictive power of a machine learning algorithm decreases as the dimensionality increases with a fixed number of training examples.
-
 To reducde the dimensionality of the input, several different methods were used.
@@
 \end{figure}
-\subsection{Classifiers}
 
+\subsection{Training Method and Classifiers}
+For each classifier, a "leave one (subject) out" cross validation procedure was used to avoid overfitting.
+For N=6 iterations, N-1 subjects functioned as training set, while the other subject functioned as validation set.
+The classifier accuracy is the average accuracy over all interations.
 We compare the following classifiers:
 
-Gaussian Naive Bayes (GNB).\cite{bk:2004aplay}
+Gaussian Naive Bayes (GNB).
+This classifier builds a generative model underlying a class, by estimating the means and the standard deviations of the input features from the training set\cite{bk:2004aplay}.
+It assumes a Gaussian distribution and strongly independent input features.
+Bayes rule is used to estimate the probability of a class (picture or sentence), given the input features of the validation set.
 

### 2012-11-10 article/article.tex
mozes

@@
 The disadvantage of using ROIs is that they require \textit{a priori} knowledge.
-When this knowledge is inaccurate, it results in an inductive bias.
+When this knowledge is inaccurate, classification accuracy may not be optimal.
 
@@
 Gaussian Naive Bayes (GNB).
-This classifier builds a generative model underlying a class, by estimating the means and the standard deviations of the input features from the training set\cite{bk:2004aplay}.
+This classifier builds a generative model underlying a class, by estimating the means and the standard deviations of the input features from the training set \cite{bk:2004aplay}.
 It assumes a Gaussian distribution and strongly independent input features.
 Bayes rule is used to estimate the probability of a class (picture or sentence), given the input features of the validation set.
+Two approaches were used to estimate the parameters of the model.
+In the first method, averages of the selected scans belonging to a stimulus were used to estimate the parameters, while in the second method, the parameters were estimated for each scan.
+In both methods, the resulting probability of a class was the sum of log-likelyhoods per scan and the class with maximum probability was selected.
+The advantage of the first method is that fewer parameters need to be estimated with greater amount of examples.
+It thus reduces the space and time complexity, and variance of the model.
+However, because it is likely that the assumptions of equal means and standard deviations for each scan do not hold, the second approach may decrease the inductive bias and enhance the accuracy.
 

### 2012-11-10 article/article.tex
mozes
results

--- 2012-11-10 article/article.tex
@@
 Columns depict the used classifiers.
-For the classifiers depicted with *, we used different structure for the data that we fit to them, as described in Section~\ref{}.
+For the classifiers depicted with *, we used different structure for the data, as described in Section~\ref{}.
+
+Furthermore, in order to find out how much percent sure we are that the actually accuracy of the classifiers is between some regions we performed an exact binomial test in each case, from which we obtained the confidence interval.
+The results were rounded to one decimal point.
+For the following comparison we are only looking at the case where \emph{Norm/Vox-M-SD ROIs} was used.
 
 From Table~\ref{tab:results} we can distinguish the fact that when PCA is used for feature extraction, the performance of Naive Bayes is lower than when other feature selection methods are used.
-By using an exact binomial test we found out that the confidence interval of the predictive performance of Naive Bayes when PCA is used is $95\% CI: 56.9 - 65.8$.
+The confidence interval of the predictive performance of Naive Bayes when PCA is used is 95\% $CI: 56.9 - 65.8$ which suggests that...
 
-We can also distinguish the fact that when different structure is used for the data, the performance increases,
+Comparing classifiers when the same data structure is used, we can distinguish the fact that Multivariate performs better than Naive Bayes.
+Also that Linear SVM performs better than Naive Bayes, though not significantly.
+The confidence interval of Naive Bayes is 95\% $CI: 68.9-77.0$ and of Linear SVM is 95\% $CI: 72.6-80.4$, which suggests that their performances do not differ significantly.
 
-Comparing classifiers when the same structure is used, it is observed that Linear SVM performs better than Naive Bayes.
-The same remains true when comparing the Multivariate case with Naive Bayes.
-The former one has better performance.
+Moreover, when different data structure is used the performance increases, though not significantly.
+This can be seen by comparing the performance of Naive Bayes when different data structure is used.
+In the first case it is 73.13\% while in the second case is 76.88\%.
+The confidence interval in the first case is 95\% $CI: 68.9 - 77.0$ and in the second case is 95\% $CI: 72.8 - 80.6$.
 
-%mention confidence interval (how much percent sure we are that the actual accuracy of the classifier is between these regions) exact binomial test
+Last but not least, we can distinguish the fact that Multivariate outperforms all the other classifiers, though we did not try a different data structure in the case of Linear SVM.
+However, we still do not expect that the performance of Linear SVM will be higher that this of Multivariate.
+The confidence interval of Multivariate is 95\% $CI: 75.7-83.1$.
 

### 2012-11-10 article/article.tex

@@
 \end{figure}
-\subsection{Classifiers}
 
-We compare the following classifiers:
+\subsection{Training Method and Classifiers}
+For each classifier, a "leave one (subject) out" cross validation procedure was used to avoid overfitting.
+For N=6 iterations, N-1 subjects functioned as training set, while the other subject functioned as validation set.
+The classifier accuracy is the average accuracy over all interations.
+We compared the following classifiers:
 
-Gaussian Naive Bayes (GNB).\cite{bk:2004aplay}
+Gaussian Naive Bayes (GNB).
+This classifier builds a generative model underlying a class, by estimating the means and the standard deviations of the input features from the training set \cite{bk:2004aplay}.
+It assumes a Gaussian distribution and strongly independent input features.
+Bayes rule is used to estimate the probability of a class (picture or sentence), given the input features of the validation set.
+Two approaches were used to estimate the parameters of the model.
+In the first method, averages of the selected scans belonging to a stimulus were used to estimate the parameters, while in the second method, the parameters were estimated for each scan.
+In both methods, the resulting probability of a class was the sum of log-likelyhoods per scan and the class with maximum probability was selected.
+The advantage of the first method is that fewer parameters need to be estimated with greater amount of examples.
+It thus reduces the space and time complexity, and variance of the model.
+However, because it is likely that the assumptions of equal means and standard deviations for each scan do not hold, the second approach may decrease the inductive bias and enhance the accuracy.
 

### 2012-11-10 article/article.tex
mozes

--- 2012-11-10 article/article.tex
@@
 The classifier accuracy is the average accuracy over all interations.
-We compare the following classifiers:
+We compared the following classifiers:
 
@@
 In the first method, averages of the selected scans belonging to a stimulus were used to estimate the parameters, while in the second method, the parameters were estimated for each scan.
-In both methods, the resulting probability of a class was the sum of log-likelyhoods per scan and the class with maximum probability was selected.
+In both methods, the resulting probability of a class was the sum of log-likelihoods per scan and the class with maximum probability was selected.
 The advantage of the first method is that fewer parameters need to be estimated with greater amount of examples.
@@
 
-Multivariate Gaussian.
+Multivariate Gaussian (MVG).
+In MVG, independence between feature vectors is not assumed.
+Instead, the mutivariate distrbution is used to estimate the probability of the classes, with the covariance matrix and the means as parameters.
+Again, the increase in parameters to be estimated compared to GNB (the correlations) may decrease generality, while also decreasing inductive bias.
 

### 2012-11-11 article/article.tex
mozes
spelling in article and added discussion conclusion

@@
 Specifically, the aforementioned studies try to infer the cognitive state of a subject from the brain activity that is evoked in a subject by performing that trial.
-For example, in an experiment where subjects look at a set of pictures (e.g.
-a star and a plus), they would be in a certain cognitive state that is the result of looking at these pictures.
+For example, in an experiment where subjects look at a set of pictures (e.g., a star and a plus), they would be in a certain cognitive state that is the result of looking at these pictures.
 One could then attempt to use machine learning on the fMRI signal that is evoked by the the brain activity associated with these cognitive states, and use this to learn what patterns of activity in the brain are associated with that cognitive state.
@@
 To avoid this problem it is recommended to only use linear classifiers when decoding cognitive states from fMRI data.
-The classifiers that we compare in this paper are Gaussian Naive Bayesian (GNB) classifiers, Linear Support Vector Machine (LSVM), and a Multivariate Gaussian (MVG) classifier.\\
+The classifiers that we compare in this paper are a Gaussian Naive Bayesian (GNB) classifier, Linear Support Vector Machine (LSVM), and a Multivariate Gaussian (MVG) classifier.\\
 \indent The goal of this study is to compare the classification performance of the cognitive states associated with looking at a picture or looking at a sentence between subjects.
@@
 First, groups of voxels that are anatomically defined as belonging to a certain brain region, a so-called region of interest (ROI), can be averaged into a single \emph{supervoxel}, which can be compared between subjects.
-Second, an anatomically defined talairach coordinate system can be used that maps the voxels of each subject onto an anatomically defined coordinate.
+Second, a talairach coordinate system can be used that maps the voxels of each subject onto an anatomically defined coordinate.
 The ROI averaging it is a rather crude method, and a lot of information of the individual voxels is lost.
@@
 In total, we have data available for approximately 4800 voxels per scan, 54 scans per trial, 40 trials per subject (excluding baseline trials) and 6 subjects.
-Voxels have been grouped into 25 regions of interest (see \textit{Feature selection}) that may be of use for analysis.
+Voxels have been grouped into 25 ROIs (see \textit{Feature selection}).
 
@@
 This problem is formally known as the \textit{Hughes effect}\cite{Hughes1054102} which states that the predictive power of a machine learning algorithm decreases as the dimensionality increases with a fixed number of training examples.
-To reducde the dimensionality of the input, several different methods were used.
+To reduce the dimensionality of the input, several different methods were used.
 
-ROI mapping.
-ROI mapping is a form of feature selection, wherein only a subset of the N-dimensional input space is used.
+\emph{ROI mapping.} ROI mapping is a form of feature selection, wherein only a subset of the N-dimensional input space is used.
 As described in the introduction, ROIs are anatomically defined brain areas classified as relevant by an expert in the field.
 We used 7 of the 25 defined ROIs, labeled  'CALC' 'LIPL' 'LT' 'LTRIA' 'LOPER' 'LIPS' and  'LDLPFC'.
-These ROIs have previously been found to be most relevant \cite{wa:2003betw}.
+These ROIs have previously been found to be most relevant in discriminating between the cognitive states of looking at sentences or pictures \cite{wa:2003betw}.
 To further reduce the number of features, average voxel values within these regions are taken as new input features.
@@
 
-Principal Component Analysis (PCA).
-As feature extraction method, we applied PCA.
+\emph{Principal Component Analysis (PCA).} As feature extraction method, we applied PCA.
 Ideally, the input vectors of the PCA should be the standardized Talairach coordinates, because we are conducting between-subject classification.
@@
 
-Temporal dimensionality reduction.
-As described in the introduction, the peak of the BOLD response is approximately 5 seconds after a stimulus is presented.
+\emph{Temporal dimensionality reduction.} The BOLD response peaks at approximately 5 seconds after a stimulus is presented (see Fig.
+1).
 This means that the first scans after a stimulus may not represent activity related to this stimulus and should not be used for classification.
@@
   	\caption{Haemodynamic BOLD response for ROI "CALC".
-The red line depicts a sentence-first trial (S-P) and the blue line depicts a picture-first trials (P-S).
+The red line depicts a sentence-picture trial (S-P) and the blue line depicts a picture-sentence trial (P-S).
 Average Bold activity for all subjects is plotted against the scan number.
@@
 For each classifier, a "leave one (subject) out" cross validation procedure was used to avoid overfitting.
-For N=6 iterations, N-1 subjects functioned as training set, while the other subject functioned as validation set.
-The classifier accuracy is the average accuracy over all interations.
+For N=6 iterations, N-1 subjects functioned as training set, while the remaining subject served as validation set.
+The classifier accuracy is the average accuracy over all iterations.
 We compared the following classifiers:
 
-Gaussian Naive Bayes (GNB).
-This classifier builds a generative model underlying a class, by estimating the means and the standard deviations of the input features from the training set \cite{bk:2004aplay}.
-It assumes a Gaussian distribution and strongly independent input features.
+\emph{Gaussian Naive Bayes (GNB).} This classifier builds a generative model underlying a class, by estimating the means and the standard deviations of the input features from the training set \cite{bk:2004aplay}.
+It assumes a Gaussian distribution and independent input features.
 Bayes rule is used to estimate the probability of a class (picture or sentence), given the input features of the validation set.
@@
 Rows indicate various feature selection methods.
-We applied PCA to the original data, while the rest methods were applied to the data belonging to the selected ROIs (cf.~Section~\ref{sec:dimensionalityReduction}).
+We applied PCA to the original data, while the other classifiers were applied to the data belonging to the selected ROIs (cf.~Section~\ref{sec:dimensionalityReduction}).
 For instance, \emph{Norm/Vox-M-SD ROIs} indicates that we normalised over the voxel values of the same voxel index using the mean and standard deviation, while \emph{Norm-M-SD ROIs} indicates that we normalised over the voxel values of the same scan index using the mean and standard deviation.
@@
 
-Furthermore, in order to find out how much percent sure we are that the actually accuracy of the classifiers is between some regions we performed an exact binomial test in each case, from which we obtained the confidence interval.
+Furthermore, to test the significance of the differences in performance between the different methods we performed an exact binomial test in each case, from which we obtained the confidence interval.
 The results were rounded to one decimal point.
@@
 
-Last but not least, we can distinguish the fact that Multivariate outperforms all the other classifiers, though we did not try a different data structure in the case of Linear SVM.
+Finally, we can distinguish the fact that Multivariate outperforms all the other classifiers, though we did not try a different data structure in the case of Linear SVM.
 However, we still do not expect that the performance of Linear SVM will be higher that this of Multivariate.
@@
 
+In this paper we have replicated earlier findings of decoding the cognitive states of looking at sentences or pictures from fMRI data.
+Specifically, by observing the patterns of activity for a group of subjects, we are able to predict with above-chance accuracy what the cognitive state of an unseen subject is.
+This shows that machine learning algorithms can be employed to capture useful information from noisy data, and generalise this information between subjects.\\
+\indent We found that even though all of our classification procedures yielded well above chance-level classification performance, the differences between the classifiers was very seldom significant.
+Still, we consistently found that the Multivariate Gaussian (MVG) classifier outperformed the Gaussian Naive Bayes (GNB) classifier and Linear Support Vector machine (LSV).
+This finding suggests that the interactions between different brain areas contain information about the cognitive state, and that these functional connectivity generalises to an extent to different subjects.\\
+\indent Still, the uncertainty in our predictions are relatively high, with maximum classification performance reaching close to 80\% accuracy for only two classes.
+The noisy nature of the fMRI data therefore makes it very difficult to draw hard conclusions when decoding a single cognitive state from a brain scan.
+This is something to be considered before employing fMRI as a diagnostic tool, or as evidence in court.\\
+\indent Overall we can say that machine learning provides a feasible way of extracting useful information from fMRI data.
+Furthermore, it avoids a strict multiple comparisons correction that is required in null hypothesis testing by using cross-validation.
+Also, the identification of patterns of activity associated with tasks is encouraged, rather than attempting to identifying brain regions that are involved in tasks.
+In doing this, much can be learned about how information is represented and processed in the brain.
+
 \bibliographystyle{abbrv}

### 2012-11-11 article/article.tex
py

--- 2012-11-10 article/article.tex
@@
 \label{sec:methods}
-General approach consisting of three steps, plus basic data acquisition and preprocessing.
+The general approach for applying machine learning techniques to fMRI data consists of the following steps:
 \begin{enumerate}
-  \item Feature selection
+  \item Preprocessing
+  \item Dimensionality reduction
   \item Training
@@
 \end{enumerate}
+The choices made in each of the steps affect the classifier accuracy.
+For a complete picture of the classification accuracies, we therefore try several methods at each step.
+In the following sections, these methods will be described.
 
-- comparison
-- classifying subtrials vs trials (ie 54 vs 27)
 - analysis / tools (python etc)
 
-\subsection{Data acquisition and dimensions}
+\subsection{Data Acquisition and Dimensions}
 The dataset used for our experiments was originally collected by Marcel Just et al.
@@
 In MVG, independence between feature vectors is not assumed.
-Instead, the mutivariate distrbution is used to estimate the probability of the classes, with the covariance matrix and the means as parameters.
+Instead, the mutivariate distribution is used to estimate the probability of the classes, with the covariance matrix and the means as parameters.
 Again, the increase in parameters to be estimated compared to GNB (the correlations) may decrease generality, while also decreasing inductive bias.

### 2012-11-11 article/article.tex
results

@@
 \subsection{Training Method and Classifiers}
-For each classifier, a "leave one (subject) out" cross validation procedure was used to avoid overfitting.
-For N=6 iterations, N-1 subjects functioned as training set, while the remaining subject served as validation set.
+\label{sec:trainingMethodsAndClassifiers}
+
+For each classifier, a ``leave one (subject) out" cross validation procedure was used to avoid overfitting.
+For $N=6$ iterations, $N-1$ subjects functioned as training set, while the remaining subject served as validation set.
 The classifier accuracy is the average accuracy over all iterations.
@@
 Columns depict the used classifiers.
-For the classifiers depicted with *, we used different structure for the data, as described in Section~\ref{}.
+For the classifiers depicted with *, we used a different method for structuring the data, as described in Section~\ref{sec:trainingMethodsAndClassifiers}.
 

### 2012-11-11 article/article.tex
Added PCA plot (correct version)

--- 2012-11-11 article/article.tex
@@
 relationships between stimuli'' (pattern characterization)~\cite{pereira2009machine}.
-While the main focus of this paper is on the application of machine learning techniques in pattern discrimination, we will briefly discuss these two other applications.
+While the main focus of this paper is on the application of machine learning techniques in pattern discrimination, we will briefly discuss these two other applications.\\
+
+Given a classifier that is able to discriminate on a dataset and identify a certain cognitive state, we may ask ourselves which parts of the brain contain the information leading to this identification.
+\emph{Pattern localization} addresses this problem by selecting the voxels that contribute the strongest and most reliably to the classification.
+Surprisingly, when taking the intersection of selected voxels across multiple subjects (for example, when performing cross-validation), this intersection is only a fairly small subset of voxels~\cite{pereira2009machine}.
 

### 2012-11-11 article/article.tex
results

--- 2012-11-11 article/article.tex
@@
 As a comparison measure, we used the performance of the classifiers for predicting the cognitive state that a subject was in.
-The predictive performance of Naive Bayes, Linear SVM, and Multivariate was compared.
+The predictive performance of GNB, LSVM, and MVG was compared.
 We observed that different feature selection and extraction methods, as well as different structure on the data affects the predictive performance of the classifiers.
@@
 Rows indicate various feature selection methods.
-We applied PCA to the original data, while the other classifiers were applied to the data belonging to the selected ROIs (cf.~Section~\ref{sec:dimensionalityReduction}).
+We applied PCA to the original data, while the rest feature selection methods were applied to the data belonging to the selected ROIs (cf.~Section~\ref{sec:dimensionalityReduction}).
 For instance, \emph{Norm/Vox-M-SD ROIs} indicates that we normalised over the voxel values of the same voxel index using the mean and standard deviation, while \emph{Norm-M-SD ROIs} indicates that we normalised over the voxel values of the same scan index using the mean and standard deviation.
@@
 The results were rounded to one decimal point.
-For the following comparison we are only looking at the case where \emph{Norm/Vox-M-SD ROIs} was used.
+For the rest of this section and to explain some of the results we are looking only at the case when \emph{Norm/Vox-M-SD ROIs} is used as a feature selection method.
 
-From Table~\ref{tab:results} we can distinguish the fact that when PCA is used for feature extraction, the performance of Naive Bayes is lower than when other feature selection methods are used.
-The confidence interval of the predictive performance of Naive Bayes when PCA is used is 95\% $CI: 56.9 - 65.8$ which suggests that...
+From Table~\ref{tab:results} we can distinguish the fact that when PCA is used for feature extraction, the performance of GNB is lower than when other feature selection methods are used.
+The confidence interval of the predictive performance of GNB when PCA is used is 95\% $CI: 56.9 - 65.8$ which suggests that...
 
-Comparing classifiers when the same data structure is used, we can distinguish the fact that Multivariate performs better than Naive Bayes.
-Also that Linear SVM performs better than Naive Bayes, though not significantly.
-The confidence interval of Naive Bayes is 95\% $CI: 68.9-77.0$ and of Linear SVM is 95\% $CI: 72.6-80.4$, which suggests that their performances do not differ significantly.
+Comparing classifiers when the same data structure is used, we can distinguish the fact that MVG performs better than GNB.
+Also that LSVM performs better than GNB, though not significantly.
+The confidence interval of GNB is 95\% $CI: 68.9-77.0$ and of LSVM is 95\% $CI: 72.6-80.4$, which suggests that their performances do not differ significantly.
 
-Moreover, when different data structure is used the performance increases, though not significantly.
-This can be seen by comparing the performance of Naive Bayes when different data structure is used.
+Moreover, when different method for structuring the data is used the performance increases, though not significantly.
+This can be seen by comparing the performance of GNB when different method for structuring the data is used.
 In the first case it is 73.13\% while in the second case is 76.88\%.
@@
 
-Finally, we can distinguish the fact that Multivariate outperforms all the other classifiers, though we did not try a different data structure in the case of Linear SVM.
-However, we still do not expect that the performance of Linear SVM will be higher that this of Multivariate.
-The confidence interval of Multivariate is 95\% $CI: 75.7-83.1$.
+Finally, we can distinguish the fact that MVG outperforms all the other classifiers, though we did not try a different data structure in the case of LSVM.
+However, we still do not expect that the performance of LSVM will be higher that this of MVG.
+The confidence interval of MVG is 95\% $CI: 75.7-83.1$.
 

### 2012-11-11 article/article.tex
peel

--- 2012-11-11 article/article.tex
@@
 relationships between stimuli'' (pattern characterization)~\cite{pereira2009machine}.
-While the main focus of this paper is on the application of machine learning techniques in pattern discrimination, we will briefly discuss these two other applications.\\
-
-Given a classifier that is able to discriminate on a dataset and identify a certain cognitive state, we may ask ourselves which parts of the brain contain the information leading to this identification.
-\emph{Pattern localization} addresses this problem by selecting the voxels that contribute the strongest and most reliably to the classification.
-Surprisingly, when taking the intersection of selected voxels across multiple subjects (for example, when performing cross-validation), this intersection is only a fairly small subset of voxels~\cite{pereira2009machine}.
+While the main focus of this paper is on the application of machine learning techniques in pattern discrimination, we will briefly discuss these two other applications.
 
@@
 \subsection{Training Method and Classifiers}
-\label{sec:trainingMethodsAndClassifiers}
-
-For each classifier, a ``leave one (subject) out" cross validation procedure was used to avoid overfitting.
-For $N=6$ iterations, $N-1$ subjects functioned as training set, while the remaining subject served as validation set.
+For each classifier, a "leave one (subject) out" cross validation procedure was used to avoid overfitting.
+For N=6 iterations, N-1 subjects functioned as training set, while the remaining subject served as validation set.
 The classifier accuracy is the average accuracy over all iterations.
@@
 Two approaches were used to estimate the parameters of the model.
-In the first method, averages of the selected scans belonging to a stimulus were used to estimate the parameters, while in the second method, the parameters were estimated for each scan.
+In the first method, averages of the selected scans belonging to a stimulus were used to estimate the parameters, while in the second method, the parameters were estimated for each scan separately.
 In both methods, the resulting probability of a class was the sum of log-likelihoods per scan and the class with maximum probability was selected.
@@
 
-Linear Support Vector Machine.
+Linear Support Vector Machine (LSVM).
+The LSVM is a discriminant based approach to classification.
+For an elaborate description see \cite{bk:2004aplay}.
+For each selected scan belonging to a stimulus, the LSVM classifies it as either a picture or a sentence.
+The final classification of the stimulus is
 
@@
 Columns depict the used classifiers.
-For the classifiers depicted with *, we used a different method for structuring the data, as described in Section~\ref{sec:trainingMethodsAndClassifiers}.
+For the classifiers depicted with *, we used different structure for the data, as described in Section~\ref{}.
 

### 2012-11-11 article/article.tex
Expanded application

--- 2012-11-11 article/article.tex
@@
 relationships between stimuli'' (pattern characterization)~\cite{pereira2009machine}.
-While the main focus of this paper is on the application of machine learning techniques in pattern discrimination, we will briefly discuss these two other applications.
+While the main focus of this paper is on the application of machine learning techniques in pattern discrimination, we will briefly discuss these two other applications.\\
+
+Given a classifier that is able to discriminate on a dataset and identify a certain cognitive state, we may ask ourselves which parts of the brain contain the information leading to this identification.
+\emph{Pattern localization} addresses this problem by selecting the voxels that contribute the strongest and most reliably to the classification.
+Surprisingly, when taking the intersection of selected voxels across multiple subjects (for example, when performing cross-validation), this intersection is only a fairly small subset of voxels~\cite{pereira2009machine}.
+The reason for this is that brains differ between subjects in their size, shape, and location of particular cells.
+As a result, activation patterns often do not line up well.
+One of the strategies to deal with this is to use the set of voxels active for each subject as seeds in a clustering algorithm to find similar voxels.\\
+
+In \emph{pattern characterization}, the question of \emph{how} class information is represented takes a central place.
+While this paper will focus on linear classifiers, some studies in this area resort to a more sophisticated approach and revolve around the question how to correspond relationships in stumulus with those in fMRI data.
+For example,~\cite{} researches predicting fMRI response for new words based on their similarity to words for which the response is known.
 
@@
 Two approaches were used to estimate the parameters of the model.
-In the first method, averages of the selected scans belonging to a stimulus were used to estimate the parameters, while in the second method, the parameters were estimated for each scan separately.
+In the first method, averages of the selected scans belonging to a stimulus were used to estimate the parameters, while in the second method, the parameters were estimated for each scan.
 In both methods, the resulting probability of a class was the sum of log-likelihoods per scan and the class with maximum probability was selected.
@@
 
-Linear Support Vector Machine (LSVM).
-The LSVM is a discriminant based approach to classification.
-For an elaborate description see \cite{bk:2004aplay}.
-For each selected scan belonging to a stimulus, the LSVM classifies it as either a picture or a sentence.
-The final classification of the stimulus is
+Linear Support Vector Machine.
 

### 2012-11-11 article/article.tex
peel

--- 2012-11-11 article/article.tex
@@
 In the first method, averages of the selected scans belonging to a stimulus were used to estimate the parameters, while in the second method, the parameters were estimated for each scan separately.
-In both methods, the resulting probability of a class was the sum of log-likelihoods per scan and the class with maximum probability was selected.
+In both methods, the resulting log-probability of a class was the sum of log-likelihoods per scan and the class with maximum log-probability was selected.
 The advantage of the first method is that fewer parameters need to be estimated with greater amount of examples.

### 2012-11-11 article/article.tex
methods

--- 2012-11-11 article/article.tex
@@
 \usepackage{mathptmx}
-\usepackage{graphicx,xcolor,booktabs, tabularx,amsmath}
+\usepackage{graphicx,xcolor,booktabs, tabularx, amsmath}
 \usepackage{times}
@@
 \emph{Gaussian Naive Bayes (GNB).} This classifier builds a generative model underlying a class, by estimating the means and the standard deviations of the input features from the training set \cite{bk:2004aplay}.
-It assumes a Gaussian distribution and independent input features.
-Bayes rule is used to estimate the probability of a class (picture or sentence), given the input features of the validation set.
+It assumes a Gaussian distribution (cf.~Eq.~\ref{eq:gaussianDistribution}) and independent input features.
+
+\begin{equation}
+\label{eq:gaussianDistribution}
+p(X|\mu, \sigma) = \frac{1}{\sigma\sqrt{2\pi}}\exp\Big[-\frac{(x - \bar{\mu})^2}{2\sigma^2}\Big]
+\end{equation}
+
+where $\mu$ denotes the mean and $\sigma$ the standard deviation.
+Bayes rule is used to estimate the probability of a class (picture or sentence), given the input features of the validation set (cf.~Eq.~\ref{eq:bayesRule}).
+
+\begin{equation}
+\label{eq:bayesRule}
+P(C_{i}|X) = \frac{p(X|C_{i})P(C_{i})}{P(X)}
+\end{equation}
+
+where $C_i$ denotes the probability of the class---sentence or picture.
 Two approaches were used to estimate the parameters of the model.
-In the first method, averages of the selected scans belonging to a stimulus were used to estimate the parameters, while in the second method, the parameters were estimated for each scan separately.
+In the first method, averages of the selected scans belonging to a stimulus were used to estimate the parameters, while in the second method, the parameters were estimated for each scan.
 In both methods, the resulting probability of a class was the sum of log-likelihoods per scan and the class with maximum probability was selected.

### 2012-11-11 article/article.tex
Expanded 2 (characterization)

--- 2012-11-11 article/article.tex
@@
 \emph{Gaussian Naive Bayes (GNB).} This classifier builds a generative model underlying a class, by estimating the means and the standard deviations of the input features from the training set \cite{bk:2004aplay}.
-It assumes a Gaussian distribution and independent input features.
-Bayes rule is used to estimate the probability of a class (picture or sentence), given the input features of the validation set.
+It assumes a Gaussian distribution (cf.~Eq.~\ref{eq:gaussianDistribution}) and independent input features.
+
+\begin{equation}
+\label{eq:gaussianDistribution}
+p(X|\mu, \sigma) = \frac{1}{\sigma\sqrt{2\pi}}\exp\Big[-\frac{(x - \bar{\mu})^2}{2\sigma^2}\Big]
+\end{equation}
+
+where $\mu$ denotes the mean and $\sigma$ the standard deviation.
+Bayes rule is used to estimate the probability of a class (picture or sentence), given the input features of the validation set (cf.~Eq.~\ref{eq:bayesRule}).
+
+\begin{equation}
+\label{eq:bayesRule}
+P(C_{i}|X) = \frac{p(X|C_{i})P(C_{i})}{P(X)}
+\end{equation}
+
+where $C_i$ denotes the probability of the class---sentence or picture.
 Two approaches were used to estimate the parameters of the model.

### 2012-11-11 article/article.tex

@@
 Surprisingly, when taking the intersection of selected voxels across multiple subjects (for example, when performing cross-validation), this intersection is only a fairly small subset of voxels~\cite{pereira2009machine}.
+The reason for this is that brains differ between subjects in their size, shape, and location of particular cells.
+As a result, activation patterns often do not line up well.
+One of the strategies to deal with this is to use the set of voxels active for each subject as seeds in a clustering algorithm to find similar voxels.\\
+
+In \emph{pattern characterization}, the question of \emph{how} class information is represented takes a central place.
+While this paper will focus on linear classifiers, some studies in this area resort to a more sophisticated approach and revolve around the question how to correspond relationships in stumulus with those in fMRI data.
+For example,~\cite{mitchell2008predicting} researches predicting fMRI response for new words based on their similarity to words for which the response is known.
 

### 2012-11-11 article/article.tex
Rewrote methods introduction

@@
 \label{sec:methods}
-The general approach for applying machine learning techniques to fMRI data consists of the following steps:
+We now continue by discussing the methods applied in our research. The general approach taken for each of the algorithms discussed consists of the following steps:
 \begin{enumerate}
@@
 \end{enumerate}
-The choices made in each of the steps affect the classifier accuracy.
-For a complete picture of the classification accuracies, we therefore try several methods at each step.
-In the following sections, these methods will be described.
-
-- analysis / tools (python etc)
+Preprocessing was done to transform the dataset into a format suitable for input into the classification algorithm.
+Dimensionality reduction was a direct result of the large dimensionality of the dataset and may consist of both feature extraction as well as feature selection.
+Finally, the approach taken to training and testing the classifier deals with preventing overfitting the classifier to training data and verifying accuracy for cross-subject classification.
+The choices made in each of these steps may affect classifier accuracy, so to obain a complete understanding of classification accuracies we have tried several methods at each step.
+In the following sections, these methods will be described in more detail.\\
+The implementation of the algorithms was done using Python combined with modules tailored to mathematics (NumPy, SciPy) and machine learning (ScikitLearn).
+The original dataset was made available as a MatLab file and we used preprocessing in MatLab to create a number of files containing modified versions of this data, such as normalization and ROI grouping (see section~\ref{preprocessing}).
+The source code for the algorithms is available from the authors on request.
 
@@
 \subsection{Preprocessing}
+\label{ssec:preprocessing}
 

### 2012-11-11 article/article.tex
methods

--- 2012-11-11 article/article.tex
@@
 The classifier accuracy is the average accuracy over all iterations.
-We compared the following classifiers:
+We compared the following classifiers:\\
 
-\emph{Gaussian Naive Bayes (GNB).} This classifier builds a generative model underlying a class, by estimating the means and the standard deviations of the input features from the training set \cite{bk:2004aplay}.
+\textbf{\emph{Gaussian Naive Bayes (GNB).}} This classifier builds a generative model underlying a class, by estimating the means and the standard deviations of the input features from the training set \cite{bk:2004aplay}.
 It assumes a Gaussian distribution (cf.~Eq.~\ref{eq:gaussianDistribution}) and independent input features.
@@
 In the first method, averages of the selected scans belonging to a stimulus were used to estimate the parameters, while in the second method, the parameters were estimated for each scan separately.
-In both methods, the resulting log-probability of a class was the sum of log-likelihoods per scan and the class with maximum log-probability was selected.
+In both methods the decision of the classifier is given by Eq.~\ref{eq:maxLikelihood}, which is the maximum likelihood of the two classes over all scans.
+
+\begin{equation}
+\label{eq:maxLikelihood}
+\operatorname{arg\,max}_i \sum_{t=1}^{N}{\log{P(C_{i}|X)}}
+\end{equation}
+
+where $t$ denotes the scan index.
 The advantage of the first method is that fewer parameters need to be estimated with greater amount of examples.
@@
 
-Linear Support Vector Machine (LSVM).
-The LSVM is a discriminant based approach to classification.
+\textbf{\emph{Linear Support Vector Machine (LSVM).}} The LSVM is a discriminant based approach to classification.
 For an elaborate description see \cite{bk:2004aplay}.
@@
 
-Multivariate Gaussian (MVG).
-In MVG, independence between feature vectors is not assumed.
+\textbf{\emph{Multivariate Gaussian (MVG).}} In MVG, independence between feature vectors is not assumed.
 Instead, the mutivariate distribution is used to estimate the probability of the classes, with the covariance matrix and the means as parameters.

### 2012-11-11 article/article.tex
methods

@@
 \usepackage{times}
+\usepackage{breqn}
 \newcommand{\todo}[1]{\textbf{\textcolor{blue}{Todo: #1 }}}
@@
 \label{eq:maxLikelihood}
-\operatorname{arg\,max}_i \sum_{t=1}^{N}{\log{P(C_{i}|X)}}
+\operatorname{arg\,max}_i \sum_{t=1}^{N}{\log{P(C_{i}|x^{t})}}
 \end{equation}
@@
 For each selected scan belonging to a stimulus, the LSVM classifies it as either a picture or a sentence.
-The final classification of the stimulus is
+The final classification of the stimulus is:
+
+\begin{equation}
+\label{eq:argmaxsvm}
+\operatorname{arg\,max}_i \sum_{t=1}^{N}{b_{i}^{t}}
+\end{equation}
+
+where $t$ denotes the scan index and $b$ is given by:
+
+\begin{equation}
+b_{i} = \left\{
+\begin{array}{l l}
+1 \quad \text{if $C_{i}$ is classified}\\
+0 \quad \text{otherwise}\\
+\end{array} \right.
+\end{equation}
 
 \textbf{\emph{Multivariate Gaussian (MVG).}} In MVG, independence between feature vectors is not assumed.
-Instead, the mutivariate distribution is used to estimate the probability of the classes, with the covariance matrix and the means as parameters.
+Instead, the multivariate distribution is used to estimate the probability of the classes, with the covariance matrix and the means as parameters.
 Again, the increase in parameters to be estimated compared to GNB (the correlations) may decrease generality, while also decreasing inductive bias.
 
-- same for each algorithm:
-    - What features did we select?
-    - How did we train?
-    - How did we test?
-        - How did we calculate performance?
-- algorithms:
-    - gaussian naive-bayes (gnb)
-        - pca + gnb
-    - multivariate
-    - svm
-    - why multivariate is not a priori better than naive: less assumption, higher generalizability, better estimation parameters.
+\begin{dmath}
+\label{eq:multivariate}
+g_{i}(\mathbf{x}) = -\frac{d}{2}\log{2\pi}-\frac{1}{2}\log{|\Sigma_{i}|}-\frac{1}{2}(\mathbf{x}-\mu_{i})^{T}\Sigma_{i}^{-1}(\textbf{x}-\mu_{i})+\log{P(C_{i})}
+\end{dmath}
 

### 2012-11-11 article/article.tex
Preprocessing draft

--- 2012-11-11 article/article.tex
@@
 
-Here we will need a nice picture, perhaps a pagewide one with labels of data for 1 trial?
-
 \subsection{Preprocessing}
 \label{ssec:preprocessing}
+The dataset contains activation values for each voxels with respect to a baseline activation value for a subject (measured before performing the task).
+To make optimal use of the information value of each voxel, we have created normalized copies of the dataset to use as input for the algorithms.
+The effect of this preprocessing step depends on assumptions made about the form of our data, which are not available.
+Instead, we had to verify the effect by looking at the resulting classification results (see table~\ref{results}).
+As a first step, we applied row normalization by setting the mean activation value to zero for each trial.
+The assumption was that the relative activation has more predictive power than the actual value.
+We then continued by transforming these values to z-scores (by setting the standard-deviation to 1) but this turned out to negatively impact accuracy (indicating that the transformation lost information).
+For completeness, we also performed normalization of higher-order properties (such as skew and kurtosis) but found a significant further decrease in accuracy.
+We therefore did not further researched performance on these sets.
+One other approach we tried was the application of column normalization, setting the mean value across trials to zero.
 
@@
 The classifier accuracy is the average accuracy over all iterations.
-We compared the following classifiers:\\
+We compared the following classifiers:
 
-\textbf{\emph{Gaussian Naive Bayes (GNB).}} This classifier builds a generative model underlying a class, by estimating the means and the standard deviations of the input features from the training set \cite{bk:2004aplay}.
+\emph{Gaussian Naive Bayes (GNB).} This classifier builds a generative model underlying a class, by estimating the means and the standard deviations of the input features from the training set \cite{bk:2004aplay}.
 It assumes a Gaussian distribution (cf.~Eq.~\ref{eq:gaussianDistribution}) and independent input features.
@@
 In the first method, averages of the selected scans belonging to a stimulus were used to estimate the parameters, while in the second method, the parameters were estimated for each scan separately.
-In both methods the decision of the classifier is given by Eq.~\ref{eq:maxLikelihood}, which is the maximum likelihood of the two classes over all scans.
-
-\begin{equation}
-\label{eq:maxLikelihood}
-\operatorname{arg\,max}_i \sum_{t=1}^{N}{\log{P(C_{i}|X)}}
-\end{equation}
-
-where $t$ denotes the scan index.
+In both methods, the resulting log-probability of a class was the sum of log-likelihoods per scan and the class with maximum log-probability was selected.
 The advantage of the first method is that fewer parameters need to be estimated with greater amount of examples.
@@
 
-\textbf{\emph{Linear Support Vector Machine (LSVM).}} The LSVM is a discriminant based approach to classification.
+Linear Support Vector Machine (LSVM).
+The LSVM is a discriminant based approach to classification.
 For an elaborate description see \cite{bk:2004aplay}.
@@
 
-\textbf{\emph{Multivariate Gaussian (MVG).}} In MVG, independence between feature vectors is not assumed.
+Multivariate Gaussian (MVG).
+In MVG, independence between feature vectors is not assumed.
 Instead, the mutivariate distribution is used to estimate the probability of the classes, with the covariance matrix and the means as parameters.

### 2012-11-11 article/article.tex
peel

--- 2012-11-11 article/article.tex
@@
 Instead, the multivariate distribution is used to estimate the probability of the classes, with the covariance matrix and the means as parameters.
-Again, the increase in parameters to be estimated compared to GNB (the correlations) may decrease generality, while also decreasing inductive bias.
-
+The log-probability of a class becomes:
 \begin{dmath}
@@
 \end{dmath}
+
+where $d$ denotes the dimensionality of $x$ and $\Sigma$ denotes the covariance matrix.
+The parameters are estimated per scan and the decision method is similar to GNB.
+The increase in parameters to be estimated compared to GNB (the correlations) may decrease generality, while also decreasing inductive bias.
 

### 2012-11-11 article/article.tex
preprocessing

--- 2012-11-11 article/article.tex
@@
 \label{ssec:preprocessing}
-The dataset contains activation values for each voxels with respect to a baseline activation value for a subject (measured before performing the task).
+The dataset contains activation values for each voxels with respect to a baseline activation value for a subject, measured before performing the task.
 To make optimal use of the information value of each voxel, we have created normalized copies of the dataset to use as input for the algorithms.
 The effect of this preprocessing step depends on assumptions made about the form of our data, which are not available.
-Instead, we had to verify the effect by looking at the resulting classification results (see table~\ref{results}).
+Instead, we had to verify the effect by looking at the resulting classification results (cf.~Table~\ref{tab:results}).
 As a first step, we applied row normalization by setting the mean activation value to zero for each trial.
 The assumption was that the relative activation has more predictive power than the actual value.
-We then continued by transforming these values to z-scores (by setting the standard-deviation to 1) but this turned out to negatively impact accuracy (indicating that the transformation lost information).
+We then continued by transforming these values to z-scores (by setting the standard-deviation to 1) but this turned out to negatively impact accuracy, indicating that the transformation lost information.
 For completeness, we also performed normalization of higher-order properties (such as skew and kurtosis) but found a significant further decrease in accuracy.

### 2012-11-11 article/article.tex
preprocessing

@@
 Instead, we had to verify the effect by looking at the resulting classification results (cf.~Table~\ref{tab:results}).
+
 As a first step, we applied row normalization by setting the mean activation value to zero for each trial.
 The assumption was that the relative activation has more predictive power than the actual value.
-We then continued by transforming these values to z-scores (by setting the standard-deviation to 1) but this turned out to negatively impact accuracy, indicating that the transformation lost information.
-For completeness, we also performed normalization of higher-order properties (such as skew and kurtosis) but found a significant further decrease in accuracy.
+We then continued by transforming these values to z-scores by setting the standard-deviation to 1.
+However, this turned out to negatively impact accuracy, indicating that the transformation lost information.
+For completeness, we also performed normalization of higher-order properties, such as skew and kurtosis, but found a significant further decrease in accuracy.
 We therefore did not further researched performance on these sets.
@@
 
-\emph{ROI mapping.} ROI mapping is a form of feature selection, wherein only a subset of the N-dimensional input space is used.
+\textbf{\emph{ROI mapping.}} ROI mapping is a form of feature selection, wherein only a subset of the N-dimensional input space is used.
 As described in the introduction, ROIs are anatomically defined brain areas classified as relevant by an expert in the field.
@@
 The parameters are estimated per scan and the decision method is similar to GNB.
-The increase in parameters to be estimated compared to GNB (the correlations) may decrease generality, while also decreasing inductive bias.
+The increase in parameters to be estimated compared to GNB (the correlations) may decrease generalization, while also decreasing inductive bias.
 

### 2012-11-11 article/article.tex
preprocessing

@@
 
-\emph{Principal Component Analysis (PCA).} As feature extraction method, we applied PCA.
+\textbf{\emph{Principal Component Analysis (PCA).}} As feature extraction method, we applied PCA.
 Ideally, the input vectors of the PCA should be the standardized Talairach coordinates, because we are conducting between-subject classification.
@@
 
-\emph{Temporal dimensionality reduction.} The BOLD response peaks at approximately 5 seconds after a stimulus is presented (see Fig.
+\textbf{\emph{Temporal dimensionality reduction.}} The BOLD response peaks at approximately 5 seconds after a stimulus is presented (see Fig.
 1).

### 2012-11-11 article/article.tex
preprocessing: mentiond across subjects

@@
 The dataset contains activation values for each voxels with respect to a baseline activation value for a subject, measured before performing the task.
-To make optimal use of the information value of each voxel, we have created normalized copies of the dataset to use as input for the algorithms.
+However, each subject may respond differently in terms of strength of voxel activation, making it harder to classify across subjects.
+To counter this, and make optimal use of the information carried by each voxel, we have created normalized copies of the dataset to use as input for the algorithms.
 The effect of this preprocessing step depends on assumptions made about the form of our data, which are not available.

### 2012-11-11 article/article.tex
results

@@
 {\small
-\begin{tabular}{c|c|c|c|c}
-& \textbf{Naive Bayes} & \textbf{Linear SVM} & \textbf{Naive Bayes*} & \textbf{Multivariate*} \\ \hline
+\begin{tabular}{|c|c|c|c|c|}
+\hline
+\textbf{Extraction/Selection Method} & \textbf{Naive Bayes} & \textbf{Linear SVM} & \textbf{Naive Bayes*} & \textbf{Multivariate*} \\ \hline
 PCA & 61.46 &  - & - & - \\
@@
 Norm/Vox-M ROIs & 73.13 & 76.67 & 76.88 & 79.58 \\
+\hline
 \end{tabular}}

### 2012-11-11 article/article.tex
peel

--- 2012-11-11 article/article.tex
@@
 These ROIs have previously been found to be most relevant in discriminating between the cognitive states of looking at sentences or pictures \cite{wa:2003betw}.
-To further reduce the number of features, average voxel values within these regions are taken as new input features.
-As other higher-order features such as the standard deviation, kurtosis and skewness within ROIs did not increase predictability, these were not included in the subsequent analysis.
+To further reduce the number of features, average voxel values and standard deviations within these regions are taken as new input features.
+As other higher-order features such as $kurtosis$ and $skewness$ within ROIs did not increase predictability, these were not included in the subsequent analysis.
 The disadvantage of using ROIs is that they require \textit{a priori} knowledge.

### 2012-11-11 article/article.tex
results

--- 2012-11-11 article/article.tex
@@
 We applied PCA to the original data, while the rest feature selection methods were applied to the data belonging to the selected ROIs (cf.~Section~\ref{sec:dimensionalityReduction}).
-For instance, \emph{Norm/Vox-M-SD ROIs} indicates that we normalised over the voxel values of the same voxel index using the mean and standard deviation, while \emph{Norm-M-SD ROIs} indicates that we normalised over the voxel values of the same scan index using the mean and standard deviation.
+For instance, \emph{Norm/Vox-M ROIs} indicates that we normalised over the voxel values of the same voxel index using only the mean, while \emph{Norm-M-SD ROIs} indicates that we normalised over the voxel values of the same scan index using the mean and standard deviation.
 Columns depict the used classifiers.
-For the classifiers depicted with *, we used a different method for structuring the data, as described in Section~\ref{sec:trainingMethodsAndClassifiers}.
+For the classifiers depicted with *, we used a different method for structuring the data as described in Section~\ref{sec:trainingMethodsAndClassifiers}.
 
@@
 The results were rounded to one decimal point.
-For the rest of this section and to explain some of the results we are looking only at the case when \emph{Norm/Vox-M-SD ROIs} is used as a feature selection method.
+For the rest of this section and to explain some of the results we are looking only at the case when \emph{Norm/Vox-M ROIs} is used as a feature selection method.
 

### 2012-11-11 article/article.tex
PCA Components results added

--- 2012-11-11 article/article.tex
@@
 These ROIs have previously been found to be most relevant in discriminating between the cognitive states of looking at sentences or pictures \cite{wa:2003betw}.
-To further reduce the number of features, average voxel values and standard deviations within these regions are taken as new input features.
-As other higher-order features such as $kurtosis$ and $skewness$ within ROIs did not increase predictability, these were not included in the subsequent analysis.
+To further reduce the number of features, average voxel values within these regions are taken as new input features.
+As other higher-order features such as the standard deviation, kurtosis and skewness within ROIs did not increase predictability, these were not included in the subsequent analysis.
 The disadvantage of using ROIs is that they require \textit{a priori} knowledge.
@@
 {\small
-\begin{tabular}{|c|c|c|c|c|}
-\hline
-\textbf{Extraction/Selection Method} & \textbf{Naive Bayes} & \textbf{Linear SVM} & \textbf{Naive Bayes*} & \textbf{Multivariate*} \\ \hline
+\begin{tabular}{c|c|c|c|c}
+& \textbf{Naive Bayes} & \textbf{Linear SVM} & \textbf{Naive Bayes*} & \textbf{Multivariate*} \\ \hline
 PCA & 61.46 &  - & - & - \\
@@
 Norm/Vox-M ROIs & 73.13 & 76.67 & 76.88 & 79.58 \\
-\hline
 \end{tabular}}
@@
 From Table~\ref{tab:results} we can distinguish the fact that when PCA is used for feature extraction, the performance of GNB is lower than when other feature selection methods are used.
-The confidence interval of the predictive performance of GNB when PCA is used is 95\% $CI: 56.9 - 65.8$ which suggests that...
+The confidence interval of the predictive performance of GNB when PCA is used is 95\% $CI: 56.9 - 65.8$ which is significantly worse than other dimension reduction techniques.
+This suggests that maximizing the variance of the data does not maximize information.
+Another explanation could be that the data is not suited to be aligned by voxel index (see section~\ref{sec:dimensionalityReduction}).
+The performance listed in Table~\ref{tab:results} is for PCA using 149 components.
+This value was found by performing classification with a varying amount of components, the results for which are listed in figure~\ref{fig:pca_components}.
+
+\begin{figure}
+	\centering
+	\includegraphics[width=90mm]{figures/pca_plot_by_index.png}
+  	\caption{Classification accuracy of GNB after performing PCA with a varying amount of components.}
+  	\label{fig:pca_components}
+\end{figure}
 

### 2012-11-11 article/article.tex
peel

--- 2012-11-11 article/article.tex
@@
 
-where $d$ denotes the dimensionality of $x$ and $\Sigma$ denotes the covariance matrix.
+where $g_{i}(\mathbf{x})$ denotes the log-probability of a class, $d$ denotes the dimensionality of $x$ and $\Sigma$ denotes the covariance matrix.
 The parameters are estimated per scan and the decision method is similar to GNB.
@@
 We applied PCA to the original data, while the rest feature selection methods were applied to the data belonging to the selected ROIs (cf.~Section~\ref{sec:dimensionalityReduction}).
-For instance, \emph{Norm/Vox-M ROIs} indicates that we normalised over the voxel values of the same voxel index using only the mean, while \emph{Norm-M-SD ROIs} indicates that we normalised over the voxel values of the same scan index using the mean and standard deviation.
+For instance, \emph{Norm/Vox-M-SD ROIs} indicates that we normalised over the voxel values of the same voxel index using the mean and standard deviation, while \emph{Norm-M-SD ROIs} indicates that we normalised over the voxel values of the same scan index using the mean and standard deviation.
 Columns depict the used classifiers.
-For the classifiers depicted with *, we used a different method for structuring the data as described in Section~\ref{sec:trainingMethodsAndClassifiers}.
+For the classifiers depicted with *, we used a different method for structuring the data, as described in Section~\ref{sec:trainingMethodsAndClassifiers}.
 
@@
 The results were rounded to one decimal point.
-For the rest of this section and to explain some of the results we are looking only at the case when \emph{Norm/Vox-M ROIs} is used as a feature selection method.
+For the rest of this section and to explain some of the results we are looking only at the case when \emph{Norm/Vox-M-SD ROIs} is used as a feature selection method.
 

### 2012-11-11 article/article.tex

--- 2012-11-11 article/article.tex
@@
 From Table~\ref{tab:results} we can distinguish the fact that when PCA is used for feature extraction, the performance of GNB is lower than when other feature selection methods are used.
-The confidence interval of the predictive performance of GNB when PCA is used is 95\% $CI: 56.9 - 65.8$ which is significantly worse than other dimension reduction techniques.
-This suggests that maximizing the variance of the data does not maximize information.
-Another explanation could be that the data is not suited to be aligned by voxel index (see section~\ref{sec:dimensionalityReduction}).
-The performance listed in Table~\ref{tab:results} is for PCA using 149 components.
-This value was found by performing classification with a varying amount of components, the results for which are listed in figure~\ref{fig:pca_components}.
-
-\begin{figure}
-	\centering
-	\includegraphics[width=90mm]{figures/pca_plot_by_index.png}
-  	\caption{Classification accuracy of GNB after performing PCA with a varying amount of components.}
-  	\label{fig:pca_components}
-\end{figure}
+The confidence interval of the predictive performance of GNB when PCA is used is 95\% $CI: 56.9 - 65.8$ which suggests that...
 

### 2012-11-11 article/article.tex
spelling

@@
 relationships between stimuli'' (pattern characterization)~\cite{pereira2009machine}.
-While the main focus of this paper is on the application of machine learning techniques in pattern discrimination, we will briefly discuss these two other applications.\\
+While the main focus of this paper is on the application of machine learning techniques in pattern discrimination, we will briefly discuss these two other applications.
 
@@
 Preprocessing was done to transform the dataset into a format suitable for input into the classification algorithm.
-Dimensionality reduction was a direct result of the large dimensionality of the dataset and may consist of both feature extraction as well as feature selection.
+Dimensionality reduction was required to cope with the high dimensionality of the dataset and may consist of both feature extraction as well as feature selection.
 Finally, the approach taken to training and testing the classifier deals with preventing overfitting the classifier to training data and verifying accuracy for cross-subject classification.
@@
 \subsection{Data Acquisition and Dimensions}
-The dataset used for our experiments was originally collected by Marcel Just et al.
-from Carnegie Mellon University's Center for Cognitive Brain Imaging and published as the StarPlus fMRI dataset.
+The dataset used for our experiments was originally collected by Marcel Just and colleagues from Carnegie Mellon University's Center for Cognitive Brain Imaging and published as the StarPlus fMRI dataset.
 This set consists of fMRI data for 6 subjects that was obtained while having each subject participate in a number of trials.

### 2012-11-11 article/article.tex
spelling and style

@@
 	\includegraphics[width=90mm]{figures/scan_hemodynamic_BOLD_roi}
-  	\caption{Haemodynamic BOLD response for ROI "CALC".
+  	\caption{Haemodynamic response for ROI "CALC".
 The red line depicts a sentence-picture trial (S-P) and the blue line depicts a picture-sentence trial (P-S).
-Average Bold activity for all subjects is plotted against the scan number.
-The coloured regions depict an interval of 1 standard deviation.}
+Average BOLD signal for all subjects is plotted against the scan number.
+The coloured regions depict the mean BOLD signal +/- 1 standard deviation.}
   	\label{fig:boldroi}
@@
 	\includegraphics[width=90mm]{figures/pca_plot_by_index.png}
-  	\caption{Classification accuracy of GNB after performing PCA with a varying amount of components.}
+  	\caption{Classification accuracy of GNB after performing PCA with an increasing number of components.}
   	\label{fig:pca_components}

### 2012-11-11 article/article.tex
PCA results

--- 2012-11-11 article/article.tex
@@
 From Table~\ref{tab:results} we can distinguish the fact that when PCA is used for feature extraction, the performance of GNB is lower than when other feature selection methods are used.
-The confidence interval of the predictive performance of GNB when PCA is used is 95\% $CI: 56.9 - 65.8$ which is significantly worse than other dimension reduction techniques.
+The confidence interval of the predictive performance of GNB when PCA is used is 95\% $CI: 56.9 - 65.8$ which is significantly worse than applying GNB in combination with other feature selection methods.
 This suggests that maximizing the variance of the data does not maximize information.

### 2012-11-11 article/article.tex
results

@@
 \hline
-\textbf{Extraction/Selection Method} & \textbf{Naive Bayes} & \textbf{Linear SVM} & \textbf{Naive Bayes*} & \textbf{Multivariate*} \\ \hline
+\textbf{Dimensionality Reduction Method} & \textbf{Naive Bayes} & \textbf{Linear SVM} & \textbf{Naive Bayes*} & \textbf{Multivariate*} \\ \hline
 PCA & 61.46 &  - & - & - \\

### 2012-11-11 article/article.tex
Proof-read discussion/conclusion

--- 2012-11-11 article/article.tex
@@
 The implementation of the algorithms was done using Python combined with modules tailored to mathematics (NumPy, SciPy) and machine learning (ScikitLearn).
-The original dataset was made available as a MatLab file and we used preprocessing in MatLab to create a number of files containing modified versions of this data, such as normalization and ROI grouping (see section~\ref{preprocessing}).
+The original dataset was made available as a MatLab file and we used preprocessing in MatLab to create a number of files containing modified versions of this data, such as normalization and ROI grouping (see section~\ref{ssec:preprocessing}).
 The source code for the algorithms is available from the authors on request.
@@
 To further reduce the number of features, average voxel values within these regions are taken as new input features.
-As other higher-order features such as the standard deviation, kurtosis and skewness within ROIs did not increase predictability, these were not included in the subsequent analysis.
+As with preprocessing, other higher-order features such as the standard deviation, kurtosis and skewness within ROIs did not increase predictability and were not included in the subsequent analysis.
 The disadvantage of using ROIs is that they require \textit{a priori} knowledge.
@@
 This means that the first scans after a stimulus may not represent activity related to this stimulus and should not be used for classification.
-To examine which scans provide maximal class information, we plotted the average BOLD activity per scan for all trials of all subjects, given a trial type and ROI.
+To examine which scans prWindows API error 87: The parameter is incorrect.ovide maximal class information, we plotted the average BOLD activity per scan for all trials of all subjects, given a trial type and ROI.
 As can be seen in Figure~\ref{fig:boldroi}, scans 10 to 20 provide maximal information for the first stimulus, while scans 23 to 33 provide maximal information for the second stimulus.
@@
 	\includegraphics[width=90mm]{figures/scan_hemodynamic_BOLD_roi}
-  	\caption{Haemodynamic response for ROI "CALC".
+  	\caption{Haemodynamic BOLD response for ROI "CALC".
 The red line depicts a sentence-picture trial (S-P) and the blue line depicts a picture-sentence trial (P-S).
-Average BOLD signal for all subjects is plotted against the scan number.
-The coloured regions depict the mean BOLD signal +/- 1 standard deviation.}
+Average Bold activity for all subjects is plotted against the scan number.
+The coloured regions depict an interval of 1 standard deviation.}
   	\label{fig:boldroi}
@@
 From Table~\ref{tab:results} we can distinguish the fact that when PCA is used for feature extraction, the performance of GNB is lower than when other feature selection methods are used.
-The confidence interval of the predictive performance of GNB when PCA is used is 95\% $CI: 56.9 - 65.8$ which is significantly worse than applying GNB in combination with other feature selection methods.
+The confidence interval of the predictive performance of GNB when PCA is used is 95\% $CI: 56.9 - 65.8$ which is significantly worse than applying GNB in combination with the other methods.
 This suggests that maximizing the variance of the data does not maximize information.
@@
 	\includegraphics[width=90mm]{figures/pca_plot_by_index.png}
-  	\caption{Classification accuracy of GNB after performing PCA with an increasing number of components.}
+  	\caption{Classification accuracy of GNB after performing PCA with a varying amount of components.}
   	\label{fig:pca_components}
@@
 This shows that machine learning algorithms can be employed to capture useful information from noisy data, and generalise this information between subjects.\\
-\indent We found that even though all of our classification procedures yielded well above chance-level classification performance, the differences between the classifiers was very seldom significant.
-Still, we consistently found that the Multivariate Gaussian (MVG) classifier outperformed the Gaussian Naive Bayes (GNB) classifier and Linear Support Vector machine (LSV).
+\indent We found that even though all of our classification procedures yielded well above chance-level classification performance, the difference between the classifiers was seldomly significant.
+Still, we consistently found that the Multivariate Gaussian (MVG) classifier outperformed the Gaussian Naive Bayes (GNB) classifier and Linear Support Vector Machine (LSVM).
 This finding suggests that the interactions between different brain areas contain information about the cognitive state, and that these functional connectivity generalises to an extent to different subjects.\\

### 2012-11-11 article/article.tex

@@
 {\small
-\begin{tabular}{c|c|c|c|c}
-& \textbf{Naive Bayes} & \textbf{Linear SVM} & \textbf{Naive Bayes*} & \textbf{Multivariate*} \\ \hline
+\begin{tabular}{|c|c|c|c|c|}
+\hline
+\textbf{Dimensionality Reduction Method} & \textbf{Naive Bayes} & \textbf{Linear SVM} & \textbf{Naive Bayes*} & \textbf{Multivariate*} \\ \hline
 PCA & 61.46 &  - & - & - \\

### 2012-11-11 article/article.tex
peel
Read-through, typos

@@
 \caption{Performance results (in \%) of the three classifiers on different methods for feature selection and structure.
-\emph{M} stands for \emph{Mean}, \emph{Norm} for \emph{Normalisation}, \emph{SD} for \emph{Standard Deviation}, and \emph{Norm/Vox} for \emph{Normalisation per Voxel index}.}
+\emph{M} stands for \emph{Mean}, \emph{Norm} for \emph{Normalisation}, \emph{SD} for \emph{Standard Deviation}, and \emph{Norm/Vox} for \emph{Normalisation per Voxel index}.
+Columns marked with an asterisk used the total log-likelihood as a decision function (see section~\ref{sec:trainingMethodsAndClassifiers}).}
 \label{tab:results}
@@
 
-Moreover, when different method for structuring the data is used the performance increases, though not significantly.
-This can be seen by comparing the performance of GNB when different method for structuring the data is used.
-In the first case it is 73.13\% while in the second case is 76.88\%.
+Moreover, when a different method for structuring the data is used the performance increases, though not significantly.
+This can be seen by comparing the performance of GNB on different datasets.
+In the first case performance is 73.13\% while in the second case it is 76.88\%.
 The confidence interval in the first case is 95\% $CI: 68.9 - 77.0$ and in the second case is 95\% $CI: 72.8 - 80.6$.
 
-Finally, we can distinguish the fact that MVG outperforms all the other classifiers, though we did not try a different data structure in the case of LSVM.
-However, we still do not expect that the performance of LSVM will be higher that this of MVG.
+Finally, we can distinguish the fact that MVG outperforms all the other classifiers, though we did not try a different decision function in the case of LSVM.
+However, we still do not expect that the performance of LSVM will be higher than that of MVG.
 The confidence interval of MVG is 95\% $CI: 75.7-83.1$.

### 2012-11-11 article/article.tex
results

--- 2012-11-11 article/article.tex
@@
 We applied PCA to the original data, while the rest feature selection methods were applied to the data belonging to the selected ROIs (cf.~Section~\ref{sec:dimensionalityReduction}).
-For instance, \emph{Norm/Vox-M-SD ROIs} indicates that we normalised over the voxel values of the same voxel index using the mean and standard deviation, while \emph{Norm-M-SD ROIs} indicates that we normalised over the voxel values of the same scan index using the mean and standard deviation.
+For instance, \emph{Norm/Vox-M ROIs} indicates that we normalised over the voxel values of the same voxel index using the mean and standard deviation, while \emph{Norm-M-SD ROIs} indicates that we normalised over the voxel values of the same scan index using the mean and standard deviation.
 Columns depict the used classifiers.
@@
 
-Furthermore, to test the significance of the differences in performance between the different methods we performed an exact binomial test in each case, from which we obtained the confidence interval.
-The results were rounded to one decimal point.
-For the rest of this section and to explain some of the results we are looking only at the case when \emph{Norm/Vox-M-SD ROIs} is used as a feature selection method.
+Furthermore, to test the significance of the differences in performance between the different methods we performed a Chi-square test for interesting contrasts, and obtained a confidence interval for their difference in proportions.
+For the rest of this section we focus only on the case when \emph{Norm/Vox-M ROIs} is used as a feature selection method.
 
 From Table~\ref{tab:results} we can distinguish the fact that when PCA is used for feature extraction, the performance of GNB is lower than when other feature selection methods are used.
-The confidence interval of the predictive performance of GNB when PCA is used is 95\% $CI: 56.9 - 65.8$ which is significantly worse than applying GNB in combination with the other methods.
-This suggests that maximizing the variance of the data does not maximize information.
-Another explanation could be that the data is not suited to be aligned by voxel index (see section~\ref{sec:dimensionalityReduction}).
+By using an exact binomial test we found that GNB that used PCA as feature extraction method performed significantly better than chance-level, with a 95\% $CI: 56.9 - 65.8$, $p < 0.001$.
+However, a Chi-square test for equality of proportions indicated that using PCA as dimentionality reduction method performed significantly worse than a GNB classifier on the means of the selected ROI's, $\chi^2(1) = 7.7848, p < 0.01$.
+This suggests that maximizing the variance of the data does not maximize information, possibly because the low signal-to-noise ratio of fMRI data does not insure that the principle component with the highest variance also contains a lot of useful information.
+Additionally, the data may not suited to be aligned by voxel index (see section~\ref{sec:dimensionalityReduction}).
 The performance listed in Table~\ref{tab:results} is for PCA using 149 components.
-This value was found by performing classification with a varying amount of components, the results for which are listed in figure~\ref{fig:pca_components}.
+This value was found by performing classification with a varying amount of components, the results for which are listed in Figure~\ref{fig:pca_components}.
 
@@
 Comparing classifiers when the same data structure is used, we can distinguish the fact that MVG performs better than GNB.
-Also that LSVM performs better than GNB, though not significantly.
-The confidence interval of GNB is 95\% $CI: 68.9-77.0$ and of LSVM is 95\% $CI: 72.6-80.4$, which suggests that their performances do not differ significantly.
+However, a Chi-square test did not find this difference to be significant, $\chi^2(1) = 0.8807, p = 0.348$.
+Furthermore, the difference between LSVM and GNB was also not found to be significant, $\chi^2(1) = 1.4183, p = 0.234$.
 
-Moreover, when different method for structuring the data is used the performance increases, though not significantly.
-This can be seen by comparing the performance of GNB when different method for structuring the data is used.
-In the first case it is 73.13\% while in the second case is 76.88\%.
-The confidence interval in the first case is 95\% $CI: 68.9 - 77.0$ and in the second case is 95\% $CI: 72.8 - 80.6$.
-
-Finally, we can distinguish the fact that MVG outperforms all the other classifiers, though we did not try a different data structure in the case of LSVM.
-However, we still do not expect that the performance of LSVM will be higher that this of MVG.
-The confidence interval of MVG is 95\% $CI: 75.7-83.1$.
+Finally, when classifying per scan and combining these classifications into a single classification per trial (cf.~Table~\ref{tab:results} indicated with a *) the performance increases.
+When looking at the increase in performance for the GNB classifier, we see performance increases from 73.13\% to 76.88\%.
+However, a Chi-square test did not find this increase in performance to be significant, $\chi^2(1) = 1.6056, p = 0.205$.
 

### 2012-11-11 src/demo/gaussian_nb.py

@@ -153,5 +153,7 @@ for datafile in FILES:
             total_correct += correct
             total_scans += count
 
+    print total_correct
+    print total_scans
     print "Final result: %s%%" % ((total_correct / total_scans) * 100)
     print

### 2012-11-11 article/article.tex

@@
 
+\begin{figure}
+	\centering
+	\includegraphics[width=0.45\textwidth]{figures/flying_dutchman}
+  	\caption{}
+  	\label{fig:bold_response}
+\end{figure}
+
 \textbf{\emph{Principal Component Analysis (PCA).}} As feature extraction method, we applied PCA.

### 2012-11-11 article/article.tex
results

@@
 \emph{M} stands for \emph{Mean}, \emph{Norm} for \emph{Normalisation}, \emph{SD} for \emph{Standard Deviation}, and \emph{Norm/Vox} for \emph{Normalisation per Voxel index}.
-Columns marked with an asterisk used the total log-likelihood as a decision function (see section~\ref{sec:trainingMethodsAndClassifiers}).}
+Columns marked with an asterisk used the total log-likelihood as a decision function (cf.~Section~\ref{sec:trainingMethodsAndClassifiers}).}
 \label{tab:results}
@@
 This suggests that maximizing the variance of the data does not maximize information, possibly because the low signal-to-noise ratio of fMRI data does not insure that the principle component with the highest variance also contains a lot of useful information.
-Additionally, the data may not suited to be aligned by voxel index (see section~\ref{sec:dimensionalityReduction}).
+Additionally, the data may not suited to be aligned by voxel index (cf.~Section~\ref{sec:dimensionalityReduction}).
 The performance listed in Table~\ref{tab:results} is for PCA using 149 components.

### 2012-11-11 article/article.tex
picture

--- 2012-11-11 article/article.tex
@@
 By using this technique, voxels or brain areas can be identified that differ in their average BOLD response across conditions.\\
-\begin{figure}
-	\centering
-	\includegraphics[width=60mm]{figures/f1_bold_response}
-  	\caption{Above is the hypothetical neural response to some stimulus.
-Below is the BOLD response that is evoked by the short spike of neural activity.
-The neural activity shown above is attempted to be inferred from the BOLD response below that is measured by the fMRI scanner.
-Figure is adopted from Heeger and Ress \cite{he:2002na}.}
-  	\label{fig:bold_response}
-\end{figure}
 \indent Recent studies indicate that this approach may have some serious limitations.
@@
 The classifiers that we compare in this paper are a Gaussian Naive Bayesian (GNB) classifier, Linear Support Vector Machine (LSVM), and a Multivariate Gaussian (MVG) classifier.\\
+\begin{figure}
+	\centering
+	\includegraphics[width=60mm]{figures/f1_bold_response}
+  	\caption{Above is the hypothetical neural response to some stimulus.
+Below is the BOLD response that is evoked by the short spike of neural activity.
+The neural activity shown above is attempted to be inferred from the BOLD response below that is measured by the fMRI scanner.
+Figure is adopted from Heeger and Ress \cite{he:2002na}.}
+  	\label{fig:bold_response}
+\end{figure}
 \indent The goal of this study is to compare the classification performance of the cognitive states associated with looking at a picture or looking at a sentence between subjects.
@@
 The implementation of the algorithms was done using Python combined with modules tailored to mathematics (NumPy, SciPy) and machine learning (ScikitLearn).
-The original dataset was made available as a MatLab file and we used preprocessing in MatLab to create a number of files containing modified versions of this data, such as normalization and ROI grouping (see section~\ref{ssec:preprocessing}).
+The original dataset was made available as a MatLab file and we used preprocessing in MatLab to create a number of files containing modified versions of this data, such as normalization and ROI grouping (see section~\ref{preprocessing}).
 The source code for the algorithms is available from the authors on request.
@@
 To further reduce the number of features, average voxel values within these regions are taken as new input features.
-As with preprocessing, other higher-order features such as the standard deviation, kurtosis and skewness within ROIs did not increase predictability and were not included in the subsequent analysis.
+As other higher-order features such as the standard deviation, kurtosis and skewness within ROIs did not increase predictability, these were not included in the subsequent analysis.
 The disadvantage of using ROIs is that they require \textit{a priori} knowledge.
@@
 This means that the first scans after a stimulus may not represent activity related to this stimulus and should not be used for classification.
-To examine which scans prWindows API error 87: The parameter is incorrect.ovide maximal class information, we plotted the average BOLD activity per scan for all trials of all subjects, given a trial type and ROI.
+To examine which scans provide maximal class information, we plotted the average BOLD activity per scan for all trials of all subjects, given a trial type and ROI.
 As can be seen in Figure~\ref{fig:boldroi}, scans 10 to 20 provide maximal information for the first stimulus, while scans 23 to 33 provide maximal information for the second stimulus.
@@
 We applied PCA to the original data, while the rest feature selection methods were applied to the data belonging to the selected ROIs (cf.~Section~\ref{sec:dimensionalityReduction}).
-For instance, \emph{Norm/Vox-M ROIs} indicates that we normalised over the voxel values of the same voxel index using the mean and standard deviation, while \emph{Norm-M-SD ROIs} indicates that we normalised over the voxel values of the same scan index using the mean and standard deviation.
+For instance, \emph{Norm/Vox-M-SD ROIs} indicates that we normalised over the voxel values of the same voxel index using the mean and standard deviation, while \emph{Norm-M-SD ROIs} indicates that we normalised over the voxel values of the same scan index using the mean and standard deviation.
 Columns depict the used classifiers.
@@
 
-Furthermore, to test the significance of the differences in performance between the different methods we performed a Chi-square test for interesting contrasts, and obtained a confidence interval for their difference in proportions.
-For the rest of this section we focus only on the case when \emph{Norm/Vox-M ROIs} is used as a feature selection method.
+Furthermore, to test the significance of the differences in performance between the different methods we performed an exact binomial test in each case, from which we obtained the confidence interval.
+The results were rounded to one decimal point.
+For the rest of this section and to explain some of the results we are looking only at the case when \emph{Norm/Vox-M-SD ROIs} is used as a feature selection method.
 
 From Table~\ref{tab:results} we can distinguish the fact that when PCA is used for feature extraction, the performance of GNB is lower than when other feature selection methods are used.
-By using an exact binomial test we found that GNB that used PCA as feature extraction method performed significantly better than chance-level, with a 95\% $CI: 56.9 - 65.8$, $p < 0.001$.
-However, a Chi-square test for equality of proportions indicated that using PCA as dimentionality reduction method performed significantly worse than a GNB classifier on the means of the selected ROI's, $\chi^2(1) = 7.7848, p < 0.01$.
-This suggests that maximizing the variance of the data does not maximize information, possibly because the low signal-to-noise ratio of fMRI data does not insure that the principle component with the highest variance also contains a lot of useful information.
-Additionally, the data may not suited to be aligned by voxel index (see section~\ref{sec:dimensionalityReduction}).
+The confidence interval of the predictive performance of GNB when PCA is used is 95\% $CI: 56.9 - 65.8$ which is significantly worse than applying GNB in combination with other feature selection methods.
+This suggests that maximizing the variance of the data does not maximize information.
+Another explanation could be that the data is not suited to be aligned by voxel index (see section~\ref{sec:dimensionalityReduction}).
 The performance listed in Table~\ref{tab:results} is for PCA using 149 components.
-This value was found by performing classification with a varying amount of components, the results for which are listed in Figure~\ref{fig:pca_components}.
+This value was found by performing classification with a varying amount of components, the results for which are listed in figure~\ref{fig:pca_components}.
 
@@
 Comparing classifiers when the same data structure is used, we can distinguish the fact that MVG performs better than GNB.
-However, a Chi-square test did not find this difference to be significant, $\chi^2(1) = 0.8807, p = 0.348$.
-Furthermore, the difference between LSVM and GNB was also not found to be significant, $\chi^2(1) = 1.4183, p = 0.234$.
+Also that LSVM performs better than GNB, though not significantly.
+The confidence interval of GNB is 95\% $CI: 68.9-77.0$ and of LSVM is 95\% $CI: 72.6-80.4$, which suggests that their performances do not differ significantly.
 
-Finally, when classifying per scan and combining these classifications into a single classification per trial (cf.~Table~\ref{tab:results} indicated with a *) the performance increases.
-When looking at the increase in performance for the GNB classifier, we see performance increases from 73.13\% to 76.88\%.
-However, a Chi-square test did not find this increase in performance to be significant, $\chi^2(1) = 1.6056, p = 0.205$.
+Moreover, when different method for structuring the data is used the performance increases, though not significantly.
+This can be seen by comparing the performance of GNB when different method for structuring the data is used.
+In the first case it is 73.13\% while in the second case is 76.88\%.
+The confidence interval in the first case is 95\% $CI: 68.9 - 77.0$ and in the second case is 95\% $CI: 72.8 - 80.6$.
+
+Finally, we can distinguish the fact that MVG outperforms all the other classifiers, though we did not try a different data structure in the case of LSVM.
+However, we still do not expect that the performance of LSVM will be higher that this of MVG.
+The confidence interval of MVG is 95\% $CI: 75.7-83.1$.
 
@@
 This shows that machine learning algorithms can be employed to capture useful information from noisy data, and generalise this information between subjects.\\
-\indent We found that even though all of our classification procedures yielded well above chance-level classification performance, the difference between the classifiers was seldomly significant.
-Still, we consistently found that the Multivariate Gaussian (MVG) classifier outperformed the Gaussian Naive Bayes (GNB) classifier and Linear Support Vector Machine (LSVM).
+\indent We found that even though all of our classification procedures yielded well above chance-level classification performance, the differences between the classifiers was very seldom significant.
+Still, we consistently found that the Multivariate Gaussian (MVG) classifier outperformed the Gaussian Naive Bayes (GNB) classifier and Linear Support Vector machine (LSV).
 This finding suggests that the interactions between different brain areas contain information about the cognitive state, and that these functional connectivity generalises to an extent to different subjects.\\

### 2012-11-11 article/article.tex
yes

--- 2012-11-11 article/article.tex
@@
 
+\begin{figure}
+	\centering
+	\includegraphics[width=0.45\textwidth]{figures/flying_dutchman}
+  	\caption{ROIs.
+These ROIs have been selected as input for the classifiers.
+From top left to bottom right, descending vertical slices of the brain are depicted.
+brown = LDLPFC, red = LIPS, green = LOPER, yellow = LTRIA, orange = LT, light blue = LIPL, blue = CALC}
+  	\label{fig:fd}
+\end{figure}
+
 \textbf{\emph{ROI mapping.}} ROI mapping is a form of feature selection, wherein only a subset of the N-dimensional input space is used.
 As described in the introduction, ROIs are anatomically defined brain areas classified as relevant by an expert in the field.
-We used 7 of the 25 defined ROIs, labeled  'CALC' 'LIPL' 'LT' 'LTRIA' 'LOPER' 'LIPS' and  'LDLPFC'.
+We used 7 of the 25 defined ROIs, labeled  'CALC' 'LIPL' 'LT' 'LTRIA' 'LOPER' 'LIPS' and  'LDLPFC' (see Figure~\ref{fig:fd}).
 These ROIs have previously been found to be most relevant in discriminating between the cognitive states of looking at sentences or pictures \cite{wa:2003betw}.
@@
 When this knowledge is inaccurate, classification accuracy may not be optimal.
-
-\begin{figure}
-	\centering
-	\includegraphics[width=0.45\textwidth]{figures/flying_dutchman}
-  	\caption{}
-  	\label{fig:bold_response}
-\end{figure}
 

### 2012-11-11 article/article.tex
yes

@@
 Rows indicate various feature selection methods.
-We applied PCA to the original data, while the rest feature selection methods were applied to the data belonging to the selected ROIs (cf.~Section~\ref{sec:dimensionalityReduction}).
+We applied PCA to the original data, while for the rest feature selection methods were applied to the data belonging to the selected ROIs (cf.~Section~\ref{sec:dimensionalityReduction}).
 For instance, \emph{Norm/Vox-M ROIs} indicates that we normalised over the voxel values of the same voxel index using the mean and standard deviation, while \emph{Norm-M-SD ROIs} indicates that we normalised over the voxel values of the same scan index using the mean and standard deviation.
@@
 This suggests that maximizing the variance of the data does not maximize information, possibly because the low signal-to-noise ratio of fMRI data does not insure that the principle component with the highest variance also contains a lot of useful information.
-Additionally, the data may not suited to be aligned by voxel index (cf.~Section~\ref{sec:dimensionalityReduction}).
+Additionally, the data may not be well suited to be aligned by voxel index (cf.~Section~\ref{sec:dimensionalityReduction}).
 The performance listed in Table~\ref{tab:results} is for PCA using 149 components.

### 2012-11-11 article/article.tex
yes

@@
 In \emph{pattern characterization}, the question of \emph{how} class information is represented takes a central place.
-While this paper will focus on linear classifiers, some studies in this area resort to a more sophisticated approach and revolve around the question how to correspond relationships in stumulus with those in fMRI data.
+While this paper will focus on linear classifiers, some studies in this area resort to a more sophisticated approach and revolve around the question how to correspond relationships in stimulus with those in fMRI data.
 For example,~\cite{mitchell2008predicting} researches predicting fMRI response for new words based on their similarity to words for which the response is known.

### 2012-11-12 article/article.tex
results

@@
 Specifically, by observing the patterns of activity for a group of subjects, we are able to predict with above-chance accuracy what the cognitive state of an unseen subject is.
-This shows that machine learning algorithms can be employed to capture useful information from noisy data, and generalise this information between subjects.\\
-\indent We found that even though all of our classification procedures yielded well above chance-level classification performance, the difference between the classifiers was seldomly significant.
+This shows that machine learning algorithms can be employed to capture useful information from noisy data, and generalise this information between subjects.
+
+Wang et al.
+\cite{wa:2003betw} used GNB and Support Vector Machine (SVM) as cross-subject classifiers in a sentence versus picture study, in a similar context as our conducted study.
+When using the averaging technique (averaging all voxels in an ROI into a supervoxel) and normalization, GNB yielded a predictive performance of 74.3\%, while SVM 75.3\%.
+
+We found that even though all of our classification procedures yielded well above chance-level classification performance, the difference between the classifiers was seldomly significant.
 Still, we consistently found that the Multivariate Gaussian (MVG) classifier outperformed the Gaussian Naive Bayes (GNB) classifier and Linear Support Vector Machine (LSVM).
-This finding suggests that the interactions between different brain areas contain information about the cognitive state, and that these functional connectivity generalises to an extent to different subjects.\\
-\indent Still, the uncertainty in our predictions are relatively high, with maximum classification performance reaching close to 80\% accuracy for only two classes.
+This finding suggests that the interactions between different brain areas contain information about the cognitive state, and that these functional connectivity generalises to an extent to different subjects.
+
+Still, the uncertainty in our predictions are relatively high, with maximum classification performance reaching close to 80\% accuracy for only two classes.
 The noisy nature of the fMRI data therefore makes it very difficult to draw hard conclusions when decoding a single cognitive state from a brain scan.
-This is something to be considered before employing fMRI as a diagnostic tool, or as evidence in court.\\
-\indent Overall we can say that machine learning provides a feasible way of extracting useful information from fMRI data.
+This is something to be considered before employing fMRI as a diagnostic tool, or as evidence in court.
+
+Overall we can say that machine learning provides a feasible way of extracting useful information from fMRI data.
 Furthermore, it avoids a strict multiple comparisons correction that is required in null hypothesis testing by using cross-validation.

### 2012-11-12 article/article.tex
results

@@
 
-Wang et al.
-\cite{wa:2003betw} used GNB and Support Vector Machine (SVM) as cross-subject classifiers in a sentence versus picture study, in a similar context as our conducted study.
-When using the averaging technique (averaging all voxels in an ROI into a supervoxel) and normalization, GNB yielded a predictive performance of 74.3\%, while SVM 75.3\%.
+To compare our results with other related studies, Wang et al.
+\cite{wa:2003betw} used GNB and Support Vector Machine (SVM) as cross-subject classifiers in a sentence versus picture study.
+When using the averaging technique (averaging all voxels in an ROI into a supervoxel) and normalization, GNB yielded a predictive performance of 74.3\%, and SVM 75.3\%.
+Our results denote a higher performance, where using normalization GNB yielded a performance of 76.67\% and LSVM 76.88\%.
 

### 2012-11-12 article/article.tex
review

@@
 In this paper, we present a comparison of the performance of machine learning algorithms in between-subjects classification of cognitive states, using fMRI data.
-We found that averaging across ROIs gave the best classification performance.
+We found that averaging across Regions of Interest (ROIs) gave the best classification performance.
 All classifiers that we used performed above-chance, and we found that a Multivariate Gaussian classifier outperformed a Naive Bayesian classifier and a Support Vector Machine.
@@
 
-Traditionally, data from each voxel, or from a group of voxels, are analysed with analysis of variance.
-Voxels are usually convoluted with a haemodynamic response function in an attempt to infer the spiking neural activity from the slowly building and dropping BOLD response \cite{he:2002na} (see Figure~\ref{fig:bold_response}).
+Traditionally, data from each voxel, or from a group of voxels, are analyzed with analysis of variance.
+Voxels are usually convoluted with a haemodynamic response function in an attempt to infer the spiking neural activity from the slowly building and dropping BOLD response \cite{he:2002na} (cf.~Figure~\ref{fig:bold_response}).
 In this setup, experimental conditions are the independent variables and a general linear model is fitted to each voxel, or group of voxels.
-By using this technique, voxels or brain areas can be identified that differ in their average BOLD response across conditions.\\
-\indent Recent studies indicate that this approach may have some serious limitations.
+By using this technique, voxels or brain areas can be identified that differ in their average BOLD response across conditions.
+
+Recent studies indicate that this approach may have some serious limitations.
 One study used a dataset of 1326 subjects, which is much higher than the usual 15 to 25 subjects found in fMRI studies \cite{thy:2012very}.
 This study explored what happened to a simple contrast between BOLD signal while watching faces and baseline activity, when increasing the number of subjects included in the analysis.
-They found that, as the number of analysed subjects increased, an increasing number of brain areas that extended well beyond the expected visual- and face recognition areas showed significant deviation from baseline.
+They found that, as the number of analyzed subjects increased, an increasing number of brain areas that extended well beyond the expected visual- and face recognition areas showed significant deviation from baseline.
 Another study, using only three subjects, repeated a simple letter/number discrimination task and visual stimulation.
@@
 These findings are perhaps not surprising, because everything we know about the brain indicates that it is best seen as a network with countless connections to other parts of the brain.
-Therefore, it makes sense to also analyse signals from the brain as a pattern of information coming from the entire brain, rather than attempting to link specific brain areas to experimental conditions.\\
-\indent In fact, Machine learning algorithms have been employed in studies to learn the patterns of activity associated with some experimental condition.
+Therefore, it makes sense to also analyze signals from the brain as a pattern of information coming from the entire brain, rather than attempting to link specific brain areas to experimental conditions.
+
+In fact, Machine learning algorithms have been employed in studies to learn the patterns of activity associated with some experimental condition.
 Subsequently, these learned patterns can be used on unseen subjects in order to predict experimental conditions from their brain activity patterns.
-If these predictions are accurate then it means that the identified patterns hold information about the brain activity associated with some experimental task.\\
-\indent The general goal of these machine learning studies is to identify what experimental condition a subject is in at a certain time.
+If these predictions are accurate then it means that the identified patterns hold information about the brain activity associated with some experimental task.
+
+The general goal of these machine learning studies is to identify what experimental condition a subject is in at a certain time.
 The state that a person is in during some trial (e.g., looking at a face) is referred to as \emph{cognitive state}, because the cognitive processes that are associated with being in that state are of interest.
@@
 This \emph{StarPlus} experiment has been performed by researchers where, in addition to viewing pictures, subjects also were tasked to read sentences containing some true or false statement about the pictures.
-These data have then been used to successfully infer the cognitive states of looking at either sentences or pictures within single subjects \cite{mi:2003within}, and between subjects \cite{wa:2003betw}.\\
-\indent In this study we use the same StarPlus data as in the aforementioned experiment, and use machine learning algorithms to decode the cognitive states of subjects while looking at pictures or while looking at sentences.
+These data have then been used to successfully infer the cognitive states of looking at either sentences or pictures within single subjects \cite{mi:2003within}, and between subjects \cite{wa:2003betw}.
+
+In this study we use the same StarPlus data as in the aforementioned experiment, and use machine learning algorithms to decode the cognitive states of subjects while looking at pictures or while looking at sentences.
 An interesting difference between many applications of machine learning, and this study is that most machine learning algorithms have the goal of achieving the maximum possible discrimination between conditions or classes.
@@
 To avoid this problem it is recommended to only use linear classifiers when decoding cognitive states from fMRI data.
-The classifiers that we compare in this paper are a Gaussian Naive Bayesian (GNB) classifier, Linear Support Vector Machine (LSVM), and a Multivariate Gaussian (MVG) classifier.\\
+The classifiers that we compare in this paper are a Gaussian Naive Bayesian (GNB) classifier, Linear Support Vector Machine (LSVM), and a Multivariate Gaussian (MVG) classifier.
+
 \begin{figure}
@@
 \end{figure}
-\indent The goal of this study is to compare the classification performance of the cognitive states associated with looking at a picture or looking at a sentence between subjects.
+
+The goal of this study is to compare the classification performance of the cognitive states associated with looking at a picture or looking at a sentence between subjects.
 This objective introduces an important issue, namely, how the data from the fMRI scans can be compared between subjects.
@@
 By using the Talairach coordinate system the data of each individual voxel can be used~\cite{talairach1988co}.
-However, because the error of doing this is usually in the order of a centimetre, while the size of the voxels are in the order of millimetres, the coordinates are difficult to generalise.
-Because we aim to do between-subjects comparisons the generalisation performance is important, thus we will use averaged ROIs in order to compare the classifiers \cite{mi:2004coord}.
+However, because the error of doing this is usually in the order of a centimetre, while the size of the voxels are in the order of millimetres, the coordinates are difficult to generalize.
+Because we aim to do between-subjects comparisons the generalization performance is important, thus we will use averaged ROIs in order to compare the classifiers \cite{mi:2004coord}.
 

### 2012-11-12 article/article.tex
review

@@
 \label{sec:application}
+
 Thus far, machine learning techniques have been successfully applied for decoding the orientation of a stimulus from fMRI scans \cite{ka:2005decoding}, identifying different patterns of activity for animate versus inanimate objects \cite{kr:2008RSA}, and even reconstructing visual input based on information in the fMRI scans \cite{mi:2008re}.
@@
 
-The studies mentioned before all depend on the question of whether fMRI data carry class
-information (pattern discrimination) and exploiting this information to perform classification.
-However, researchers have indicated at least two other basic questions to which classifiers can be applied, being ``where or when
-is class information represented?'' (pattern localization) and ``how is
-class information encoded'' or ``how does its encoding relate to known
-relationships between stimuli'' (pattern characterization)~\cite{pereira2009machine}.
+The studies mentioned before all depend on the question of whether fMRI data carries class information (pattern discrimination) and exploiting this information to perform classification.
+However, researchers have indicated at least two other basic questions to which classifiers can be applied, being ``where or when is class information represented?'' (pattern localization) and ``how is class information encoded'' or ``how does its encoding relate to known relationships between stimuli'' (pattern characterization)~\cite{pereira2009machine}.
 While the main focus of this paper is on the application of machine learning techniques in pattern discrimination, we will briefly discuss these two other applications.
@@
 As a result, activation patterns often do not line up well.
-One of the strategies to deal with this is to use the set of voxels active for each subject as seeds in a clustering algorithm to find similar voxels.\\
+One of the strategies to deal with this is to use the set of voxels active for each subject as seeds in a clustering algorithm to find similar voxels.
 
@@
 \label{sec:methods}
-We now continue by discussing the methods applied in our research. The general approach taken for each of the algorithms discussed consists of the following steps:
+
+We now continue by discussing the methods applied in our research. The general approach taken for each of the discussed algorithms consists of the following steps:
+
 \begin{enumerate}
@@
 \end{enumerate}
+
 Preprocessing was done to transform the dataset into a format suitable for input into the classification algorithm.
@@
 Finally, the approach taken to training and testing the classifier deals with preventing overfitting the classifier to training data and verifying accuracy for cross-subject classification.
-The choices made in each of these steps may affect classifier accuracy, so to obain a complete understanding of classification accuracies we have tried several methods at each step.
-In the following sections, these methods will be described in more detail.\\
+The choices made in each of these steps may affect classifier accuracy, so to obtain a complete understanding of classification accuracies we have tried several methods at each step.
+In the following sections, these methods will be described in more detail.
+
 The implementation of the algorithms was done using Python combined with modules tailored to mathematics (NumPy, SciPy) and machine learning (ScikitLearn).
-The original dataset was made available as a MatLab file and we used preprocessing in MatLab to create a number of files containing modified versions of this data, such as normalization and ROI grouping (see section~\ref{ssec:preprocessing}).
+The original dataset was made available as a MatLab file and we used preprocessing in MatLab to create a number of files containing modified versions of this data, such as normalization and ROI grouping (cf.~Section~\ref{sec:preprocessing}).
 The source code for the algorithms is available from the authors on request.
@@
 \subsection{Data Acquisition and Dimensions}
+
 The dataset used for our experiments was originally collected by Marcel Just and colleagues from Carnegie Mellon University's Center for Cognitive Brain Imaging and published as the StarPlus fMRI dataset.
@@
 \subsection{Preprocessing}
-\label{ssec:preprocessing}
+\label{sec:preprocessing}
 The dataset contains activation values for each voxels with respect to a baseline activation value for a subject, measured before performing the task.

### 2012-11-12 article/article.tex
review

@@
 
-The dataset used for our experiments was originally collected by Marcel Just and colleagues from Carnegie Mellon University's Center for Cognitive Brain Imaging and published as the StarPlus fMRI dataset.
+The dataset used for our experiments was originally collected by Marcel Just et al.
+from Carnegie Mellon University's Center for Cognitive Brain Imaging and published as the StarPlus fMRI dataset.
 This set consists of fMRI data for 6 subjects that was obtained while having each subject participate in a number of trials.
@@
 In total, 14 trials were baseline trials, while half of the remaining trials had a picture as first stimulus and the other half a sentence.
+
 The entire dataset consists of voxel data and meta-data providing information about the dataset.
@@
 In total, we have data available for approximately 4800 voxels per scan, 54 scans per trial, 40 trials per subject (excluding baseline trials) and 6 subjects.
-Voxels have been grouped into 25 ROIs (see \textit{Feature selection}).
+Voxels have been grouped into 25 ROIs (cf.~Section~\ref{sec:dimensionalityReduction}).
 
@@
 \label{sec:preprocessing}
+
 The dataset contains activation values for each voxels with respect to a baseline activation value for a subject, measured before performing the task.

### 2012-11-12 article/article.tex
review

@@
 \label{sec:introduction}
+
 \maketitle
+
 %% Now the actual contents of first section
@@
 Traditionally, data from each voxel, or from a group of voxels, are analyzed with analysis of variance.
-Voxels are usually convoluted with a haemodynamic response function in an attempt to infer the spiking neural activity from the slowly building and dropping BOLD response \cite{he:2002na} (cf.~Figure~\ref{fig:bold_response}).
+Voxels are usually convoluted with a hemodynamic response function in an attempt to infer the spiking neural activity from the slowly building and dropping BOLD response \cite{he:2002na} (cf.~Figure~\ref{fig:bold_response}).
 In this setup, experimental conditions are the independent variables and a general linear model is fitted to each voxel, or group of voxels.
@@
 %Useful background info: http://www.brainvoyager.com/bvqx/doc/UsersGuide/WebHelp/Content/MVPATools/MVPA_Basic_Concepts.htm
+
 Typical input for the machine learning algorithms discussed in this paper consists of a number of feature vectors, one for each fMRI scan.
@@
 In the most basic case, we could use the entire set of voxels belonging to a scan, resulting in a feature vector containing potentially (hundreds of) thousands of features.
+
 In the case that we have a very large dimension of features combined with only a limited amount of scans, our classifier may have difficulty learning the right classification function.
-This problem is formally known as the \textit{Hughes effect}\cite{Hughes1054102} which states that the predictive power of a machine learning algorithm decreases as the dimensionality increases with a fixed number of training examples.
+This problem is formally known as the \emph{Hughes effect}\cite{Hughes1054102} which states that the predictive power of a machine learning algorithm decreases as the dimensionality increases with a fixed number of training examples.
 To reduce the dimensionality of the input, several different methods were used.
@@
 \textbf{\emph{ROI mapping.}} ROI mapping is a form of feature selection, wherein only a subset of the N-dimensional input space is used.
-As described in the introduction, ROIs are anatomically defined brain areas classified as relevant by an expert in the field.
-We used 7 of the 25 defined ROIs, labeled  'CALC' 'LIPL' 'LT' 'LTRIA' 'LOPER' 'LIPS' and  'LDLPFC' (see Figure~\ref{fig:fd}).
+As described in the Section~\ref{sec:introduction}, ROIs are anatomically defined brain areas classified as relevant by an expert in the field.
+We used 7 of the 25 defined ROIs, labeled  'CALC' 'LIPL' 'LT' 'LTRIA' 'LOPER' 'LIPS' and  'LDLPFC' (cf.~Figure~\ref{fig:fd}).
 These ROIs have previously been found to be most relevant in discriminating between the cognitive states of looking at sentences or pictures \cite{wa:2003betw}.
 To further reduce the number of features, average voxel values within these regions are taken as new input features.
-As with preprocessing, other higher-order features such as the standard deviation, kurtosis and skewness within ROIs did not increase predictability and were not included in the subsequent analysis.
-The disadvantage of using ROIs is that they require \textit{a priori} knowledge.
+As with preprocessing, other higher-order features such as the standard deviation, kurtosis, and skewness within ROIs did not increase predictability and were not included in the subsequent analysis.
+The disadvantage of using ROIs is that they require \emph{a priori} knowledge.
 When this knowledge is inaccurate, classification accuracy may not be optimal.
@@
 
-\textbf{\emph{Temporal dimensionality reduction.}} The BOLD response peaks at approximately 5 seconds after a stimulus is presented (see Fig.
-1).
+\textbf{\emph{Temporal dimensionality reduction.}} The BOLD response peaks at approximately 5 seconds after a stimulus is presented (cf.~Figure~\ref{fig:bold_response}).
 This means that the first scans after a stimulus may not represent activity related to this stimulus and should not be used for classification.
 To examine which scans provide maximal class information, we plotted the average BOLD activity per scan for all trials of all subjects, given a trial type and ROI.
-As can be seen in Figure~\ref{fig:boldroi}, scans 10 to 20 provide maximal information for the first stimulus, while scans 23 to 33 provide maximal information for the second stimulus.
+As illustrated in Figure~\ref{fig:boldroi}, scans 10 to 20 provide maximal information for the first stimulus, while scans 23 to 33 provide maximal information for the second stimulus.
 The other scans were excluded.
+
 \begin{figure}
@@
 
-where $C_i$ denotes the probability of the class---sentence or picture.
+where $C_i$ denotes the probability of the class (picture or sentence).
 Two approaches were used to estimate the parameters of the model.
@@
 \textbf{\emph{Linear Support Vector Machine (LSVM).}} The LSVM is a discriminant based approach to classification.
-For an elaborate description see \cite{bk:2004aplay}.
+For an elaborate description see Chapter 10 of \cite{bk:2004aplay}.
 For each selected scan belonging to a stimulus, the LSVM classifies it as either a picture or a sentence.
@@
 The log-probability of a class becomes:
+
 \begin{dmath}

### 2012-11-12 article/article.tex
review

@@
 We applied PCA to the original data, while for the rest feature selection methods were applied to the data belonging to the selected ROIs (cf.~Section~\ref{sec:dimensionalityReduction}).
-For instance, \emph{Norm/Vox-M ROIs} indicates that we normalised over the voxel values of the same voxel index using the mean and standard deviation, while \emph{Norm-M-SD ROIs} indicates that we normalised over the voxel values of the same scan index using the mean and standard deviation.
+For instance, \emph{Norm/Vox-M ROIs} indicates that we normalized over the voxel values of the same voxel index using the mean and standard deviation, while \emph{Norm-M-SD ROIs} indicates that we normalized over the voxel values of the same scan index using the mean and standard deviation.
 Columns depict the used classifiers.
@@
 By using an exact binomial test we found that GNB that used PCA as feature extraction method performed significantly better than chance-level, with a 95\% $CI: 56.9 - 65.8$, $p < 0.001$.
-However, a Chi-square test for equality of proportions indicated that using PCA as dimentionality reduction method performed significantly worse than a GNB classifier on the means of the selected ROI's, $\chi^2(1) = 7.7848, p < 0.01$.
+However, a Chi-square test for equality of proportions indicated that using PCA as dimensionality reduction method performed significantly worse than a GNB classifier on the means of the selected ROIs, $\chi^2(1) = 7.7848, p < 0.01$.
 This suggests that maximizing the variance of the data does not maximize information, possibly because the low signal-to-noise ratio of fMRI data does not insure that the principle component with the highest variance also contains a lot of useful information.

### 2012-11-12 article/article.tex
nonono

--- 2012-11-12 article/article.tex
@@
 Still, we consistently found that the Multivariate Gaussian (MVG) classifier outperformed the Gaussian Naive Bayes (GNB) classifier and Linear Support Vector Machine (LSVM).
-This finding suggests that the interactions between different brain areas contain information about the cognitive state, and that these functional connectivity generalises to an extent to different subjects.
+This finding suggests that the interactions between different brain areas contain information about the cognitive state, and that this functional connectivity generalises to an extent to different subjects.
 

### 2012-11-12 article/article.tex
review

@@
 Specifically, by observing the patterns of activity for a group of subjects, we are able to predict with above-chance accuracy what the cognitive state of an unseen subject is.
-This shows that machine learning algorithms can be employed to capture useful information from noisy data, and generalise this information between subjects.
+This shows that machine learning algorithms can be employed to capture useful information from noisy data, and generalize this information between subjects.
 
@@
 Still, we consistently found that the Multivariate Gaussian (MVG) classifier outperformed the Gaussian Naive Bayes (GNB) classifier and Linear Support Vector Machine (LSVM).
-This finding suggests that the interactions between different brain areas contain information about the cognitive state, and that this functional connectivity generalises to an extent to different subjects.
+This finding suggests that the interactions between different brain areas contain information about the cognitive state, and that these functional connectivity generalizes to an extent to different subjects.
 
@@
 Furthermore, it avoids a strict multiple comparisons correction that is required in null hypothesis testing by using cross-validation.
-Also, the identification of patterns of activity associated with tasks is encouraged, rather than attempting to identifying brain regions that are involved in tasks.
+Also, the identification of patterns of activity associated with tasks is encouraged, rather than attempting to identify brain regions that are involved in tasks.
 In doing this, much can be learned about how information is represented and processed in the brain.

### 2012-11-12 article/article.tex
review

@@
 %% indicate IEEE Member or Student Member in form indicated below
-\author{Mattijs Meiboom, Robin Mills, Tijn Schouten, Spyros Ioakeimidis}
+\author{Mattijs Meiboom, Robin Mills, Spyros Ioakeimidis, Tijn Schouten}
 \authorfooter{
@@
 Still, we consistently found that the Multivariate Gaussian (MVG) classifier outperformed the Gaussian Naive Bayes (GNB) classifier and Linear Support Vector Machine (LSVM).
-This finding suggests that the interactions between different brain areas contain information about the cognitive state, and that these functional connectivity generalizes to an extent to different subjects.
+This finding suggests that the interactions between different brain areas contain information about the cognitive state, and that this functional connectivity generalises to an extent to different subjects.
 

### 2012-11-12 article/article.tex
review
nonono

@@
 Columns depict the used classifiers.
-For the classifiers depicted with *, we used a different method for structuring the data, as described in Section~\ref{sec:trainingMethodsAndClassifiers}.
+For the classifiers depicted with *, we used the method of estimating the parameters per scan, as described in Section~\ref{sec:trainingMethodsAndClassifiers}.
 

### 2012-11-12 src/demo/pca_by_coord.py
renamed files

@@ -1,186 +1,0 @@
-from __future__ import division
-import scipy.io as sio
-# import pylab
-from scipy import std as sstd
-from scipy import mean as smean
-from functions import *
-from settings import ORIG_FILES
-from sklearn.naive_bayes import GaussianNB
-#from sklearn.preprocessing import normalize
-from math import fsum
-
-# Data
-print "Loading subjects..."
-subjects = [sio.loadmat(ORIG_FILES['Orig'] % (index + 1)) for index in range(6)]
-print "Loaded subjects!"
-
-# Configuration
-FIRST_STIMULUS_SCAN_INDICES = range(10, 20)
-SECOND_STIMULUS_SCAN_INDICES = range(27, 37)
-COMPONENTS = 148
-
-# Rois are ignored if filter coords is false
-#ROIS = ['CALC', 'LIPL', 'LT', 'LTRIA', 'LOPER', 'LIPS', 'LDLPFC']
-ROIS = []
-
-FLAG_PCA = True
-FLAG_NORM = False
-FLAG_FILTER_COORDS = True
-
-# If true, average the prediction per set of scans
-FLAG_PER_SCAN = False
-
-# Just some debugging output
-# for coord in validCoords:
-#     print "Regions for coord (%s,%s,%s)" % (coord[0], coord[1], coord[2])
-#     for subject in subjects:
-#         voxel_index = subject['meta']['coordToCol'][0][0][coord[0]][coord[1]][coord[2]]
-#         print subject['meta']['colToROI'][0][0][voxel_index - 1]
-
-'''
-PCA ALGORITHM
-1) Apply filtering on voxels (for each subject)
-2) Construct data matrix for PCA:
-- each row is a scan from the range of applicable scans (split for type of trial - PS, SP, PS+SP?)
-- each column is a value for a certain coordinate
-3) PCA
-4) Split matrix based on subject
-5) Use one part for training
-6) Use the other part for testing
-'''
-# 1: Get list of coordinates
-validCoords = []
-trials = getValidTrialIndices(subjects[0])
-if FLAG_FILTER_COORDS:
-    print "Constructing nonempty coordinates across subjects"
-    coords = getCoordinatesForSubject(subjects[0], ROIS)
-    print len(coords)
-    for coord in coords:
-        voxels = getVoxelsForCoordinate(coord[0], coord[1], coord[2], subjects)
-        if voxels:
-            validCoords.append(coord)
-    if len(validCoords) == 0:
[…]
